### 2013-02-20 ISMB/conclusion_ISMB.tex
ismb tex

%!TEX root = main_ISMB.tex
\section{Conclusion}

\label{sec:conclusion}

In this article, we described a novel algorithm, \ourprog, for the RNA design problem, i.e. the design of an RNA sequence adopting a predefined secondary structure as minimal free-energy fold.
Implementing a global sampling approach, it optimizes affinity towards the target secondary structure, while granting the user full control over the \GCContent of the resulting sequences.
This extended control does not necessarily induce additional computational demands, and we showed the linear complexity of both the preprocessing stage and the generation of candidate sequences for the design, allowing for the design of larger and more complex secondary structures in a matter of minutes on a single processor (e.g. $\sim$28 mins for 100 candidate sequences for a $\sim$1500nts 16s rRNA). We evaluated the method on a benchmark composed of target secondary structures extracted from the \RNASTRAND database. We observed good overall success rate, with the notable exception of very low targeted \GCContent ($10\%$), and a good to excellent entropy within designed candidates.
Finally, we implemented an hybrid approach, using the \RNAinverse software as a post-processing step for unpaired regions. This approach greatly increased the success rate of the method, allowing for the design of highly diverse candidates for almost all of the structures in our benchmark, while largely preserving the targeted \GCContent.

In the future, we would like to complement this study by further investigating the potential of hybrid local/global -- or {\em glocal} -- approaches.
A global sampling approach would capture the positive aspects of design, optimizing affinity towards a given structure while allowing the specification of expressive systems of constraints.
Designed sequences would serve as a seed for a restricted local approach which, by breaking unwanted symmetries, would perform the negative part of the design, 
while ideally maintaining obedience to the constraints. Another perspective of this work is the incorporation of the full Turner energy model, which should in principle yield better designs for unpaired regions.

### 2013-02-20 ISMB/introduction_ISMB.tex
ismb tex

%!TEX root = main_ISMB.tex
\section{Introduction}
\label{sec:introduction}

At the core of the emerging field of synthetic biology resides our capacity to design and re-engineer molecules with target functions. RNA molecules are well tailored for such applications. The ease to synthesize them (they are directly transcribed from DNA) and the broad diversity of catalytic and regulation functions they can perform enable to integrate \textit{de-novo} logic  circuits within living cells \cite{Rodrigo:2012fk} or re-program existing regulation mechanisms \cite{Chang:2012uq}. Future advances and applications of these techniques in gene-therapy studies will strongly rely on efficient computational methods to design and re-engineer RNA molecules.

Most of RNA functions are, at least partially, encoded by the three-dimensional molecular structures, which are themselves primarily determined by the secondary structures. The development of efficient algorithms for designing RNA sequences with pre-defined secondary structures is thus a milestone to enter the synthetic biology era. \RNAinverse pioneered RNA secondary structure design algorithms. It has been developed and distributed with the Vienna RNA package \cite{Hofacker:1994}. However, only posterior experimental studies revealed the potential and practical impact of these techniques. Thereby, during the last 6 years many improvements and variants of \RNAinverse have been proposed. Conceptually, almost all of existing algorithms follow the same approach. First a seed sequence is selected, then a local search strategy is used to mutate the seed and find, in its vicinity, a sequence with desired folding properties. Using this strategy, \INFORNA \cite{Busch:2006uq}, \RNASSD \cite{Aguirre-Hernandez:2007kx} and \NUPACK \cite{Zadeh:2011fk} significantly improved the performance of RNA secondary structure design algorithms. More recent research studies aimed to include more constraints in the selection criteria. \RNAexinv focused on the design of sequences with enhanced thermodynamical and mutational robustness \cite{Avihoo:2011fk}, while \frankenstein enables to design RNA with multiple target structures \cite{Lyngso:2012vn}.

We recently introduced with \RNAensign a novel paradigm for the search strategy of RNA secondary structure design algorithm \cite{Levin:2012kx}. Instead of a local search approach, we proposed a global sampling strategy of the mutational landscape based on the \RNAmutants algorithm~\cite{Waldispuhl2008}. This methodology offered promising performances, but suffered from  prohibitive runtime and memory consumption. Following our work, Garcia-Martin and co-workers proposed \RNAiFOLD, an alternate methodology that uses constraint programming techniques to prune the mutational landscape. While also suffering from prohibitive running times, it is worth noting that this latter algorithm also proposes a seed-less approach to the RNA secondary structure design problem.

In this paper, we introduce \ourprog, a RNA secondary structure design algorithm that benefits of our recent algorithmic advances \cite{Reinharz:2013aa} to expand our original \RNAensign algorithm \cite{Levin:2012kx}. \ourprog addresses previous limitations of \RNAensign and offers new functionalities. First, while our previous program had a running time complexity of $\mathcal{O}(n^5)$, \ourprog now runs in linear-time and space complexity, allowing it to demonstrate similar speeds as any local search algorithm. Next, \ourprog is \textit{seed-less}. Unlike \RNAensign, it does not require a seed sequence to initiate its search. Finally, \ourprog implements a novel algorithm using weighted sampling techniques~\cite{Bodini2010} that enables us to control, for the first time, \textit{explicitly} the \GCContent of the solution. This functionality is essential because wild sequences within living organisms often present medium or low \GCContent, presumably to offer better transcription rates and/or structural plasticity. Previous programs do not allow to control this parameter and tend to output sequences with high \GCContent{}s.

We demonstrate the performance of our algorithms on a set of real RNA structures extracted from the \RNASTRAND database \cite{andronescu2008rna}. To complete this study, we develop an hybrid method combining our global sampling approach with local search strategies such as the one implemented in \RNAinverse.  Remarkably, our glocal methodology overcomes both local and global approaches  for sampling sequences with a specific GC content and target structure.

### 2013-02-20 ISMB/main_ISMB.tex
ismb tex

\documentclass[11pt]{article}
\usepackage[top=1in, bottom=1in,right=1in,left=1in]{geometry}                % See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   % ... or a4paper or a5paper or ... 
%\geometry{landscape}                % Activate for for rotated page geometry
%\usepackage[parfill]{parskip}    % Activate to begin paragraphs with an empty line rather than an indent
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm} %pour les newtheorem
\usepackage{epstopdf}
\usepackage{xspace}
\usepackage{tcolorbox}
\usepackage{relsize}
\usepackage[numbers]{natbib}
\usepackage[colorinlistoftodos]{todonotes}

%\usepackage{subfig} %Subfigures
\usepackage{subcaption} %Subfigures

\usepackage{tikz}
\usepackage{amsmath}
\usepackage{ifthen}
\usepackage{relsize}
\usetikzlibrary{positioning}
\usetikzlibrary{matrix}
\usetikzlibrary{calc}
\usetikzlibrary{shapes}
\usetikzlibrary{arrows}
\usetikzlibrary{fit}
\usetikzlibrary{decorations}
\usetikzlibrary{snakes}
\pgfdeclarelayer{background}
\pgfdeclarelayer{backbackground}
\pgfdeclarelayer{foreground}
\pgfsetlayers{backbackground,background,main,foreground}

\usepackage[noend,ruled,vlined]{algorithm2e}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}

\newcommand{\ourprog}{\texttt{IncaRNAtion}\xspace}

%\usepackage[applemac]{inputenc} %for the encoding

\newcommand{\RNAmutants}{\texttt{RNAmutants}\xspace}
\newcommand{\RNApyro}{\texttt{RNApyro}\xspace}
\newcommand{\RNAinverse}{\texttt{RNAinverse}\xspace}
\newcommand{\RNASSD}{\texttt{RNA-SSD}\xspace}
\newcommand{\INFORNA}{\texttt{INFO-RNA}\xspace}
\newcommand{\NUPACK}{\texttt{NUPACK}\xspace}
\newcommand{\RNAiFOLD}{\texttt{RNAiFOLD}\xspace}
\newcommand{\frankenstein}{\texttt{Frnakenstein}\xspace}
\newcommand{\RNAexinv}{\texttt{RNAexinv}\xspace}
\newcommand{\rnaDesign}{\texttt{rnaDesign}\xspace}
\newcommand{\RNAensign}{\texttt{RNA-ensign}\xspace}
\newcommand{\GCContent}{\Gb\Cb-content\xspace}
\newcommand{\SFold}{\texttt{SFold}\xspace}

\newcommand{\red}[1]{{\color{red}#1}}
\newcommand{\farna}{\texttt{FARNA}\xspace}
\newcommand{\mcfoldmcsym}{\texttt{MC-Pipeline}\xspace}
\newcommand{\mcfold}{\texttt{MC-Fold}\xspace}
\newcommand{\mcsym}{\texttt{MC-Sym}\xspace}
\newcommand{\nast}{\texttt{NAST}\xspace}
\newcommand{\ifoldrna}{\texttt{iFoldRNA}\xspace}
\newcommand{\rnafold}{\texttt{RNAfold}\xspace}
\newcommand{\rnasubopt}{\texttt{RNAsubopt}\xspace}
\newcommand{\rnawolf}{\texttt{RNAwolf}\xspace}
\newcommand{\rnastructure}{\texttt{RNAstructure}\xspace}
\newcommand{\contrafold}{\texttt{contrafold}\xspace}
\newcommand{\unafold}{\texttt{unafold}\xspace}
\newcommand{\rnadd}{\texttt{RNA2D3D}\xspace}
\newcommand{\assemble}{\texttt{assemble}\xspace}
\newcommand{\fred}{\texttt{FR3D}\xspace}
\newcommand{\rnajunction}{\texttt{RNAjunction}\xspace}
\newcommand{\rnamotif}{\texttt{RNAmotif}\xspace}
\newcommand{\treefolder}{\texttt{TreeFolder}\xspace}
\newcommand{\barnacle}{\texttt{BARNACLE}\xspace}
\newcommand{\contextfold}{\texttt{contextfold}\xspace}

\newcommand{\RNASTRAND}{{\sf RNA STRAND}\xspace}

%\newcommand{\citep}{\cite}
\newcommand{\Z}[2]{\mathcal{Z}{\substack{(#1)\\ [#2]}}}
\newcommand{\Y}[3]{\mathcal{Y}_{\substack{(#1)\\ [#3]}}^{#2}}
\newcommand{\B}{\mathcal{B}}
\newcommand{\Kron}{\delta}
\newcommand{\ub}{\bullet}

\newcommand{\Ab}{{\sf{A}}\xspace}
\newcommand{\Cb}{{\sf{C}}\xspace}
\newcommand{\Gb}{{\sf{G}}\xspace}
\newcommand{\Ub}{{\sf{U}}\xspace}

\newcommand{\gc}{{\#\text{{\sf gc}}}}
\newcommand{\PE}[1]{E(#1)}
\newcommand{\EI}{\text{EI}}
\newcommand{\ES}{\text{{\sf Es}}}
\newcommand{\ISO}{\text{ISO}}
\newcommand{\Prob}{\text{P}}
\newcommand{\Target}{\text{\sf S}}
\newcommand{\CNBP}[1]{\gamma(#1)}

\newcommand{\CFramed}[2][gray]{\begin{tcolorbox}[colframe=#1,title=To be included only if space allows\ldots,colback=white]#1\end{tcolorbox}}
\renewcommand{\CFramed}[2][gray]{}

\title{A weighted sampling algorithm for the design of RNA sequences with targeted secondary structure and nucleotides distribution}
\author{Vladimir Reinharz$^1$, Yann Ponty$^{2,*}$, J\'er\^{o}me Waldisp\"{u}hl$^{1,*}$}
\date{$^1$ School of Computer Science, McGill University, Montreal, Canada, $^2$ Laboratoire d'informatique, \'Ecole Polytechnique, Palaiseau, France.\\ \small $^*$Corresponding authors: \texttt{jeromew@cs.mcgill.ca}, \texttt{yann.ponty@lix.polytechnique.fr}}

%Last thing before BEGIN DOC for links to appear in pdf

%  \usepackage{hyperref}
%  \hypersetup{colorlinks}
%              citecolor=black,
%              filecolor=black%
%              linkcolor=black%
%              urlcolor=black%
%              pdftex}

\begin{document}
\maketitle
%\section{}
%\subsection{}
\begin{abstract}
\noindent
\textbf{Motivations:} The design of RNA sequences folding into predefined secondary structures is a milestone for many synthetic biology and gene therapy studies. Most of the current software use similar local search strategies (i.e. a random seed is progressively adapted to acquire the desired folding propertues) and more importantly do not allow the user to control explicitly the nucleotide distribution such as the \GCContent in their sequences. However, the latter is an important criteria for large-scale applications as it could presumably be used to design sequences with better transcription rates and/or structural plasticity.\\
\noindent
\textbf{Results:}
In this paper, we introduce \ourprog, a novel algorithm to design RNA sequences folding into target secondary structures with a predefined nucleotide distribution.  \ourprog uses a global sampling approach and weighted sampling techniques. We show that our approach is fast (i.e. running time comparable or better than local search methods), seed-less (we remove the bias of the seed in local search heuristics), and successfully generates high-quality sequences (i.e. thermodynamically stable) for any \GCContent. To complete this study, we develop an hybrid method combining our global sampling approach with local search strategies. Remarkably, our glocal methodology overcomes both local and global approaches  for sampling sequences with a specific GC content and target structure.\\
\noindent
\textbf{Key words:} RNA, secondary structure, design, weighted sampling, GC-content.
\end{abstract}

\newpage

\input{introduction_ISMB.tex}

\input{methods_ISMB.tex}

\input{results_ISMB.tex}

\input{conclusion_ISMB.tex}

\input{acknowledgments_ISMB.tex}

\newpage

\bibliographystyle{plainnat}
\bibliography{RNApyro}

\newpage

\input{supplementary_ISMB.tex}

\end{document}

### 2013-02-20 ISMB/methods_ISMB.tex
ismb tex

%!TEX root = main_ISMB.tex
\section{Methods}
\label{sec:methods}

We introduce a probabilistic model for the design of RNA sequences with a specific \GCContent and folding into a predefined secondary structure.
For the sake of simplicity, we choose to base this proof-of-concept implementation on a simplified free-energy function $\ES(\cdot)$, which only considers the contributions of 
stacked canonical base-pairs. We show how a modification of the dynamic programming scheme used in the \RNAmutants allows for the sampling of good and diverse design candidates, in linear time and space complexities.

%To that purpose, a Boltzmann weighted distribution is used, based on a pseudo-energy function $\PE{\cdot}$ which includes contributions for both the free-energy and its putative isostericity towards a multiple sequence alignment. In this model, the probability that the nucleotide at a given position needs to be mutated (i.e. corresponds to a sequencing error) can be computed using a variant of the \emph{Inside-Outside algorithm}~\cite{Lari1990}.
%
%\subsection{Probabilistic model}
%Let $\Omega$ be an gap-free RNA alignment sequence, $S$ its associated secondary structure, 
%then any sequence $s$ is assigned a probability proportional to its Boltzmann factor
%\begin{align*}
%  \mathcal{B}(s) &= e^\frac{-\PE{s}}{RT}, &&\text{with}&\PE{s}&:=\alpha\cdot\ES(s,S)+(1-\alpha)\cdot\EI(s,S,\Omega),
%\end{align*}
%where $R$ is the Boltzmann constant, $T$ the temperature in Kelvin, $\ES(s)$ and $\EI(s,S,\Omega)$ 
%are the free-energy and isostericity contributions respectively (further described below), and $\alpha\in[0,1]$ is an arbitrary parameter that sets the relative weight for both contributions.
%
%\subsubsection{Energy contribution}
\subsection{Definitions}

A targeted secondary structure $\Target$ is abstracted as a non-crossing arc-annotated sequence,  where 
$\Target_i$ stands for the base-pairing position of position $i$ in $\Target$ if any (and, reciprocally, $\Target_{\Target_i}=i$), or $-1$ otherwise. 
Let us denote by $\gc(s)$ the number of occurrences of \Gb and \Cb in an RNA sequence $s$.

\subsubsection{Simplified energy model}
The free-energy contribution in our pseudo-energy model corresponds to an additive stacking-pairs model, using individual values from the Turner 2004 model (retrieved from the NNDB~\cite{Turner2010}). Given a candidate sequence $s$ for a secondary structure $S$, the free-energy of $S$ on $s$ is given by
\begin{align*}
  \ES(s,S) = \sum_{\substack{(i,j)\to (i',j')\in S\\ \text{stacking pairs}}}\ES^{\beta}_{s_is_j\to s_{i'}s_{j'}} 
\end{align*}
where $\ES^{\beta}_{ab\to a'b'}$ is set to $0$ if $ab=\varnothing$ (no base-pair to stack onto), the tabulated free-energy of stacking pairs $(ab)/(a'b')$ in the Turner model if available, or $\beta\in[0,\infty]$ for non-Watson-Crick/Wobble pairs (i.e. not in $\{\Gb\Ub,\Ub\Gb,\Cb\Gb,\Gb\Cb, \Ab\Ub\text{ or }\Ub\Ab\}$). This latter parameter allows one to choose whether to simply penalize invalid base pairs ($\beta>0$), or forbid them altogether ($\beta = \infty$).

\subsubsection{\GCContent weighted pseudo-Boltzmann ensemble and distribution}

In order to counterbalance the documented tendency of sampling methods to generate \Gb\Cb-rich sequences~\cite{Levin:2012kx}, let us introduce a parameter $x\in\mathbb{R}^+$, whose value will influence the \GCContent of generated sequences. Let $\Target$ be a targeted secondary structure, then let us define the pseudo-Boltzmann factor $\B_{x}(s,S)$ of a sequence $s$ such that
\begin{equation}
\B_{x, \Target}(s) = e^{\frac{-\ES(s,\Target)}{RT}}\cdot x^{\gc(s)}
\label{def:genBoltz}
\end{equation}
where $R$ is the Boltzmann constant and $T$ the temperature in Kelvin.

Summing the pseudo-Boltzmann factor over all possible sequences of a given length $|\Target|$ gives the pseudo-partition function $\mathcal{Z}_{x,\Target}$, from which one defines the pseudo-Boltzmann probability $\Prob_{x,\Target}(s)$ of each sequence $s$, respectively such that 
\begin{align}\mathcal{Z}_{x,\Target} &= \sum_{\substack{|s|=|\Target|}}\B_{x,\Target}(s)& \text{and}&&
\Prob_{x,\Target}(s) &= \frac{\B_{x,\Target}(s)}{\mathcal{Z}_{x,\Target}}.\label{def:distribution}\end{align}

\subsection{Sampling the pseudo-Boltzmann ensemble}

Let us now describe a linear-time algorithm to sample sequences at random in the pseudo-Boltzmann distribution. This algorithm follows the general principles of the recursive approach to random generation~\cite{Wilf1977}, pioneered in the context of RNA by the \SFold algorithm~\cite{Ding2003}. The algorithm starts by precomputing the partition function restricted to each valid interval, and then performs a series of recursive stochastic backtracks, using precomputed values to decide on the probability of each alternative.

\subsubsection{Precomputing the pseudo-partition function}\label{sec:pf}
\begin{figure}
\resizebox{\textwidth}{!}{\input{FigStochasticBacktrack.pgf}}
\caption{Three main cases in the stochastic backtrack procedure over an interval $[i,j]$: Either position $i$ is left unpaired (top), a base-pair $(i,j)$ is formed, stacking onto $(i+1,j-1)$ (middle), or $i$ is paired somewhere else, defining two regions on which subsequent recursive calls are needed (bottom). Positions indicated in red are assigned during this stage of the backtrack.\label{fig:stochastic}}
\end{figure}

Firstly, a dynamic programming algorithm computes $\mathcal{Z}\substack{(i,j)\\ [a,b]}$ the pseudo-partition function restricted to an interval $[i,j]$, assuming its (previously chosen) flanking nucleotides are $a$ and $b$ respectively. 
Since the only sequence over an empty interval is the empty sequence, having energy $0$, one has
\begin{equation}
	\forall i \in [0,n-1]:\, \Z{i+1,i}{a,b}=1.
	\label{eq:Z_in}
\end{equation}

The recursion consists in four terms, depending on the base-pairing status and context of $(i,j)$:
\begin{equation}
	\Z{i,j}{a,b}:=\left\{
  \begin{array}{ll}
  		\displaystyle
      \sum_{a'\in \B}  
      x^{\gc(a')}
      \cdot\Z{i+1,j}{a',b} &\begin{array}{@{}l@{}}\text{If }\Target_{i}=-1\\ \text{\relsize{-1}Position $i$ unpaired,}\end{array}\\
      \displaystyle
      \sum_{a',b'\in \B^2}
			 x^{\gc(a'.b')}
			 \cdot e^{\frac{-\ES^{\beta}_{ab \to a'b'}}{RT}}
			 \cdot \Z{i+1,j-1}{a',b'}&
\begin{array}{@{}l@{}}\text{Elif }\Target_i=j \land \Target_{i-1}=j+1\\ \text{\relsize{-1}$(i,j)$ stacking pair onto $(i$-$1,j$+$1)$,}\end{array}
			 \\
			 \displaystyle
      \sum_{a',b'\in \B^2}
      x^{\gc(a'.b')}
			\cdot e^{\frac{-\ES^{\beta}_{\varnothing\to a'b'}}{RT}}
      \cdot\Z{i+1,k-1}{a',b'}
      \cdot\Z{k+1,j}{b',b} & 
\begin{array}{@{}l@{}}\text{Elif }\Target_i=k \land i < k \leq j\\
\text{\relsize{-1}Junction/bulge $\Rightarrow$ split, no stack,}
\end{array}
\\
      0 &\text{Otherwise.}
	\end{array}\right.
\label{eq:Z_rec}
\end{equation}

Despite its apparent quadratic complexity, the recurrence described in Subsection~\ref{sec:pf} can be computed in linear $\Theta(|\Target|)$ time and space, owing to the fact that, on any recursive call, the position $j$ is entirely determined by $i$.

\begin{theorem}Starting from the complete region $[0,n-1]$, Equation~\eqref{eq:Z_rec} can be computed in $\Theta(|\Target|)$ time and space complexity.
\end{theorem}
\begin{proof}

Consider the {\bf covering base-pair} $\CNBP{i}$ of a position $i$, defined as the base-pair $(i',j')$ such that $i'<i<j'$ and which minimizes $i-i'$. For the sake of simplicity, and without loss of generality, let us assume that there exists a virtual base-pair $(-1,n)$ in $\Target$, so that any position in $[0,n-1]$ has a covering base-pair.
Then, it is easily shown by induction that, starting from the complete interval $[0,n-1]$, the recursive calls only evaluate $\mathcal{Z}$ over intervals $[i,j]$ such that $j=j'-1$, $\CNBP{i}=[i',j']$. 
\CFramed{This property clearly holds initially for the complete interval $[0,n-1]$, owing to the virtual base-pair $(-1,n)$. Furthermore, if it holds for a given $(i,j)$, $\CNBP{i}=[i',j+1]$, then:
\begin{itemize} 
  \item If $i$ is unpaired (Case 1), then $\CNBP{i+1}=\CNBP{i}=[i',j+1]$, and the property is satisfied by the subsequent call over $[i+1,j]$. 
  \item If $i$ is paired with $j$ and stacking onto $(i-1,j+1)$ (Case 2), then one has $\CNBP{i+1}=[i,j]$ and the  subsequent call over $[i+1,j-1]$ satisfies the property.
  \item If $i$ is paired with some $k$ (Case 3), then $\CNBP{i+1}=[i,k]$ and $\CNBP{k+1}=[i',j+1]$. Therefore the subsequent calls, respectively over $[i+1,k-1]$ and $[k+1,j]$, both satisfy the property.
\end{itemize}
The case of an empty sequence does not cause a recursive call and must not be verified.
}
Since, in any interval $[i,j]$ satisfying the property, the ending position $j$ is completely determined by $i$, then there exists at most $\Theta(|\Target|)$ intervals that need to be evaluated, and we conclude on the overall complexity with the remark that $a$, $b$, $a'$ and $b'$ vary over bounded ranges and only contribute to a constant factor in the complexity.
\end{proof}

\subsubsection{Stochastic backtrack}
Once the pseudo-partition functions have been memorized, a stochastic backtrack starts from the complete interval $[0,|\Target|-1]$ and, at each step, chooses a suitable assignment for one or several positions, using probabilities derived from the precomputation. One or several recursive calls over regions that are left unassigned are then performed, as described by Algorithm~\ref{alg:back}, and illustrated by Figure~\ref{fig:stochastic}.
\begin{algorithm}[t]
\DontPrintSemicolon
	\SetAlgoLined
\SetKwFunction{Backtrack}{Backtrack}
\SetKwFunction{Random}{Random}
	rand $\leftarrow$ \Random$\left(\Z{i,j}{a,b}\right)$\tcp*[r]{Draw random number in $[0,\Z{i,j}{a,b}[$}
 \lIf(\tcp*[f]{Empty sequence}){$j<i$}{\Return{$\varepsilon$}}
	\Else{
  k $\leftarrow S_i$\;
	\If(\tcp*[f]{Unpaired}){$(k = -1)$}{
		\For{$a'\in\B$}{
			rand $\leftarrow$ rand $- x^{\gc(a')}\cdot \Z{i+1,j}{a',b}$\;
			\lIf{$\text{rand}<0$}\Return{$a'.\Backtrack\left(i+1,j,a',b,\Target\right)$}\;
		}
	}
	\lElse{
		\If(\tcp*[f]{Stacking pairs}){$(k=j) \land (S_{i-1}=j+1)$}{
			\For{$(a',b')\in\B\times\B$}{
				rand $\leftarrow$ rand $-
			 	x^{\gc(a'.b')}
			 	\cdot e^{\frac{-\ES^{\beta}_{ab \to a'b'}}{RT}}
			 	\cdot \Z{i+1,j-1}{a',b'}	$\;
				\lIf{$\text{rand}<0$}{
					\Return{$a'.\Backtrack\left(i+1,j-1,a',b',\Target\right). b'$}\;		
				}
			}
		}
		\lElse{
			\If(\tcp*[f]{Paired, not stacked}){$(\Target_i=k) \land (i < k \leq j)$}{
				\For{$(a',b')\in\B\times\B$}{
					rand $\leftarrow$ rand $-	
       x^{\gc(a'.b')}
					\cdot e^{\frac{-\ES^{\beta}_{\varnothing\to a'b'}}{RT}}
  	    	\cdot\Z{i+1,k-1}{a',b'}
	    	  \cdot\Z{k+1,j}{b',b}$\;	
 					\If{$\text{rand}<0$}{
						\Return{$a'
						.\Backtrack\left(i+1,k-1,a',b',\Target\right)
						.b'
						.\Backtrack\left(k+1,j,b',b,\Target\right)
						$}\;	
					}
				}
		 }
		}
	}
 }
\caption{\protect\Backtrack$\left(i,j,a,b,\Target\right)$\label{alg:back}}
\end{algorithm}

Since, on each recursive call, the algorithm assigns at least one nucleotide to a -- previously unassigned -- position, and that the number of executions of the loops are bounded by a constant, then the complexity of the algorithm grows linearly on the targeted length.

\subsubsection{Rejecting unsuitable candidates}

Note that, although the parameter $x$ is introduced to match the expected \GCContent with the targeted one, nothing prevents Algorithm~\ref{alg:back} to generate sequences of arbitrary \GCContent. Therefore, we use a rejection-based approach~\cite{Bodini2010} previously used by the authors in a similar context~\cite{Waldispuhl2011}.

It was established~\cite{Waldispuhl2011} that, for each value of $x$, there exists constants $\mu_x$ and $\sigma_x$ such that the \GCContent distribution observed in sequences sampled with respect to the above distribution converges asymptotically towards a normal law having expectation in $\mu\cdot n\cdot(1+o(1))$ and standard deviation in $\sigma\cdot\sqrt{n}\cdot(1+o(1))$.
Furthermore, it can easily be shown that the expected \GCContent is a continuous and strictly increasing monotonic  function of $x$, whose limits are $0$ when $x=0$ and $n$ when $x\to +\infty$.

Consequently, for any targeted \GCContent $gc\in[0\%,100\%]$, there exists a unique value $x_{gc}$ such that generated sequences feature, on the average, the right \GCContent. Such a value can be determined using any numerical recipe, and a simple bisection procedure, further described in earlier work~\cite{Waldispuhl2011}, is used in our implementation. Furthermore, the concentration of the distribution, asserted by its limited standard deviation, implies a growth in $\Theta(\sqrt{n})$ (resp. $\Theta(1)$, i.e. constant) for the expected number of attempts before generating a sequence having the exact targeted \GCContent  (resp. within $[gc-\kappa,gc+\kappa]$, where $\kappa\in\Theta(1/\sqrt n)$).

\subsection{Postprocessing unpaired regions: A local/global (glocal) hybrid approach}
\label{subsec:glocal_method}
Due to our simplified our energy model, unpaired regions are not subject to design constraints other than the \GCContent, leading to modest probabilities for refolded design candidates to match the targeted structure. To improve these performances and test the complementarity  of our global sampling approach with previous contributions based on local search, we used the \RNAinverse software to redesign unpaired regions. We specified a constraint mask to prevent stacking base-pairs from being modified and, whenever necessary, reetablished their content content a posteriori as \RNAinverse has been witnessed to take some liberties with constraints masks. As shown in Table~\ref{table:impact_on_gc} (Supplementary material), this postprocessing does not drastically alter the \GCContent, so the glocal approach reasonably addresses the contrained \GCContent design problem.

### 2013-02-20 ISMB/results_ISMB.tex
ismb tex

%!TEX root = main_ISMB.tex
\section{Results}
\label{sec:results}

\subsection{Implementation}
\label{sec:implementation}
Our software, \ourprog, was implemented in {\tt Python~2.7}. We used
\RNAinverse from the \textit{Vienna Package 2.0}~\cite{Hofacker:1994}.
All time benchmarks were run on a one AMD Opteron(tm) Processor 6278  at 2.4 GHz with cache of 512 KB.
 We used as parameter $\beta=15$, the penalty for an invalid base pair.

We present the average times spent by \ourprog and \RNAinverse to generate one sequence
with the required \GCContent in Figure~\ref{fig:time}. As expected, the time grows linearly
in function of the length of the structures for \ourprog.  A highly interesting feature
is the huge decrease in time of computation when there is $15\%$ free 
nucleotides inside the loops of the secondary structure.

\begin{figure}[ht!]
	\centering
	\includegraphics[width=\textwidth]{Figures/time_rnastrand_clustered_rnainverse_100samples_fix}
	\caption{Average time in seconds to generate one sequence. We explicitly show 
	the time spent by \ourprog (full line) and \texttt{RNAinverse} (dashed) for various \GCContent. The first plot is in function
	of the length of the structures. The second is in function of the stack
	density (i.e. $2\cdot\#stacks/length$) and the latter in function of 
	the proportion of free nucleotides (i.e. unpaired bases) within loops.}
	\label{fig:time}	
\end{figure}

\subsection{Dataset}
To evaluate the quality of our method, we used secondary
structures from the \RNASTRAND database~\cite{andronescu2008rna}.
Those are known secondary structure from a variety of organisms.
We select the same subset of $50$ structures as in~\cite{Levin:2012kx}, 
whose length ranges between $20$ and $100$ nucleotides.

 To ease the visualization of the results, we clustered together strucures
 having similar length, stacks density and proportion of free-nucleotides in loops. 
  We present in Figure~\ref{fig:bins} the distribution of structures in each category.

 \begin{figure}[ht!]
 	\centering
	\includegraphics[width=\textwidth]{Figures/bins_distribution.png}
	\caption{Number of secondary structures per bin, according to our three clustering criteria.}
	\label{fig:bins}
 \end{figure}

\subsection{Design}
 To benchmark our method, we first started by sampling about $100$ sequences per structure. After correcting the free nucleotides with
 \texttt{RNAinverse} we computed their MFE with \texttt{RNAfold} from the \textit{Vienna Package 2.0}~\cite{Hofacker:1994}.

{\em Définir critères d'évaluations !!!}

A first criterion is which percentage of sequences have as MFE the target
secondary structures. We are also interested by the number of structures
for which we can design at least one sequence with the desired MFE.
We show the results when using only \ourprog in 
 Figure~\ref{fig:mfe_struct_solved_noinverse}. As mentioned earlier, the high rate of failure
is due to the fact that no selection criterion is applied to
unpaired nucleotides. A local strategy is thus needed.
After processing \ourprog sequences with \RNAinverse to 
locally optimize the free nucleotides, we obtained the results 
shown in Figure~\ref{fig:mfe_struct_solved}. We observed
that length is, in general, not a good predictor for the hardness of designing a structure. 
Instead, a high number of free nucleotides in the structure seems to be a 
good measure of its design hardness. 
 The difficulty of 
designing sequences for targets with a high number of free nucleotides 
 in loops appears clearly in the last column of Figure~\ref{fig:mfe_struct_solved}.
Even with a \GCContent of $50\%$ and more, when at least
one solutions was found for almost all structures, most of the samples 
do not fold into the target. A few structures remained unsolved under
all \GCContent.

\begin{figure}[ht!]	
	\centering
	\includegraphics[width=\textwidth]{Figures/mfe_struct_solve_nornainverse.png}
	\caption{Quality of \ourprog results. The first row shows the percentage
	of sampled sequences folding into the target. The second shows the 	
	proportion	of structures for which at least one correct sequence was 
	sampled.}
	\label{fig:mfe_struct_solved_noinverse}	
\end{figure}

\begin{figure}[ht!]	
	\centering
	\includegraphics[width=\textwidth]{Figures/mfe_struct_solved}
	\caption{Quality of \texttt{Incarnation+RNAinverse} results. The first row shows the percentage
	of sampled sequences folding into the target. The second shows the 	
	proportion	of structures for which at least one correct sequence was 
	sampled.}
	\label{fig:mfe_struct_solved}	
\end{figure}

To evaluate the global quality of \ourprog sequences, we show
in Figure~\ref{fig:ss_sens} the ratio of well predicted base pairs in the
MFE structure of our sampled sequences. We can observe that, in all cases,
the hardest sequence to design are those with an extremely low \GCContent. As anticipated, those are the sequences with the weakest bonds.
We notice that the most accurate sequences yield a \GCContent
of $70\pm 10\%$.

As discussed in Sec.~\ref{sec:implementation}, we notice a highly decreased
computational time needed to generate the sequences with $15\%$ free 
nucleotides in the loops. We remark that those structures also yield 
a much lower structural sensitivity.

\begin{figure}[ht!]
 	\centering
	\includegraphics[scale=0.45]{Figures/rnastrand_clustered_rnainverse_100samples_struct_sens.png}
	\caption{Structural sensitivity (i.e. $\#$ well predicted base pairs / $\#$ base pairs in target) of the sampled sequences MFE. }
	\label{fig:ss_sens}	
\end{figure}

\subsection{Samples properties}

In this section, we further analyze the generated sequences that fold into the 
target structure.

A desirable feature in sequence design, is to produce samples with a high
diversity. Figure~\ref{fig:nb_sols_entropy} shows the number of correct
solutions as their average entropy and base pair entropy, since 
\ourprog  only influences the distribution of nucleotides inside 
stacked base pairs. The most constraining factor for generating valid
 solutions seems to be  low \GCContent and the percentage of free nucleotides.

%Our method is able to generate

Also, a critical properties in RNA sequence design is 
the frequency of the MFE. 
The sequence should be in its target conformation long enough to
perform the desired action. Presented in Figure~\ref{fig:freq} we see that
there is a slow decline of the frequency with the increase in size. Yet,
for an average \GCContent, the frequency stays over $10\%$ even
at size of $100$ nucleotides.

\begin{figure}[ht!]
	\centering
	\includegraphics[width=\textwidth]{Figures/nb_sols_entropy.png}
	\caption{Number of solutions generated and their average entropy. 
	The last row presents only the entropy of the nucleotides inside base 
	pairs.}
	\label{fig:nb_sols_entropy}
\end{figure}

\begin{figure}[ht!]
	\centering
	\includegraphics[width=\textwidth]{Figures/freq.png}
	\caption{Average Boltzmann probability of the MFE.}
	\label{fig:freq}
\end{figure}

\subsection{Global sampling vs Local search vs Glocal approach}

Here, we are interested in estimating the impact of the design methodology on the performances. More precisely, we aim to determine the merits of a global sampling approach (\ourprog), compared to a glocal procedure (\ourprog + \RNAinverse) and a local search methodology (\RNAinverse).

Our results (See Fig. \ref{fig:rnainverse} in Supplementary Material) indicate that a glocal approach produces sequences that are consistently more diverse than those obtained with both {\em pure} methods alone. Furthermore, while \ourprog outperforms \RNAinverse for extreme pressures on the \GCContent, \RNAinverse produces more entropic sequences than \ourprog for the central \GCContent regime ($30\%-70\%$).

It should be noted, however, than \RNAinverse produces sequences whose \GCContent is extremely concentrated (see Table~\ref{tab:nb_rnainv} in Supplementary Material), thereby requiring a large number of runs before producing a suitable \GCContent sequence. Therefore, a rejection algorithm based on \RNAinverse alone should not be considered a realistic solution for the design with constrained \GCContent.

{\em Ajouter considérations Base Pair entropy: Se rappeler, que moins, c'est bien !}

% \ourprog outperforms other strategies for non-centered . In particular, the base pair entropy as well as diversity of sequences are favorable to \ourprog. This advantage is even more striking for low targeted \GCContent{}s regimes.
%To complete this benchmark, we added the results obtained with the glocal procedure (\ourprog + \RNAinverse). We note that the results are slightly lower than those of \ourprog alone. However, we must stress that the \ourprog data consider only the sample of \ourprog already satisfying the MFE criterion. In fact \RNAinverse enable to ``correct'' a lot of sample sampled by \ourprog but without the good folding properties. Hence, our conclusion here is that the glocal approach enable us to conserve the same level of performance but to drastically improve the success rate of our methodology.

\label{fig:rnainverse}

### 2013-02-20 ISMB/supplementary_ISMB.tex
ismb tex

%!TEX root = main_ISMB.tex
\section{Supplementary data}
\subsection{Benchmark \ourprog vs \RNAinverse}
Using the same dataset of $50$ structures, we generated $100$ samples
per structure with \RNAinverse. They yield for most parameters
a high entropy.
Table.~\ref{tab:nb_rnainv} contains the number of sample with a given \GCContent generated by \RNAinverse. The vast majority resides around $50\%-70\%$, showing one of its major deficiencies. 
 For all structures that have been solved 
by the three methods, only \RNAinverse, only \texttt{Incarnation} and
\texttt{Incarnation} followed by \RNAinverse, 
we present for every different concentration of \GCContent
the average sample entropy. and the average base pairs entropy in Figure.~\ref{fig:rnainverse}.

\begin{table}[h!]
	\begin{center}
		\begin{tabular}{|c|ccccc|}
		\hline
		Target \GCContent & 10 & 30 & 50 & 70 & 90\\ \hline
   $\#$\RNAinverse samples& 11 & 437 & 3362 & 1155 & 35\\ \hline
		\end{tabular}
	\end{center}
  \caption{Overall \GCContent distribution for sequences designed using \RNAinverse.}
	\label{tab:nb_rnainv}
\end{table}

\begin{figure}[ht!]
	\centering
	%\hspace{-5em}
	\includegraphics[scale=0.4]{Figures/RNAinverse_data_100.png}
	\caption{Entropy and \GCContent  for structures solved by
	the 3 methods.}
	\label{fig:rnainverse}
\end{figure}
{\em Récupérer Base-pair entropy}

%The same analysis but only for structures solved by \texttt{Incarnation}
%and \texttt{Incarnation} followed by \RNAinverse, is presented in 
%Fig.~\ref{fig:inc_rnainv}
%
%\begin{figure}
%	\centering
%	\includegraphics{}
%	\caption{Entropy and \texttt{C+G} content for structures solved by
%	the \texttt{Incarnation} and \texttt{Incarnation}$+$\RNAinverse.}
%	\label{fig:inc_rnainv}
%\end{figure}
\subsection{Impact of local search on the \GCContent of \ourprog output}

These data demonstrate that the local search heuristic used to design nucleotides in loop regions has a very limited impact on the \GCContent. For each class of \GCContent, we show in Table~\ref{table:impact_on_gc}, the observed \GCContent of the sequence sampled by \ourprog, and then we report the observed \GCContent after the \RNAinverse step (as defined in Section \ref{subsec:glocal_method}). Our results show that the \GCContent is relatively well conserved.

\begin{table}[h!]
\begin{center}
\begin{tabular}{|c|c|c|}
\hline
target \GCContent & \ourprog & \ourprog + \RNAinverse \\
\hline
10\% & 0.15 & 0.21\\
30\% & 0.30 & 0.33\\
50\% & 0.48 & 0.49\\
70\% & 0.71 & 0.69\\
90\% & 0.83 & 0.78\\
\hline
\end{tabular}
\end{center}
\caption{Observed \GCContent of solutions returned by \ourprog (2nd column) and after the application of the local search heuristic (3rd column).}
\label{table:impact_on_gc}
\end{table}

### 2013-02-20 src/IncaRNAtion.py

@@ -1,3 +1,29 @@
+###############################################################################
+#Copyright (c) 2013, Vladimir Reinharz, Yann Ponty & Jerome Waldispuhl        #
+#All rights reserved.                                                         #
+#                                                                             #
+#Redistribution and use in source and binary forms, with or without           #
+#modification, are permitted provided that the following conditions are met:  #
+#* Redistributions of source code must retain the above copyright             #
+#notice, this list of conditions and the following disclaimer.                #
+#* Redistributions in binary form must reproduce the above copyright          #
+#notice, this list of conditions and the following disclaimer in the          #
+#documentation and/or other materials provided with the distribution.         #
+#* Neither the name of the <organization> nor the                             #
+#names of its contributors may be used to endorse or promote products         #
+#derived from this software without specific prior written permission.        #
+#                                                                             #
+#THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"  #
+#AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE    #
+#IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE   #
+#ARE DISCLAIMED. IN NO EVENT SHALL <COPYRIGHT HOLDER> BE LIABLE FOR ANY       #
+#DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES   #
+#(INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR SERVICES; #
+#LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION) HOWEVER CAUSED AND  #
+#ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT LIABILITY, OR TORT   #
+#(INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY OUT OF THE USE OF THIS#
+#SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF SUCH DAMAGE                  #
+###############################################################################
 import itertools
 import random
 import math

### 2013-03-01 src/IncaRNAtion.py
iupac code added

@@ -46,7 +46,25 @@ except ImportError:
 
 sys.setrecursionlimit(10000)
 
-BASES = ['A','C','G','U']
+IUPACBASES = {
+  'A':['A'],
+  'C':['C'],
+  'G':['G'],
+  'U':['U'],
+  'N':['A','C','G','U'],
+  'R':['A','G'],
+  'Y':['C','U'],
+  'S':['G','C'],
+  'W':['A','U'],
+  'K':['G','U'],
+  'M':['A','C'],
+  'B':['C','G','U'],
+  'D':['A','G','U'],
+  'H':['A','C','U'],
+  'V':['A','C','G']
+  }
+BASES = []
+
 BOLTZMANN = 0.0019872041
 global T
 T = 310.15
@@ -54,7 +72,7 @@ T = 310.15
 #Just generate all possible combination with maxint
 global STACKING_ENERGY
 STACKING_ENERGY = {k:sys.maxint for k in itertools.product(
-                  BASES, repeat=4)}
+                  IUPACBASES['N'], repeat=4)}
 #Adjust with turner04
 #The order of the nucleotides is from 5' -> 3'
 STACKING_ENERGY.update({('A', 'A', 'U', 'U'):-0.9,
@@ -413,7 +431,7 @@ def forward(profile,ref_seq,struct,(i,j),(a,b),alpha):
   else:
     k = struct[i]
     if k==-1:
-      for a2 in BASES:
+      for a2 in BASES[i]:
         pro = profile[i][a2]
         f = forward(profile,ref_seq,struct,
                     (i+1,j),
@@ -421,8 +439,8 @@ def forward(profile,ref_seq,struct,(i,j),(a,b),alpha):
                     alpha)
         result +=  pro*f
     elif i < k <= j: #If k > j we return 0
-      for a2 in BASES:
-        for b2 in BASES:
+      for a2 in BASES[i]:
+        for b2 in BASES[k]:
           pro = profile[i][a2]*profile[k][b2]
           #if not stacked outside (or border, then no stack possible)
           if i==0 or j==len(struct)-1 or not (j==k and struct[i-1]==j+1):
@@ -466,7 +484,7 @@ def backward(profile,ref_seq,struct,(i,j),(a,b),alpha):
   else:
     k = struct[i]
     if k==-1:
-      for a2 in BASES:
+      for a2 in BASES[i]:
         pro = profile[i][a2]
         back = backward(profile,ref_seq,struct,
[…]

### 2013-03-03 ISMB/introduction_ISMB.tex
typos

@@
 Finally, \ourprog implements a novel algorithm using weighted sampling techniques~\cite{Bodini2010} that enables us to control, for the first time, \textit{explicitly} the \GCContent of the solution.
-This functionality is essential because wild sequences within living organisms often present medium or low \GCContent, presumably to offer better transcription rates and/or structural plasticity.
+This functionality is essential because wild-type sequences within living organisms often present medium or low \GCContent, presumably to offer better transcription rates and/or structural plasticity.
 Previous programs do not allow to control this parameter and tend to output sequences with high \GCContent{}s.

### 2013-03-03 ISMB/main_ISMB.tex
typos

@@
 \textbf{Motivations:} The design of RNA sequences folding into predefined secondary structures is a milestone for many synthetic biology and gene therapy studies.
-Most of the current software use similar local search strategies (i.e.
+Most of the current software uses similar local search strategies (i.e.
 a random seed is progressively adapted to acquire the desired folding propertues) and more importantly do not allow the user to control explicitly the nucleotide distribution such as the \GCContent in their sequences.
-However, the latter is an important criteria for large-scale applications as it could presumably be used to design sequences with better transcription rates and/or structural plasticity.\\
+However, the latter is an important criterion for large-scale applications as it could presumably be used to design sequences with better transcription rates and/or structural plasticity.\\
 \noindent

### 2013-03-03 ISMB/methods_ISMB.tex
typos

@@
 stacked canonical base-pairs.
-We show how a modification of the dynamic programming scheme used in the \RNAmutants allows for the sampling of good and diverse design candidates, in linear time and space complexities.
+We show how a modification of the dynamic programming scheme used in \RNAmutants allows for the sampling of good and diverse design candidates, in linear time and space complexities.
 
@@
 
-Despite its apparent quadratic complexity, the recurrence described in Subsection~\ref{sec:pf} can be computed in linear $\Theta(|\Target|)$ time and space, owing to the fact that, on any recursive call, the position $j$ is entirely determined by $i$.
+Despite its apparent quadratic complexity, the recurrence described in Subsection~\ref{sec:pf} can be computed in linear $\Theta(|\Target|)$ time and space, owing to the fact that, on any recursive call, the position $k$ is entirely determined by $i$.
 
@@
 \label{subsec:glocal_method}
-Due to our simplified our energy model, unpaired regions are not subject to design constraints other than the \GCContent, leading to modest probabilities for refolded design candidates to match the targeted structure.
+Due to our simplified energy model, unpaired regions are not subject to design constraints other than the \GCContent, leading to modest probabilities for refolded design candidates to match the targeted structure.
 To improve these performances and test the complementarity  of our global sampling approach with previous contributions based on local search, we used the \RNAinverse software to redesign unpaired regions.
-We specified a constraint mask to prevent stacking base-pairs from being modified and, whenever necessary, reetablished their content content a posteriori as \RNAinverse has been witnessed to take some liberties with constraints masks.
+We specified a constraint mask to prevent stacking base-pairs from being modified and, whenever necessary, reestablished their content a posteriori as \RNAinverse has been witnessed to take some liberties with constraints masks.
 As shown in Table~\ref{table:impact_on_gc} (Supplementary material), this postprocessing does not drastically alter the \GCContent, so the glocal approach reasonably addresses the contrained \GCContent design problem.

### 2013-03-03 ISMB/results_ISMB.tex
fix typo
typos

@@
  To ease the visualization of the results, we clustered together strucures
- having similar length, stacks density and proportion of free-nucleotides in loops. 
+ having similar length, stacks density and proportion of free nucleotides in loops. 
   We present in Figure~\ref{fig:bins} the distribution of structures in each category.
@@
 \subsection{Design}
- To benchmark our method, we first started by sampling about $100$ sequences per structure. After correcting the free nucleotides with
+ To benchmark our method, we first started by sampling about $1000$ sequences per structure. After correcting the free nucleotides with
  \texttt{RNAinverse} we computed their MFE with \texttt{RNAfold} from the \textit{Vienna Package 2.0}~\cite{Hofacker:1994}.
@@
 Even with a \GCContent of $50\%$ and more, when at least
-one solutions was found for almost all structures, most of the samples 
+one solution was found for almost all structures, most of the samples 
 do not fold into the target. A few structures remained unsolved under
@@
 
-Also, a critical properties in RNA sequence design is 
+Also, a critical propertie in RNA sequence design is 
 the frequency of the MFE. 

### 2013-03-03 ISMB/supplementary_ISMB.tex
typos

@@
 a high entropy.
-Table.~\ref{tab:nb_rnainv} contains the number of sample with a given \GCContent generated by \RNAinverse. The vast majority resides around $50\%-70\%$, showing one of its major deficiencies. 
+Table.~\ref{tab:nb_rnainv} contains the number of samples with a given \GCContent generated by \RNAinverse. The vast majority resides around $50\%-70\%$, showing one of its major deficiencies. 
  For all structures that have been solved 
@@
 we present for every different concentration of \GCContent
-the average sample entropy. and the average base pairs entropy in Figure.~\ref{fig:rnainverse}.
+the average sample entropy and the average base pairs entropy in Figure.~\ref{fig:rnainverse}.
 

### 2013-03-08 ISMB/methods_ISMB.tex
Typos + Changed sampling figure

@@
 
-A targeted secondary structure $\Target$ is abstracted as a non-crossing arc-annotated sequence,  where 
+A targeted secondary structure $\Target$ is given as a non-crossing arc-annotated sequence,  where 
 $\Target_i$ stands for the base-pairing position of position $i$ in $\Target$ if any (and, reciprocally, $\Target_{\Target_i}=i$), or $-1$ otherwise. 
-Let us denote by $\gc(s)$ the number of occurrences of \Gb and \Cb in an RNA sequence $s$.
+In addition, we denote by $\gc(s)$ the number of occurrences of \Gb and \Cb in an RNA sequence $s$.
 
 \subsubsection{Simplified energy model}
-The free-energy contribution in our pseudo-energy model corresponds to an additive stacking-pairs model, using individual values from the Turner 2004 model (retrieved from the NNDB~\cite{Turner2010}).
+We use a simplified free-energy model which only includes additive contributions from stacking base-pairs.
+Using individual values from the Turner 2004 model (retrieved from the NNDB~\cite{Turner2010}).
 Given a candidate sequence $s$ for a secondary structure $S$, the free-energy of $S$ on $s$ is given by
@@
 not in $\{\Gb\Ub,\Ub\Gb,\Cb\Gb,\Gb\Cb, \Ab\Ub\text{ or }\Ub\Ab\}$).
-This latter parameter allows one to choose whether to simply penalize invalid base pairs ($\beta>0$), or forbid them altogether ($\beta = \infty$).
+This latter parameter allows one to choose whether to simply penalize invalid base pairs ($\beta>0$), or forbid them altogether ($\beta = +\infty$).
 
@@
 
-In order to counterbalance the documented tendency of sampling methods to generate \Gb\Cb-rich sequences~\cite{Levin:2012kx}, let us introduce a parameter $x\in\mathbb{R}^+$, whose value will influence the \GCContent of generated sequences.
+In order to counterbalance the documented tendency of sampling methods to generate \Gb\Cb-rich sequences~\cite{Levin:2012kx}, we introduce a parameter $x\in\mathbb{R}^+$, whose value will influence the \GCContent of generated sequences.
 Let $\Target$ be a targeted secondary structure, then let us define the pseudo-Boltzmann factor $\B_{x}(s,S)$ of a sequence $s$ such that
@@
 
-Summing the pseudo-Boltzmann factor over all possible sequences of a given length $|\Target|$ gives the pseudo-partition function $\mathcal{Z}_{x,\Target}$, from which one defines the pseudo-Boltzmann probability $\Prob_{x,\Target}(s)$ of each sequence $s$, respectively such that
+Summing the pseudo-Boltzmann factor over all possible sequences of a given length $|\Target|$, one obtains the pseudo-partition function $\mathcal{Z}_{x,\Target}$, from which one defines the pseudo-Boltzmann probability $\Prob_{x,\Target}(s)$ of each sequence $s$, respectively such that
 \begin{align}\mathcal{Z}_{x,\Target} &= \sum_{\substack{|s|=|\Target|}}\B_{x,\Target}(s)& \text{and}&&
@@
 Firstly, a dynamic programming algorithm computes $\mathcal{Z}\substack{(i,j)\\ [a,b]}$ the pseudo-partition function restricted to an interval $[i,j]$, assuming its (previously chosen) flanking nucleotides are $a$ and $b$ respectively.
-Since the only sequence over an empty interval is the empty sequence, having energy $0$, one has
+Remark that the empty interval only supports the empty structure, having energy $0$, so one has
 \begin{equation}
@@
       x^{\gc(a')}
-      \cdot\Z{i+1,j}{a',b} &\begin{array}{@{}l@{}}\text{If }\Target_{i}=-1\\ \text{\relsize{-1}Position $i$ unpaired,}\end{array}\\
+      \cdot\Z{i+1,j}{a',b} &\begin{array}{@{}l@{}}\text{If }\Target_{i}=-1,\\ \text{\relsize{-1}[Position $i$ unpaired]}\end{array}\\
       \displaystyle
@@
 			 \cdot \Z{i+1,j-1}{a',b'}&
-\begin{array}{@{}l@{}}\text{Elif }\Target_i=j \land \Target_{i-1}=j+1\\ \text{\relsize{-1}$(i,j)$ stacking pair onto $(i$-$1,j$+$1)$,}\end{array}
+\begin{array}{@{}l@{}}\text{Elif }\Target_i=j \land \Target_{i-1}=j+1,\\ \text{\relsize{-1}[$(i,j)$ stacking pair onto $(i$-$1,j$+$1)$]}\end{array}
 			 \\
@@
       \cdot\Z{k+1,j}{b',b} & 
-\begin{array}{@{}l@{}}\text{Elif }\Target_i=k \land i < k \leq j\\
-\text{\relsize{-1}Junction/bulge $\Rightarrow$ split, no stack,}
+\begin{array}{@{}l@{}}\text{Elif }\Target_i=k \land i < k \leq j,\\
+\text{\relsize{-1}[Junction/bulge $\Rightarrow$ split, no stack]}
 \end{array}
@@
 
-\subsubsection{Stochastic backtrack}
+\subsubsection{\GCContent Stochastic backtrack}
 Once the pseudo-partition functions have been memorized, a stochastic backtrack starts from the complete interval $[0,|\Target|-1]$ and, at each step, chooses a suitable assignment for one or several positions, using probabilities derived from the precomputation.
@@
 To improve these performances and test the complementarity  of our global sampling approach with previous contributions based on local search, we used the \RNAinverse software to redesign unpaired regions.
-We specified a constraint mask to prevent stacking base-pairs from being modified and, whenever necessary, reestablished their content a posteriori as \RNAinverse has been witnessed to take some liberties with constraints masks.
+We specified a constraint mask to prevent stacking base-pairs from being modified and, whenever necessary, reestablished their content {\em a posteriori}, as \RNAinverse has been witnessed to take some liberties with constraints masks.
 As shown in Table~\ref{table:impact_on_gc} (Supplementary material), this postprocessing does not drastically alter the \GCContent, so the glocal approach reasonably addresses the contrained \GCContent design problem.

### 2013-03-08 ISMB/results_ISMB.tex
Typos + Changed sampling figure

@@
 \RNAinverse from the \textit{Vienna Package 2.0}~\cite{Hofacker:1994}.
-All time benchmarks were run on a one AMD Opteron(tm) Processor 6278  at 2.4 GHz with cache of 512 KB.
- We used as parameter $\beta=15$, the penalty for an invalid base pair.
+All time benchmarks were run on a single AMD Opteron(tm) 6278 Processor  at 2.4 GHz with cache of 512 KB.
+The penalty $\beta$, associated with invalid base-pairs, was set to $15$.
 
-We present the average times spent by \ourprog and \RNAinverse to generate one sequence
-with the required \GCContent in Figure~\ref{fig:time}. As expected, the time grows linearly
-in function of the length of the structures for \ourprog.  A highly interesting feature
-is the huge decrease in time of computation when there is $15\%$ free 
-nucleotides inside the loops of the secondary structure.
-
-\begin{figure}[ht!]
+\begin{figure}[t!]
 	\centering
@@
 
+Figure~\ref{fig:time} presents the average times spent running \ourprog and \RNAinverse to generate one sequence
+with the required \GCContent. As expected, the time grows linearly
+in function of the length of the structures for \ourprog.  A highly intriguing feature is the substantial decrease of the time 
+required for running \RNAinverse when $15\%$ of nucleotides are unpaired in the target structure.
+However, one must remain careful interpreting this observation, as the structures within this class all originate from the PDB, and are relatively small (for the complete STRAND DB, the average length is $\sim526$nts, compared to $\sim38$nts around 15\% unpaired bases).
+
 \subsection{Dataset}
@@
 Those are known secondary structure from a variety of organisms.
-We select the same subset of $50$ structures as in~\cite{Levin:2012kx}, 
+We considered a subset of $50$ structures selected by Levin~\emph{et al}~\cite{Levin:2012kx}, 
 whose length ranges between $20$ and $100$ nucleotides. 
-
- To ease the visualization of the results, we clustered together strucures
- having similar length, stacks density and proportion of free nucleotides in loops. 
-  We present in Figure~\ref{fig:bins} the distribution of structures in each category.
+ To ease the visualization of results, we clustered together structures
+ having similar length, stacks density and proportion of free nucleotides in loops, leading to distributions of structures shown in Figure~\ref{fig:bins}.
 
@@
 
-{\em Ajouter considérations Base Pair entropy: Se rappeler, que moins, c'est bien !}
+{\em Ajouter considérations Base Pair entropy: Se rappeler, que moins, c'est mieux !}
 

### 2013-03-08 ISMB/supplementary_ISMB.tex
Typos + Changed sampling figure

@@
 \hline
-target \GCContent & \ourprog & \ourprog + \RNAinverse \\
+\multirow{3}{*}{Target \GCContent (\%)}& \multicolumn{2}{|c|}{Designed sequences \GCContent (\%)}\\ \cline{2-3}
+ & \ourprog & \ourprog + \RNAinverse\\
+ & (Global) & (Glo{\bf c}al) \\
 \hline
-10\% & 0.15 & 0.21\\
-30\% & 0.30 & 0.33\\
-50\% & 0.48 & 0.49\\
-70\% & 0.71 & 0.69\\
-90\% & 0.83 & 0.78\\
+10\% & 15\% & 21\% \quad $\nearrow 6\%$\\
+30\% & 30\% & 33\% \quad $\nearrow 3\%$\\
+50\% & 48\% & 49\% \quad $\nearrow 1\%$\\
+70\% & 71\% & 69\% \quad $\searrow 2\%$\\
+90\% & 83\% & 78\% \quad $\searrow 5\%$\\
 \hline
@@
 \end{center}
-\caption{Observed \GCContent of solutions returned by \ourprog (2nd column) and after the application of the local search heuristic (3rd column).}
+\caption{Observed \GCContent of solutions returned by \ourprog (2nd column) and after the application of the local search postprocessing (3rd column).}
 \label{table:impact_on_gc}

### 2013-03-08 ISMB/ResponseToReviewers-Incarnation-ISMB-2013.tex
Added draft of response to reviewer

% !TEX TS-program = pdflatex
% !TEX encoding = UTF-8 Unicode

% This is a simple template for a LaTeX document using the "article" class.
% See "book", "report", "letter" for other types of document.

\documentclass[11pt,hyperref,draft]{article} % use larger type; default would be 10pt

\usepackage[utf8]{inputenc} % set input encoding (not needed with XeLaTeX)

%%% Examples of Article customizations
% These packages are optional, depending whether you want the features they provide.
% See the LaTeX Companion or other references for full information.

%%% PAGE DIMENSIONS
\usepackage{geometry} % to change the page dimensions
\geometry{a4paper} % or letterpaper (US) or a5paper or....
% \geometry{margin=2in} % for example, change the margins to 2 inches all round
% \geometry{landscape} % set up the page for landscape
%   read geometry.pdf for detailed page layout information

\usepackage{graphicx} % support the \includegraphics command and options

% \usepackage[parfill]{parskip} % Activate to begin paragraphs with an empty line rather than an indent

%%% PACKAGES
\usepackage{amsmath} % for much better looking tables
\usepackage{xspace}
\usepackage{amssymb} % for much better looking tables
\usepackage{booktabs} % for much better looking tables
\usepackage{array} % for better arrays (eg matrices) in maths
\usepackage{ntheorem} % for better arrays (eg matrices) in maths
\usepackage{paralist} % very flexible & customisable lists (eg. enumerate/itemize, etc.)
\usepackage{verbatim} % adds environment for commenting out blocks of text & for better verbatim
\usepackage{subfig} % make it possible to include more than one captioned figure/table in a single float
\usepackage{framed}
% These packages are all incorporated in the memoir class to one degree or another...

%%% HEADERS & FOOTERS
\usepackage{fancyhdr} % This should be set AFTER setting up the page geometry
\pagestyle{fancy} % options: empty , plain , fancy
\renewcommand{\headrulewidth}{0pt} % customise the layout...
\lhead{}\chead{}\rhead{}
\lfoot{}\cfoot{\thepage}\rfoot{}

%%% SECTION TITLE APPEARANCE
\usepackage{sectsty}
\allsectionsfont{\sffamily\mdseries\upshape} % (See the fntguide.pdf for font help)
% (This matches ConTeXt defaults)

%%% ToC (table of contents) APPEARANCE
\usepackage[nottoc,notlof,notlot]{tocbibind} % Put the bibliography in the ToC
\usepackage[titles,subfigure]{tocloft} % Alter the style of the Table of Contents
\renewcommand{\cftsecfont}{\rmfamily\mdseries\upshape}
\renewcommand{\cftsecpagefont}{\rmfamily\mdseries\upshape} % No bold!

\newcommand{\Answer}[1]{\noindent\textsf{\textbf{Response: }}{\sf#1}\\}
\newcommand{\Comment}[1]{\noindent\textsf{\textbf{Comment: }}{\it#1}\\[.5em]}

%%% END Article customizations

\input{macros}

%%% The "real" document content comes below...

\title{A weighted sampling algorithm for the design of RNA sequences with targeted secondary structure and nucleotides distribution}
\author{Vladimir Reinharz$^1$, Yann Ponty$^{2,*}$, J\'er\^{o}me Waldisp\"{u}hl$^{1,*}$}
\date{$^1$ School of Computer Science, McGill University, Montreal, Canada, $^2$ Laboratoire d'informatique, \'Ecole Polytechnique, Palaiseau, France.\\ \small $^*$Corresponding authors: \texttt{jeromew@cs.mcgill.ca}, \texttt{yann.ponty@lix.polytechnique.fr}}

%\date{} % Activate to display a given date or no date (if empty),
         % otherwise the current date is printed

\begin{document}
\maketitle

First of all, the authors wish to express their gratitude to the reviewers for their encouraging comments and constructive criticisms, which we tried to address in this revised version of our manuscript.

\section{Detailed responses}

\subsection{Reviewer \#1}

\Comment{The benchmark only compares to RNAinverse which is outperformed by most methods you name in your introduction. Thus, a real benchmarking would have to compare to other methods as well. For most, source code packages are available.}
\Answer{We systematically benchmarked {\tt IncaRNAtion} against most of the existing available tools, including NUPack~\cite{Zadeh:2011fk}, FRNAkenstein~\cite{Lyngso:2012vn},  RNA-SSD~\cite{Aguirre-Hernandez:2007kx}, Info-RNA~\cite{Busch:2006uq} and RNAexInv~\cite{Avihoo:2011fk}.
While the page limitations associated with a conference submission do not allow for a an exhaustive discussion, we now report in the revised manuscript some of the most striking results. TODO (Comment results)
}

\Comment{Since you have shown that RNAinverse biases the GC-content to 50-70, the GC-content increase/decrease for lower/higher IncaRNAtion values, resp., when successivly applying RNAinverse (Supp 5.2) is to be expected.}
\Answer{Absolutely. However, the point of this section was not to put the stress on this issue, but rather to show that our glocal approach, starting with a global sampling followed by a postprocessing of unpaired regions using \RNAinverse, does not {\em drift} too much towards regions of high \GCContent{}s. We rephrased the text of this section to make this conclusion appear more clearly, and changed the layout of Table 2 to make the shift in \GC more explicit.}

\Comment{Is there a way to apply sequence constraints within IncaRNAtion? This is something most end user will want.
}
\Answer{We added this feature to our implementation, which now accepts a constraint mask consisting in a sequence of IUPAC symbols (e.g. 'N' for unconstrained position, or G for Guanine-only). We also briefly mention this feature in the definition of our probabilistic model.}

\Comment{What is the time consumption/complexity of the GC-weight parameter $x_{gc}$ identification within Sec. 2.2.3? And are these values length-/structure-independent or do they have to be computed for each structure individually?}
\Answer{TODO (Yann)}

\Comment{Is the method available?}
\Answer{TODO (Jerome)}

\subsection{Reviewer \#2}

\Comment{1) The $x^{\gc(s)}$ in equation (1) seems to me a novel contributing term that is of high importance. Perhaps its introduction should be highlighted more in the text.}
\Answer{TODO (Yann)}

\Comment{2) Concerning Table 2 in the Supplementary data, the way it is written (IncaRNAtion + RNAinverse) it looks like "glocal" search heuristic whereas in the table it seems local. Perhaps there is a small typo ambiguity that can be resolved.}
\Answer{Reviewer \#1 also pointed out some inconsistency/unclarity in our previous formulation. The second column indeed showed the final \GCContent of the glocal approach, following the postprocessing of sequences with a local-search approach. While such a postprocessing tends to reestablish \GCContent{}s that are  closer to 50\% \GC, this shift remains reasonable (at most ~6\% \GC). Furthermore, this shift is very predictable and offers the user the possibility to \emph{overshoot} for a more extreme \GCContent with the global approach, knowing that the postprocessing step will tend to average it.}

\Comment{3-4) Typos\ldots}
\Answer{We fixed those typos.}

\subsection{Reviewer \#3}

\Comment{(1) There are many points where you iterate over "all sequences of length $|S|$" or all of those which can fold into $S$; it is not always clear. Why not introduce $F(s)$ for the folding space of a sequence, and $F^{-1}(S)$ for the inverse folding space of a structure, and simply use $F^{-1}(S)$ where this is what you mean?}
\Answer{While we agree that this proposition would lead to a compact and well-defined mathematical definition of the problem, we prefer a implicit definition of the compatibility of a sequence to a given structure. Namely, we always allow every possible sequence of a length $|S|$ to be a member of the inverse folding space, and assign to it some  energy penalty (typically $+\infty$, or $\beta$) if it violates some constraints (restriction on allowed base-pairs, position-specific constraints...). The inverse folding space sequences is therefore implicitly defined as the set of sequences having finite energy, and our algorithm is arguably more prone to generalization, e.g. to capture non-canonical base-pairs which are increasingly considered by the field of RNA bioinformatics.
}

\Comment{(2) You make the computation of $Z$ look like yet another set of DP recurrences, and as a consequence, you need a proof to show that it is linear in $n$. Surely, the reader wonders at that point how you can iterate over $(i,j)$ and compute an $(n,n)$ table in less than quadratic time. This mis-led understanding is created by the Nussinovic style of the recurrences.
 What you actually compute is a sparse, tree-shaped segment of that $(i,j)$-table.
 Why not represent the given structure as a tree of Empty, Singleton, Pair and Bifurcation nodes, such as 
\begin{align*}
  S<\texttt{.},B<P<\texttt{(},S<\texttt{.},E>,\texttt{)},P<\texttt{(},Em,\texttt{)}>>> \text{ for {\tt .(.)()}}  ? 
\end{align*}
 (You do not even need to explicitly represent the dots and brackets at the leaves.)\\
 Then, computation of $Z$ becomes just structural recursion on the tree,
     $Z(a,b, t) =$ defined by case analysis on t, such that, for example,
     $Z(a,b,P<'(',r,')'>)$ calls on $Z(a',b' r)$ for all $a'$ and $b'$.
And the result of the computation is just the tree annotated with the $Z_{ab}$ values. 
This way it is more transparent what is going on, and it is obvious that space and time are linear in $n$, because the size of the tree is in $O(n)$.}
\Answer{Great suggestion. TODO!}

\Comment{(3) Figure 1 is also a bit misleading. Make it clear that the stochastic choice is NOT made over the three cases depicted. The case is always fixed for given $i$, just the bases are chosen.}
\Answer{Absolutely. We changed this figure, which now explicitly distinguishes between the three different cases (different left-hand side diagrams), and also features the base emission probabilities.}

\Comment{(4) Section 2.2.3 is titled "Rejecting unsuitable candidates". It explains that candidates tend to be quite suitable, but somewhere, you should actually tell who is rejected and when ... }
\Answer{TODO}

\Comment{(5) recurring: "free-nucleotides in loops" -- isn't that the same, being "free" and in a loop? And: drop the hyphen.}
\Answer{TODO}

\bibliographystyle{amsplain}
\bibliography{RNApyro}
\end{document}

### 2013-03-08 ISMB/methods_ISMB.tex
Added draft of response to reviewer

@@
 This latter parameter allows one to choose whether to simply penalize invalid base pairs ($\beta>0$), or forbid them altogether ($\beta = +\infty$).
+Position-specific sequence constraints can also be enforced at this level (details omitted for the sake of clarity) by assigning to $\ES$ a $+\infty$ penalty (leading to a null probability) to the presence, at a given position, of a base that is incompatible with a user-specified constraint mask.
 
@@
 In order to counterbalance the documented tendency of sampling methods to generate \Gb\Cb-rich sequences~\cite{Levin:2012kx}, we introduce a parameter $x\in\mathbb{R}^+$, whose value will influence the \GCContent of generated sequences.
-Let $\Target$ be a targeted secondary structure, then let us define the pseudo-Boltzmann factor $\B_{x}(s,S)$ of a sequence $s$ such that
+Let $\Target$ be a targeted secondary structure, then the pseudo-Boltzmann factor of a sequence $s$ is $\B_{x}(s,S)$  such that
 \begin{equation}
@@
 \subsubsection{Precomputing the pseudo-partition function}\label{sec:pf}
-\begin{figure}
-\resizebox{\textwidth}{!}{\input{FigStochasticBacktrack.pgf}}
-\caption{Three main cases in the stochastic backtrack procedure over an interval $[i,j]$: Either position $i$ is left unpaired (top), a base-pair $(i,j)$ is formed, stacking onto $(i+1,j-1)$ (middle), or $i$ is paired somewhere else, defining two regions on which subsequent recursive calls are needed (bottom).
-Positions indicated in red are assigned during this stage of the backtrack.\label{fig:stochastic}}
-\end{figure}
 
@@
 
-\subsubsection{\GCContent Stochastic backtrack}
-Once the pseudo-partition functions have been memorized, a stochastic backtrack starts from the complete interval $[0,|\Target|-1]$ and, at each step, chooses a suitable assignment for one or several positions, using probabilities derived from the precomputation.
-One or several recursive calls over regions that are left unassigned are then performed, as described by Algorithm~\ref{alg:back}, and illustrated by Figure~\ref{fig:stochastic}.
+\begin{figure}
+\resizebox{\textwidth}{!}{\input{FigStochasticBacktrack.pgf}}
+\caption{Stochastic backtrack procedure over an interval $[i,j]$: Either position $i$ is left unpaired (top), a base-pair $(i,j)$ is formed, stacking onto $(i+1,j-1)$ (middle), or $i$ is paired somewhere else, defining two regions on which subsequent recursive calls are needed (bottom).
+Positions indicated in red are assigned during this stage of the backtrack.\label{fig:stochastic}}
+\end{figure}
+\subsubsection{Stochastic backtrack}
+Once the pseudo-partition functions have been memorized, a stochastic backtrack starts from the complete interval $[0,|\Target|-1]$ and, at each step, chooses a suitable assignment for one or several positions, using probabilities derived from the precomputation, as illustrated by Figure~\ref{fig:stochastic}.
+One or several recursive calls over the remaining regions are then performed, as described in Algorithm~\ref{alg:back}.
+On each recursive call, the algorithm assigns at least one nucleotide to a -- previously unassigned -- position.
+Moreover, the number of executions of each loops is bounded by a constant.
+Consequently, the complexity of Algorithm~\ref{alg:back} is in $\Theta(|\Target|)$ time and space.
+
 \begin{algorithm}[t]
@@
   k $\leftarrow S_i$\;
-	\If(\tcp*[f]{Unpaired}){$(k = -1)$}{
+	\If(\tcp*[f]{Unpaired base}){$(k = -1)$}{
 		\For{$a'\in\B$}{
 			rand $\leftarrow$ rand $- x^{\gc(a')}\cdot \Z{i+1,j}{a',b}$\;
-			\lIf{$\text{rand}<0$}\Return{$a'.\Backtrack\left(i+1,j,a',b,\Target\right)$}\;
+			\lIf{$\text{\rm rand}<0$}\Return{$a'.\Backtrack\left(i+1,j,a',b,\Target\right)$}\;
 		}
@@
 	\lElse{
-		\If(\tcp*[f]{Stacking pairs}){$(k=j) \land (S_{i-1}=j+1)$}{
+		\If(\tcp*[f]{Stacking base pair}){$(k=j) \land (S_{i-1}=j+1)$}{
 			\For{$(a',b')\in\B\times\B$}{
@@
 			 	\cdot \Z{i+1,j-1}{a',b'}	$\;
-				\lIf{$\text{rand}<0$}{
+				\lIf{$\text{\rm rand}<0$}{
 					\Return{$a'.\Backtrack\left(i+1,j-1,a',b',\Target\right). b'$}\;		
@@
 	    	  \cdot\Z{k+1,j}{b',b}$\;	
- 					\If{$\text{rand}<0$}{
+ 					\If{{\rm rand} $<0$}{
 						\Return{$a'
@@
 
-Since, on each recursive call, the algorithm assigns at least one nucleotide to a -- previously unassigned -- position, and that the number of executions of the loops are bounded by a constant, then the complexity of the algorithm grows linearly on the targeted length.
+\subsubsection{Self-adaptive sampling strategy and overall complexity}
 
-\subsubsection{Rejecting unsuitable candidates}
+Our goal is to produce a set of sequences whose \GCContent matches a prescribed value $gc$.
+An absolute tolerance $\kappa$ may be allowed, so that the \GCContent of any valid sequence must fall in  
+$[gc-\kappa,gc+\kappa]$.
+Since sequences of arbitrary \GCContent may be generated by Algorithm~\ref{alg:back}, we use a rejection-based approach~\cite{Bodini2010}, previously adapted by the authors in a similar context~\cite{Waldispuhl2011}.
+In our context, this gives an algorithm which generates $k$ valid sequences in expected time $\Theta(k\cdot n\sqrt{n})$ when $\kappa=0$ (resp.
+$\Theta(k\cdot n)$ when $\kappa$ is a positive constant) memory $\Theta(k\cdot n)$.
+Although a complete analysis can be found in an earlier contribution~\cite{Waldispuhl2011}, let us briefly outline the strategy and arguments used to analyze the complexity.
 
-Note that, although the parameter $x$ is introduced to match the expected \GCContent with the targeted one, nothing prevents Algorithm~\ref{alg:back} to generate sequences of arbitrary \GCContent.
-Therefore, we use a rejection-based approach~\cite{Bodini2010} previously used by the authors in a similar context~\cite{Waldispuhl2011}.
+This approach simply generates sets of sequences by repeatedly running the stochastic backtrack algorithm.
+The average \GCContent induced by the current value of the $x$ parameter, can then be adequately estimated from the sample, or computed exactly using recent algorithmic advances~\cite{Ponty2011}.
+The set of sequences is filtered to only retain valid sequences.
+The value of the parameter $x$ is then adapted to match the average \GCContent (induced by the value of $x$) with the targeted one.
+It can be shown that the expected \GCContent is a continuous and strictly increasing monotonic function of $x$, whose limits are $0$ when $x=0$ and $n$ when $x\to +\infty$.
+Consequently, for any targeted \GCContent $gc\in[0\%,100\%]$, there exists a unique value $x_{gc}$ such that generated sequences feature, on the average, the right \GCContent.
+In practice, a simple binary search~\cite{Waldispuhl2011} is used in our implementation, and typically converges after very few iterations.
+However the optimal value for $x$ can also be derived analytically using an interpolation after $\Theta(n)$ evaluations of $\Z{i,j}{a,b}$ for different candidate values of $x$, as previously noted~\cite{Waldispuhl2011} and implemented using the Fast-Fourier Transform~\cite{Senter2012}.
 
-It was established~\cite{Waldispuhl2011} that, for each value of $x$, there exists constants $\mu_x$ and $\sigma_x$ such that the \GCContent distribution observed in sequences sampled with respect to the above distribution converges asymptotically towards a normal law having expectation in $\mu\cdot n\cdot(1+o(1))$ and standard deviation in $\sigma\cdot\sqrt{n}\cdot(1+o(1))$.
-Furthermore, it can easily be shown that the expected \GCContent is a continuous and strictly increasing monotonic  function of $x$, whose limits are $0$ when $x=0$ and $n$ when $x\to +\infty$.
-
-Consequently, for any targeted \GCContent $gc\in[0\%,100\%]$, there exists a unique value $x_{gc}$ such that generated sequences feature, on the average, the right \GCContent.
-Such a value can be determined using any numerical recipe, and a simple bisection procedure, further described in earlier work~\cite{Waldispuhl2011}, is used in our implementation.
+Regarding the complexity, it was previously established~\cite{Waldispuhl2011} that, for each value of $x$, there exists constants $\mu_x$ and $\sigma_x$ such that the distribution of \GCContent asymptotically converges towards a normal law having expectation in $\mu\cdot n\cdot(1+o(1))$ and standard deviation in $\sigma\cdot\sqrt{n}\cdot(1+o(1))$.
 Furthermore, the concentration of the distribution, asserted by its limited standard deviation, implies a growth in $\Theta(\sqrt{n})$ (resp.
 $\Theta(1)$, i.e.
-constant) for the expected number of attempts before generating a sequence having the exact targeted \GCContent  (resp.
-within $[gc-\kappa,gc+\kappa]$, where $\kappa\in\Theta(1/\sqrt n)$).
+constant) for the expected number of attempts required to generate a valid sequence when $\kappa=0$ (resp.
+$\kappa\in\Omega(1/\sqrt n)$).
 

### 2013-03-08 ISMB/supplementary_ISMB.tex
Added draft of response to reviewer

@@
 %\end{figure}
-\subsection{Impact of local search on the \GCContent of \ourprog output}
-
-These data demonstrate that the local search heuristic used to design nucleotides in loop regions has a very limited impact on the \GCContent.
-For each class of \GCContent, we show in Table~\ref{table:impact_on_gc}, the observed \GCContent of the sequence sampled by \ourprog, and then we report the observed \GCContent after the \RNAinverse step (as defined in Section \ref{subsec:glocal_method}).
-Our results show that the \GCContent is relatively well conserved.
+\subsection{Limited impact on \GC of local-search postprocessing of \ourprog output}
+Since local search approaches tend to experience a bias towards \GC{}-rich regions, it could be expected that our glocal approach, by postprocessing unpaired regions using a local search algorithm, would suffer from such a drift.
+However, as summarized in Table~\ref{table:impact_on_gc}, we observed that the local search heuristic used to design nucleotides in loop regions has a very limited impact on the \GCContent.
+For each class of \GCContent, we reported the observed \GCContent in the sequence initially generated by \ourprog, and the observed \GCContent after the \RNAinverse postprocessing (as defined in Section \ref{subsec:glocal_method}).
+Our results show that the \GCContent is relatively well conserved (less than 6\% variation), with a general tendency of the postprocessing step to bring the \GCContent back to 50\%.
 
@@
  & \ourprog & \ourprog + \RNAinverse\\
- & (Global) & (Glo{\bf c}al) \\
+ & (Global) & ({Glocal}) \\
 \hline

### 2013-03-08 ISMB/methods_ISMB.tex
Update

@@
 
-A targeted secondary structure $\Target$ is given as a non-crossing arc-annotated sequence,  where 
+A targeted secondary structure $\Target$ of length $n$ is given as a non-crossing arc-annotated sequence,  where 
 $\Target_i$ stands for the base-pairing position of position $i$ in $\Target$ if any (and, reciprocally, $\Target_{\Target_i}=i$), or $-1$ otherwise. 
-In addition, we denote by $\gc(s)$ the number of occurrences of \Gb and \Cb in an RNA sequence $s$.
+In addition, let us denote by $\gc(s)$ the number of occurrences of \Gb and \Cb in an RNA sequence $s$.
 
@@
 Using individual values from the Turner 2004 model (retrieved from the NNDB~\cite{Turner2010}).
-Given a candidate sequence $s$ for a secondary structure $S$, the free-energy of $S$ on $s$ is given by
+Given a candidate sequence $s$ for a secondary structure $\Struct$, the free-energy of any sequence $s$ of length $|\Struct|$  is given by
 \begin{align*}
-  \ES(s,S) = \sum_{\substack{(i,j)\to (i',j')\in S\\ \text{stacking pairs}}}\ES^{\beta}_{s_is_j\to s_{i'}s_{j'}} 
+  \ES(s,\Struct) = \sum_{\substack{(i,j)\to (i',j')\in \Struct\\ \text{stacking pairs}}}\ES^{\beta}_{s_is_j\to s_{i'}s_{j'}} 
 \end{align*}
@@
 This latter parameter allows one to choose whether to simply penalize invalid base pairs ($\beta>0$), or forbid them altogether ($\beta = +\infty$).
-Position-specific sequence constraints can also be enforced at this level (details omitted for the sake of clarity) by assigning to $\ES$ a $+\infty$ penalty (leading to a null probability) to the presence, at a given position, of a base that is incompatible with a user-specified constraint mask.
+Position-specific sequence constraints can also be enforced at this level (details omitted for the sake of clarity) by assigning to $\ES$ a $+\infty$ penalty (leading to a null probability) in the presence of a base incompatible with a user-specified constraint mask.
 
@@
 In order to counterbalance the documented tendency of sampling methods to generate \Gb\Cb-rich sequences~\cite{Levin:2012kx}, we introduce a parameter $x\in\mathbb{R}^+$, whose value will influence the \GCContent of generated sequences.
-Let $\Target$ be a targeted secondary structure, then the pseudo-Boltzmann factor of a sequence $s$ is $\B_{x}(s,S)$  such that
+For any secondary structure $\Struct$, the pseudo-Boltzmann factor of a sequence $s$ is $\B^{[x]}_{\Struct}(s)$  such that
 \begin{equation}
-\B_{x, \Target}(s) = e^{\frac{-\ES(s,\Target)}{RT}}\cdot x^{\gc(s)}
+\B_{\Target}^{[x]}(s) = e^{\frac{-\ES(s,\Target)}{RT}}\cdot x^{\gc(s)}
 \label{def:genBoltz}
@@
 Summing the pseudo-Boltzmann factor over all possible sequences of a given length $|\Target|$, one obtains the pseudo-partition function $\mathcal{Z}_{x,\Target}$, from which one defines the pseudo-Boltzmann probability $\Prob_{x,\Target}(s)$ of each sequence $s$, respectively such that
-\begin{align}\mathcal{Z}_{x,\Target} &= \sum_{\substack{|s|=|\Target|}}\B_{x,\Target}(s)& \text{and}&&
+\begin{align}\mathcal{Z}_{x,\Target} &= \sum_{\substack{|s|=n}}\B_{x,\Target}(s)& \text{and}&&
 \Prob_{x,\Target}(s) &= \frac{\B_{x,\Target}(s)}{\mathcal{Z}_{x,\Target}}.\label{def:distribution}\end{align}
@@
 
-Firstly, a dynamic programming algorithm computes $\mathcal{Z}\substack{(i,j)\\ [a,b]}$ the pseudo-partition function restricted to an interval $[i,j]$, assuming its (previously chosen) flanking nucleotides are $a$ and $b$ respectively.
+Firstly, a dynamic programming algorithm computes $\Z{i,j}{a,b}$ the pseudo-partition function restricted to an interval $[i,j]$, assuming its (previously chosen) flanking nucleotides are $a$ and $b$ respectively.
 Remark that the empty interval only supports the empty structure, having energy $0$, so one has
@@
 
-Despite its apparent quadratic complexity, the recurrence described in Subsection~\ref{sec:pf} can be computed in linear $\Theta(|\Target|)$ time and space, owing to the fact that, on any recursive call, the position $k$ is entirely determined by $i$.
+Despite its apparent quadratic complexity, the recurrence described in Subsection~\ref{sec:pf} can be computed in linear $\Theta(n)$ time and space, owing to the fact that, on any recursive call, the position $k$ is entirely determined by $i$.
 
-\begin{theorem}Starting from the complete region $[0,n-1]$, Equation~\eqref{eq:Z_rec} can be computed in $\Theta(|\Target|)$ time and space complexity.
+\begin{theorem}Starting from the complete region $[0,n-1]$, Equation~\eqref{eq:Z_rec} can be computed in $\Theta(n)$ time and space complexity.
 \end{theorem}
@@
 }
-Since, in any interval $[i,j]$ satisfying the property, the ending position $j$ is completely determined by $i$, then there exists at most $\Theta(|\Target|)$ intervals that need to be evaluated, and we conclude on the overall complexity with the remark that $a$, $b$, $a'$ and $b'$ vary over bounded ranges and only contribute to a constant factor in the complexity.
+Since, in any interval $[i,j]$ satisfying the property, the ending position $j$ is completely determined by $i$, then there exists at most $\Theta(n)$ intervals that need to be evaluated, and we conclude on the overall complexity with the remark that $a$, $b$, $a'$ and $b'$ vary over bounded ranges, only contributing to a constant factor in the complexity.
 \end{proof}
@@
 \subsubsection{Stochastic backtrack}
-Once the pseudo-partition functions have been memorized, a stochastic backtrack starts from the complete interval $[0,|\Target|-1]$ and, at each step, chooses a suitable assignment for one or several positions, using probabilities derived from the precomputation, as illustrated by Figure~\ref{fig:stochastic}.
+Once the pseudo-partition functions have been memorized, a stochastic backtrack starts from the complete interval $[0,n-1]$ and, at each step, chooses a suitable assignment for one or several positions, using probabilities derived from the precomputation, as illustrated by Figure~\ref{fig:stochastic}.
 One or several recursive calls over the remaining regions are then performed, as described in Algorithm~\ref{alg:back}.
@@
 Moreover, the number of executions of each loops is bounded by a constant.
-Consequently, the complexity of Algorithm~\ref{alg:back} is in $\Theta(|\Target|)$ time and space.
+Consequently, the complexity of Algorithm~\ref{alg:back} is in $\Theta(n)$ time and space.
 
@@
 
-Regarding the complexity, it was previously established~\cite{Waldispuhl2011} that, for each value of $x$, there exists constants $\mu_x$ and $\sigma_x$ such that the distribution of \GCContent asymptotically converges towards a normal law having expectation in $\mu\cdot n\cdot(1+o(1))$ and standard deviation in $\sigma\cdot\sqrt{n}\cdot(1+o(1))$.
-Furthermore, the concentration of the distribution, asserted by its limited standard deviation, implies a growth in $\Theta(\sqrt{n})$ (resp.
+Regarding the overall complexity of the algorithm, it was previously established~\cite{Waldispuhl2011} that, for each value of $x$, there exists constants $\mu_x$ and $\sigma_x$ such that the distribution of \GCContent asymptotically converges towards a normal law having expectation in $\mu_x\cdot n\cdot(1+o(1))$ and standard deviation in $\sigma_x\cdot\sqrt{n}\cdot(1+o(1))$.
+Furthermore, the concentration of the distribution, asserted by its limited standard deviation, therefore the expected number of attempts required to generate a valid sequence when $\kappa=0$ (resp.
+$\kappa\in\Omega(1/\sqrt n)$) grows linke $\Theta(\sqrt{n})$ (resp.
 $\Theta(1)$, i.e.
-constant) for the expected number of attempts required to generate a valid sequence when $\kappa=0$ (resp.
-$\kappa\in\Omega(1/\sqrt n)$).
+a constant), leading to the announced complexities.
 

### 2013-03-08 ISMB/methods_ISMB.tex
Half way through tree rewriting of algorithm

@@
 \begin{equation}
-\B_{\Target}^{[x]}(s) = e^{\frac{-\ES(s,\Target)}{RT}}\cdot x^{\gc(s)}
+\B_{\Struct}^{[x]}(s) = e^{\frac{-\ES(s,\Struct)}{RT}}\cdot x^{\gc(s)}
 \label{def:genBoltz}
@@
 
-Summing the pseudo-Boltzmann factor over all possible sequences of a given length $|\Target|$, one obtains the pseudo-partition function $\mathcal{Z}_{x,\Target}$, from which one defines the pseudo-Boltzmann probability $\Prob_{x,\Target}(s)$ of each sequence $s$, respectively such that
-\begin{align}\mathcal{Z}_{x,\Target} &= \sum_{\substack{|s|=n}}\B_{x,\Target}(s)& \text{and}&&
-\Prob_{x,\Target}(s) &= \frac{\B_{x,\Target}(s)}{\mathcal{Z}_{x,\Target}}.\label{def:distribution}\end{align}
+Summing the pseudo-Boltzmann factor over all possible sequences of a given length $|\Struct|$, one obtains the pseudo-partition function $\mathcal{Z}_{\Struct}^{[x]}$, from which one defines the pseudo-Boltzmann probability $\Prob_{\Struct}^{[x]}(s)$ of each sequence $s$, respectively such that
+\begin{align}\mathcal{Z}_{\Struct}^{[x]} &= \sum_{\substack{|s|=n}}\B_{\Struct}^{[x]}(s)& \text{and}&&
+\Prob_{\Struct}^{[x]}(s) &= \frac{\B_{\Struct}^{[x]}(s)}{\mathcal{Z}_{\Struct}^{[x]}}.\label{def:distribution}\end{align}
 
@@
 
-Firstly, a dynamic programming algorithm computes $\Z{i,j}{a,b}$ the pseudo-partition function restricted to an interval $[i,j]$, assuming its (previously chosen) flanking nucleotides are $a$ and $b$ respectively.
-Remark that the empty interval only supports the empty structure, having energy $0$, so one has
+Firstly, a dynamic programming algorithm computes $\Z{\N,\Struct}{a,b}$ the pseudo-partition function ($x$ is omitted here for the sake of clarity) of a structure $\Struct$, assuming its (previously chosen) flanking nucleotides are $a$ and $b$ respectively, directly nested ($\N=T$) or not ($\N=F$) by some base-pair.
+Remark that the empty structure only supports the empty sequence, having energy $0$, so one has
 \begin{equation}
-	\forall i \in [0,n-1]:\, \Z{i+1,i}{a,b}=1.
+	\Z{T,\varepsilon}{a,b}=\Z{F,\varepsilon}{a,b}=1.
 	\label{eq:Z_in}
@@
 
-The recursion consists in four terms, depending on the base-pairing status and context of $(i,j)$:
-\begin{equation}
-	\Z{i,j}{a,b}:=\left\{
-  \begin{array}{ll}
-  		\displaystyle
+The recursion considers different case, depending on whether of the first position in $\Struct$ is:
+\begin{itemize}
+\item Unpaired ($\Struct = \ub \Struct'$):
+\begin{align}
+	\Z{T,\ub \Struct'}{a,b} = \Z{F,\ub \Struct'}{a,b} &:=
       \sum_{a'\in \B}  
       x^{\gc(a')}
-      \cdot\Z{i+1,j}{a',b} &\begin{array}{@{}l@{}}\text{If }\Target_{i}=-1,\\ \text{\relsize{-1}[Position $i$ unpaired]}\end{array}\\
-      \displaystyle
+      \cdot\Z{F,\Struct'}{a',b}; 
+\end{align}
+\item Paired to last position, stacking onto an existing nesting pair ($\Struct = \op \Struct' \cp$ and $\N=T$):
+\begin{align}
+	\Z{T,\op \Struct' \cp}{a,b} &:=
       \sum_{a',b'\in \B^2}
@@
 			 \cdot e^{\frac{-\ES^{\beta}_{ab \to a'b'}}{RT}}
-			 \cdot \Z{i+1,j-1}{a',b'}&
-\begin{array}{@{}l@{}}\text{Elif }\Target_i=j \land \Target_{i-1}=j+1,\\ \text{\relsize{-1}[$(i,j)$ stacking pair onto $(i$-$1,j$+$1)$]}\end{array}
-			 \\
+			 \cdot \Z{T,\Struct'}{a',b'}
+\end{align}
+\item Paired to an internal position or no stacking pair ($\Struct = \op \Struct' \cp \Struct''$ and $\N=F$, or $\Struct''\neq \varepsilon$):
+\begin{align}
+	\Z{\N,\op \Struct' \cp \Struct''}{a,b} &:=
 			 \displaystyle
@@
       \cdot\Z{i+1,k-1}{a',b'}
-      \cdot\Z{k+1,j}{b',b} & 
-\begin{array}{@{}l@{}}\text{Elif }\Target_i=k \land i < k \leq j,\\
-\text{\relsize{-1}[Junction/bulge $\Rightarrow$ split, no stack]}
-\end{array}
-\\
-      0 &\text{Otherwise.}
-	\end{array}\right.
+      \cdot\Z{k+1,j}{b',b} 
 \label{eq:Z_rec}
-\end{equation}
+\end{align}
+\end{itemize}
 
@@
 
-\begin{theorem}Starting from the complete region $[0,n-1]$, Equation~\eqref{eq:Z_rec} can be computed in $\Theta(n)$ time and space complexity.
+\begin{theorem}Starting from the target structure $\Target$, Equations~\eqref{eq:Z_in} to \eqref{eq:Z_rec} can be computed in $\Theta(n)$ time and space complexity.
 \end{theorem}
 \begin{proof}
-
-Consider the {\bf covering base-pair} $\CNBP{i}$ of a position $i$, defined as the base-pair $(i',j')$ such that $i'<i<j'$ and which minimizes $i-i'$.
-For the sake of simplicity, and without loss of generality, let us assume that there exists a virtual base-pair $(-1,n)$ in $\Target$, so that any position in $[0,n-1]$ has a covering base-pair.
-Then, it is easily shown by induction that, starting from the complete interval $[0,n-1]$, the recursive calls only evaluate $\mathcal{Z}$ over intervals $[i,j]$ such that $j=j'-1$, $\CNBP{i}=[i',j']$.
-\CFramed{This property clearly holds initially for the complete interval $[0,n-1]$, owing to the virtual base-pair $(-1,n)$. Furthermore, if it holds for a given $(i,j)$, $\CNBP{i}=[i',j+1]$, then:
-\begin{itemize} 
-  \item If $i$ is unpaired (Case 1), then $\CNBP{i+1}=\CNBP{i}=[i',j+1]$, and the property is satisfied by the subsequent call over $[i+1,j]$. 
-  \item If $i$ is paired with $j$ and stacking onto $(i-1,j+1)$ (Case 2), then one has $\CNBP{i+1}=[i,j]$ and the  subsequent call over $[i+1,j-1]$ satisfies the property.
-  \item If $i$ is paired with some $k$ (Case 3), then $\CNBP{i+1}=[i,k]$ and $\CNBP{k+1}=[i',j+1]$.
-Therefore the subsequent calls, respectively over $[i+1,k-1]$ and $[k+1,j]$, both satisfy the property.
-\end{itemize}
-The case of an empty sequence does not cause a recursive call and must not be verified.
-}
-Since, in any interval $[i,j]$ satisfying the property, the ending position $j$ is completely determined by $i$, then there exists at most $\Theta(n)$ intervals that need to be evaluated, and we conclude on the overall complexity with the remark that $a$, $b$, $a'$ and $b'$ vary over bounded ranges, only contributing to a constant factor in the complexity.
+Clearly, there only exists a constant number of possible values for the vector $(a,b,\N)$, so the dependency in $n$ may only reside in $\Struct$.
+Reasoning on the length of the subintervals, one clearly see that the cumulated lengths XXXXXXXXXXXXXXXXXXXX
 \end{proof}
@@
 \subsubsection{Stochastic backtrack}
-Once the pseudo-partition functions have been memorized, a stochastic backtrack starts from the complete interval $[0,n-1]$ and, at each step, chooses a suitable assignment for one or several positions, using probabilities derived from the precomputation, as illustrated by Figure~\ref{fig:stochastic}.
+Once the pseudo-partition functions have been memorized, a stochastic backtrack starts from the complete interval $[0,n-1]$ with any exterior bases $[a,b]$ and, at each step, chooses a suitable assignment for one or several positions, using probabilities derived from the precomputation, as illustrated by Figure~\ref{fig:stochastic}.
 One or several recursive calls over the remaining regions are then performed, as described in Algorithm~\ref{alg:back}.

### 2013-03-08 ISMB/results_ISMB.tex
Update
Half way through tree rewriting of algorithm

@@
 
-{\em Définir critères d'évaluations !!!}
-
-A first criterion is which percentage of sequences have as MFE the target
-secondary structures. We are also interested by the number of structures
-for which we can design at least one sequence with the desired MFE.
-We show the results when using only \ourprog in 
- Figure~\ref{fig:mfe_struct_solved_noinverse}. As mentioned earlier, the high rate of failure
+A first criterion is the proportion of sequences accepting the target
+secondary structures as their MFE. A second criterion is the number of structures
+for which at least one sequence with the desired MFE was produced.
+Figure~\ref{fig:mfe_struct_solved_noinverse} shows the results obtained by \ourprog for these criteria. As mentioned earlier, the high rate of failure
 is due to the fact that no selection criterion is applied to
@@
 locally optimize the free nucleotides, we obtained the results 
-shown in Figure~\ref{fig:mfe_struct_solved}. We observed
+summarized by Figure~\ref{fig:mfe_struct_solved}. We observed
 that length is, in general, not a good predictor for the hardness of designing a structure. 

### 2013-03-09 ISMB/ResponseToReviewers-Incarnation-ISMB-2013.tex
Update

@@
 
-\title{A weighted sampling algorithm for the design of RNA sequences with targeted secondary structure and nucleotides distribution}
+\title{A weighted sampling algorithm for the design of RNA sequences with targeted secondary structure and nucleotides distribution\\Response to review round \#1}
 \author{Vladimir Reinharz$^1$, Yann Ponty$^{2,*}$, J\'er\^{o}me Waldisp\"{u}hl$^{1,*}$}
@@
 For most, source code packages are available.}
-\Answer{We systematically benchmarked {\tt IncaRNAtion} against most of the existing available tools, including NUPack~\cite{Zadeh:2011fk}, FRNAkenstein~\cite{Lyngso:2012vn},  RNA-SSD~\cite{Aguirre-Hernandez:2007kx}, Info-RNA~\cite{Busch:2006uq} and RNAexInv~\cite{Avihoo:2011fk}.
+\Answer{In addition to \RNAinverse~\cite{Hofacker:1994}, we systematically benchmarked \ourprog against an extended selection of available tools, including \NUPACK~\cite{Zadeh:2011fk}, \frankenstein~\cite{Lyngso:2012vn},  \RNASSD~\cite{Aguirre-Hernandez:2007kx}, \INFORNA~\cite{Busch:2006uq} and \RNAexinv~\cite{Avihoo:2011fk}.
 While the page limitations associated with a conference submission do not allow for a an exhaustive discussion, we now report in the revised manuscript some of the most striking results.
-TODO (Comment results)
+Mainly, with the exception of \RNASSD, none of the latest version of our competitor allows an explicit control over the \GCContent, resulting in designs that are usually highly concentrated in \GCContent (with means ranging from 50\% to 80\% depending on the software).
 }
@@
 \Answer{We added this feature to our implementation, which now accepts a constraint mask consisting in a sequence of IUPAC symbols (e.g.
-'N' for unconstrained position, or G for Guanine-only).
+'N' for unconstrained position, or G for Guanine-only\ldots).
 We also briefly mention this feature in the definition of our probabilistic model.}
@@
 \Comment{(4) Section 2.2.3 is titled "Rejecting unsuitable candidates". It explains that candidates tend to be quite suitable, but somewhere, you should actually tell who is rejected and when ... }
-\Answer{TODO}
+\Answer{We completely rewrote this section}
 

### 2013-03-09 ISMB/ResponseToReviewers-Incarnation-ISMB-2013.tex
method over (except typos)

@@
 Mainly, with the exception of \RNASSD, none of the latest version of our competitor allows an explicit control over the \GCContent, resulting in designs that are usually highly concentrated in \GCContent (with means ranging from 50\% to 80\% depending on the software).
+TODO (Jerome)
 }
@@
 \Answer{Absolutely.
-However, the point of this section was not to put the stress on this issue, but rather to show that our glocal approach, starting with a global sampling followed by a postprocessing of unpaired regions using \RNAinverse, does not {\em drift} too much towards regions of high \GCContent{}s.
+However, the point of this section was not to insist on this issue, but rather to show that our glocal approach, starting with a global sampling followed by a postprocessing of unpaired regions using \RNAinverse, does not {\em drift} too much towards regions of high \GCContent{}s.
 We rephrased the text of this section to make this conclusion appear more clearly, and changed the layout of Table 2 to make the shift in \GC more explicit.}
@@
 And are these values length-/structure-independent or do they have to be computed for each structure individually?}
-\Answer{TODO (Yann)}
+\Answer{These values are structure/\GCContent-dependent, and must be recomputed for any new input.
+We made this point clearer in the description of the adaptive sampling procedure, and the influence on the complexity of the number $M$ of iterations.
+Typically, 4 or 5 iterations are required to get a good approximation of the best value $x_{gc}$, and our binary search approach provably offers a linear convergence.}
 
@@
 \Comment{1) The $x^{\gc(s)}$ in equation (1) seems to me a novel contributing term that is of high importance. Perhaps its introduction should be highlighted more in the text.}
-\Answer{TODO (Yann)}
+\Answer{We tried to make this point clearer in our revised version.}
 
@@
 Why not introduce $F(s)$ for the folding space of a sequence, and $F^{-1}(S)$ for the inverse folding space of a structure, and simply use $F^{-1}(S)$ where this is what you mean?}
-\Answer{While we agree that this proposition would lead to a compact and well-defined mathematical definition of the problem, we prefer a implicit definition of the compatibility of a sequence to a given structure.
-Namely, we always allow every possible sequence of a length $|S|$ to be a member of the inverse folding space, and assign to it some  energy penalty (typically $+\infty$, or $\beta$) if it violates some constraints (restriction on allowed base-pairs, position-specific constraints...).
-The inverse folding space sequences is therefore implicitly defined as the set of sequences having finite energy, and our algorithm is arguably more prone to generalization, e.g.
+\Answer{While we agree that this proposition would lead to compact and well-defined mathematical statements, we prefer a implicit definition of the compatibility of a sequence to a given structure.
+Namely, we always allow every possible sequence of a length $|S|$ to be a member of the inverse folding space, possibly assigning some  energy penalty to it  (typically $+\infty$, or $\beta$) when some constraints are violated (restriction on allowed base-pairs, position-specific constraints...).
+The inverse folding space is therefore implicitly defined as the set of sequences having finite energy, and our algorithm is arguably more prone to generalization, e.g.
 to capture non-canonical base-pairs which are increasingly considered by the field of RNA bioinformatics.
@@
 This way it is more transparent what is going on, and it is obvious that space and time are linear in $n$, because the size of the tree is in $O(n)$.}
-\Answer{Great suggestion. TODO!}
+\Answer{Great suggestion.
+However, adopting a pure tree-like representation would require elaborating on the correspondence between trees and structures, and might lead to technicalities in the context of stacking base-pairs.
+Therefore, we completely rewrote our equations, figure and algorithm to adopt a classic dot-parenthesis representation.
+While the end result may be less elegant than tree-like representations, it has the merit of dropping the double subscripts, leading to a counterintuitive complexity claim in our original manuscript.}
 
@@
 \Comment{(4) Section 2.2.3 is titled "Rejecting unsuitable candidates". It explains that candidates tend to be quite suitable, but somewhere, you should actually tell who is rejected and when ... }
-\Answer{We completely rewrote this section}
+\Answer{We completely rewrote this section, which now starts by explicitly reminding the problem addressed by the algorithm.}
 

### 2013-03-09 ISMB/introduction_ISMB.tex
Typos + Changed sampling figure
method over (except typos)

@@
 Unlike \RNAensign, it does not require a seed sequence to initiate its search.
-Finally, \ourprog implements a novel algorithm using weighted sampling techniques~\cite{Bodini2010} that enables us to control, for the first time, \textit{explicitly} the \GCContent of the solution.
+Finally, \ourprog implements a novel algorithm based on weighted sampling techniques~\cite{Bodini2010} that enables us to control, for the first time, \textit{explicitly} the \GCContent of the solution.
 This functionality is essential because wild-type sequences within living organisms often present medium or low \GCContent, presumably to offer better transcription rates and/or structural plasticity.
-Previous programs do not allow to control this parameter and tend to output sequences with high \GCContent{}s.
+Previous programs do not allow to control this parameter and tend to output sequences having high \GCContent{}s~\cite{Lyngso:2012vn}.
 

### 2013-03-09 ISMB/methods_ISMB.tex
method over (except typos)

@@
 
-Firstly, a dynamic programming algorithm computes $\Z{\N,\Struct}{a,b}$ the pseudo-partition function ($x$ is omitted here for the sake of clarity) of a structure $\Struct$, assuming its (previously chosen) flanking nucleotides are $a$ and $b$ respectively, directly nested ($\N=T$) or not ($\N=F$) by some base-pair.
+Firstly, a dynamic programming algorithm computes $\Z{\N,\Struct}{a,b}$ the pseudo-partition function ($x$ is omitted here for the sake of clarity) for a structure $\Struct$, assuming its (previously chosen) flanking nucleotides are $a$ and $b$ respectively, either forming a closing base-pair ($\N=\BoolTrue$) or not ($\N=\BoolFalse$).
 Remark that the empty structure only supports the empty sequence, having energy $0$, so one has
 \begin{equation}
-	\Z{T,\varepsilon}{a,b}=\Z{F,\varepsilon}{a,b}=1.
+	\Z{\BoolTrue,\varepsilon}{a,b}=\Z{\BoolFalse,\varepsilon}{a,b}= e^{-0/RT} = 1.
 	\label{eq:Z_in}
@@
 
-The recursion considers different case, depending on whether of the first position in $\Struct$ is:
-\begin{itemize}
-\item Unpaired ($\Struct = \ub \Struct'$):
+The recursion must consider different cases, depending on whether the first position in $\Struct$ is:
+\begin{description}
+\item[{\bf Case 1.}] Unpaired ($\Struct = \ub \Struct'$):
 \begin{align}
-	\Z{T,\ub \Struct'}{a,b} = \Z{F,\ub \Struct'}{a,b} &:=
+	\Z{\BoolTrue,\ub \Struct'}{a,b} = \Z{\BoolFalse,\ub \Struct'}{a,b}  &:=
       \sum_{a'\in \B}  
       x^{\gc(a')}
-      \cdot\Z{F,\Struct'}{a',b}; 
+      \cdot\Z{\BoolFalse,\Struct'}{a',b}; 
+\label{eq:Z_unpaired}
 \end{align}
-\item Paired to last position, stacking onto an existing nesting pair ($\Struct = \op \Struct' \cp$ and $\N=T$):
+\item[{\bf Case 2.}] Paired to last position ($\Struct = \op \Struct' \cp$), stacking onto a pre-existing closing pair ($\N=\BoolTrue$):
 \begin{align}
-	\Z{T,\op \Struct' \cp}{a,b} &:=
+	\Z{\BoolTrue,\op \Struct' \cp}{a,b} &:=
       \sum_{a',b'\in \B^2}
@@
 			 \cdot e^{\frac{-\ES^{\beta}_{ab \to a'b'}}{RT}}
-			 \cdot \Z{T,\Struct'}{a',b'}
+			 \cdot \Z{\BoolTrue,\Struct'}{a',b'};
+\label{eq:Z_stack}
 \end{align}
-\item Paired to an internal position or no stacking pair ($\Struct = \op \Struct' \cp \Struct''$ and $\N=F$, or $\Struct''\neq \varepsilon$):
+\item[{\bf Case 3.}] Paired to some position ($\Struct = \op \Struct' \cp \Struct''$), but not stacking ($\N=\BoolFalse$ or $\Struct''\neq \varepsilon$):
 \begin{align}
@@
 			\cdot e^{\frac{-\ES^{\beta}_{\varnothing\to a'b'}}{RT}}
-      \cdot\Z{i+1,k-1}{a',b'}
-      \cdot\Z{k+1,j}{b',b} 
+      \cdot\Z{\BoolFalse,\Struct'}{a',b'}
+      \cdot\Z{\BoolTrue,\Struct''}{b',b}.
 \label{eq:Z_rec}
 \end{align}
-\end{itemize}
+\end{description}
 
-Despite its apparent quadratic complexity, the recurrence described in Subsection~\ref{sec:pf} can be computed in linear $\Theta(n)$ time and space, owing to the fact that, on any recursive call, the position $k$ is entirely determined by $i$.
-
-\begin{theorem}Starting from the target structure $\Target$, Equations~\eqref{eq:Z_in} to \eqref{eq:Z_rec} can be computed in $\Theta(n)$ time and space complexity.
-\end{theorem}
-\begin{proof}
-Clearly, there only exists a constant number of possible values for the vector $(a,b,\N)$, so the dependency in $n$ may only reside in $\Struct$.
-Reasoning on the length of the subintervals, one clearly see that the cumulated lengths XXXXXXXXXXXXXXXXXXXX
-\end{proof}
+Remark that the number of combinations of $a$, $b$ and $\N$ remains bounded by a constant, thus the complexity of computing $\Z{\N,\Struct}{a,b}$ mainly depends on the values taken by $\Struct$ upon subsequent recursive calls.
+Such values are entirely determined by $\Struct$ at any given step of the recursion, and their dependency can be summarized in a tree having $\Theta(|\Struct|)$.
+Therefore, the computation of $\Z{\N,\Target}{a,b}$ requires $\Theta(n)$ time and space using dynamic-programming.
 
@@
 \resizebox{\textwidth}{!}{\input{FigStochasticBacktrack.pgf}}
-\caption{Stochastic backtrack procedure over an interval $[i,j]$: Either position $i$ is left unpaired (top), a base-pair $(i,j)$ is formed, stacking onto $(i+1,j-1)$ (middle), or $i$ is paired somewhere else, defining two regions on which subsequent recursive calls are needed (bottom).
-Positions indicated in red are assigned during this stage of the backtrack.\label{fig:stochastic}}
+\caption{Stochastic backtrack procedure for a given substructure $\Struct$: Either the first position is left unpaired (top), a base-pair is formed between the two extremities, stacking onto an exterior base-pair (middle), or paired without creating a stacking, defining two regions on which subsequent recursive calls are needed (bottom).
+For the empty structure (omitted here), the empty sequence is returned.
+Positions indicated in red are assigned at the current stage of the backtrack.\label{fig:stochastic}}
 \end{figure}
-\subsubsection{Stochastic backtrack}
-Once the pseudo-partition functions have been memorized, a stochastic backtrack starts from the complete interval $[0,n-1]$ with any exterior bases $[a,b]$ and, at each step, chooses a suitable assignment for one or several positions, using probabilities derived from the precomputation, as illustrated by Figure~\ref{fig:stochastic}.
-One or several recursive calls over the remaining regions are then performed, as described in Algorithm~\ref{alg:back}.
-On each recursive call, the algorithm assigns at least one nucleotide to a -- previously unassigned -- position.
-Moreover, the number of executions of each loops is bounded by a constant.
-Consequently, the complexity of Algorithm~\ref{alg:back} is in $\Theta(n)$ time and space.
 
@@
 	\SetAlgoLined
-\SetKwFunction{Backtrack}{Backtrack}
+\SetKwFunction{Backtrack}{StocBack}
 \SetKwFunction{Random}{Random}
-	rand $\leftarrow$ \Random$\left(\Z{i,j}{a,b}\right)$\tcp*[r]{Draw random number in $[0,\Z{i,j}{a,b}[$}
- \lIf(\tcp*[f]{Empty sequence}){$j<i$}{\Return{$\varepsilon$}}
-	\Else{
-  k $\leftarrow S_i$\;
-	\If(\tcp*[f]{Unpaired base}){$(k = -1)$}{
+	rand $\leftarrow$ \Random$\left(\Z{\N,\Struct}{a,b}\right)$\tcp*[r]{Draw random number in $[0,\Z{\N,\Struct}{a,b}[$}
+ \Switch{}{
+   \lCase(\tcp*[f]{Empty structure $\Rightarrow$ empty sequence}){$\Struct=\varepsilon$}{\Return{$\varepsilon$}}
+   \Case(\tcp*[f]{First position is unpaired}){$\Struct=\ub\, \Struct'$}{		
    \For{$a'\in\B$}{
-			rand $\leftarrow$ rand $- x^{\gc(a')}\cdot \Z{i+1,j}{a',b}$\;
-			\lIf{$\text{\rm rand}<0$}\Return{$a'.\Backtrack\left(i+1,j,a',b,\Target\right)$}\;
+			rand $\leftarrow$ rand $- x^{\gc(a')}\cdot \Z{\BoolFalse,\Struct'}{a',b}$\;
+			\lIf{$\text{\rm rand}<0$}\Return{$a'.\Backtrack(a',b,\BoolFalse,\Struct')$}\;
   		}
    }
-	\lElse{
-		\If(\tcp*[f]{Stacking base pair}){$(k=j) \land (S_{i-1}=j+1)$}{
-			\For{$(a',b')\in\B\times\B$}{
+   \Case(\tcp*[f]{Extremities involved in stacking base pair}){$\Struct=\text{\rm\op}\, \Struct' \,\text{\rm\cp}$  {\rm \bf and} $\N=\BoolTrue$}
+   {		
+			\For{$(a',b')\in\B\times\B$}
+    {
 				rand $\leftarrow$ rand $-
 			 	x^{\gc(a'.b')}
-			 	\cdot e^{\frac{-\ES^{\beta}_{ab \to a'b'}}{RT}}
-			 	\cdot \Z{i+1,j-1}{a',b'}	$\;
+			 	\cdot e^{{-\ES^{\beta}_{ab \to a'b'}}/{RT}}
+			 	\cdot \Z{\BoolTrue,\Struct'}{a',b'}	$\;
 				\lIf{$\text{\rm rand}<0$}{
-					\Return{$a'.\Backtrack\left(i+1,j-1,a',b',\Target\right). b'$}\;		
+					\Return{$a'.\Backtrack(a',b',\BoolTrue,\Struct'). b'$}\;		
 				}
@@
 		}
-		\lElse{
-			\If(\tcp*[f]{Paired, not stacked}){$(\Target_i=k) \land (i < k \leq j)$}{
+   \Other(\tcp*[f]{First position is paired without a stacking pair})
+   {
+				\tcp{$\Struct=\text{\rm\op}\, \Struct' \,\text{\rm\cp}\,\Struct''$}
 				\For{$(a',b')\in\B\times\B$}{
@@
 					\cdot e^{\frac{-\ES^{\beta}_{\varnothing\to a'b'}}{RT}}
-  	    	\cdot\Z{i+1,k-1}{a',b'}
-	    	  \cdot\Z{k+1,j}{b',b}$\;	
- 					\If{{\rm rand} $<0$}{
+  	    	\cdot\Z{\BoolFalse,\Struct'}{a',b'}
+	    	  \cdot\Z{\BoolTrue,\Struct''}{b',b}$\;	
+ 					\lIf{{\rm rand} $<0$}{
 						\Return{$a'
-						.\Backtrack\left(i+1,k-1,a',b',\Target\right)
+						.\Backtrack\left(a',b',\BoolFalse,\Struct'\right)
 						.b'
-						.\Backtrack\left(k+1,j,b',b,\Target\right)
+						.\Backtrack\left(b',b,\BoolTrue,\Struct''\right)
 						$}\;	
@@
  }
-	}
- }
-\caption{\protect\Backtrack$\left(i,j,a,b,\Target\right)$\label{alg:back}}
+\caption{\protect\Backtrack$\left(a,b,\N,\Struct\right)$\label{alg:back}}
 \end{algorithm}
+\subsubsection{Stochastic backtrack}
+Once the pseudo-partition functions have been computed and memorized, a stochastic backtrack starts from the target structure $\Target$ with any exterior bases $[a,b]$ and no nesting base-pair, corresponding to a call \Backtrack$\left(\varnothing,\varnothing,\BoolFalse,\Target\right)$ to Algorithm~\ref{alg:back}.
+At each step, a suitable assignment for one or several positions is chosen, using probabilities derived from the precomputation, as illustrated by Figure~\ref{fig:stochastic}.
+One or several recursive calls over the appropriate substructures are then performed.
+On each recursive call, the algorithm assigns at least one nucleotide to a -- previously unassigned -- position.
+Moreover, the number of executions of each loops is bounded by a constant.
+Consequently, the complexity of Algorithm~\ref{alg:back} is in $\Theta(n)$ time and space.
 
@@
 
-Our goal is to produce a set of sequences whose \GCContent matches a prescribed value $gc$.
+Let us remind that our goal is to produce a set of sequences whose \GCContent matches a prescribed value $gc$.
 An absolute tolerance $\kappa$ may be allowed, so that the \GCContent of any valid sequence must fall in  
@@
 Since sequences of arbitrary \GCContent may be generated by Algorithm~\ref{alg:back}, we use a rejection-based approach~\cite{Bodini2010}, previously adapted by the authors in a similar context~\cite{Waldispuhl2011}.
-In our context, this gives an algorithm which generates $k$ valid sequences in expected time $\Theta(k\cdot n\sqrt{n})$ when $\kappa=0$ (resp.
-$\Theta(k\cdot n)$ when $\kappa$ is a positive constant) memory $\Theta(k\cdot n)$.
-Although a complete analysis can be found in an earlier contribution~\cite{Waldispuhl2011}, let us briefly outline the strategy and arguments used to analyze the complexity.
+This gives an algorithm which generates $k$ valid sequences in expected time $\Theta(k\cdot n\sqrt{n})$ when $\kappa=0$ (or $\Theta(k\cdot n)$ when $\kappa$ is a positive constant) and memory in $\Theta(k\cdot n)$.
+A complete analysis of the rejection process can be found in an earlier contribution~\cite{Waldispuhl2011}, but let us briefly outline the approach, and the main arguments used to establish its complexity.
 
-This approach simply generates sets of sequences by repeatedly running the stochastic backtrack algorithm.
+Our adaptive sampling approach simply generates sets of sequences by repeatedly running the stochastic backtrack algorithm.
 The average \GCContent induced by the current value of the $x$ parameter, can then be adequately estimated from the sample, or computed exactly using recent algorithmic advances~\cite{Ponty2011}.
@@
 In practice, a simple binary search~\cite{Waldispuhl2011} is used in our implementation, and typically converges after very few iterations.
-However the optimal value for $x$ can also be derived analytically using an interpolation after $\Theta(n)$ evaluations of $\Z{i,j}{a,b}$ for different candidate values of $x$, as previously noted~\cite{Waldispuhl2011} and implemented using the Fast-Fourier Transform~\cite{Senter2012}.
+An optimal value for $x$ can also be derived analytically using interpolation after $\Theta(n)$ evaluations of $\Z{i,j}{a,b}$ for different candidate values of $x$, as previously noted~\cite{Waldispuhl2011} and implemented using the Fast-Fourier Transform~\cite{Senter2012}.
 
-Regarding the overall complexity of the algorithm, it was previously established~\cite{Waldispuhl2011} that, for each value of $x$, there exists constants $\mu_x$ and $\sigma_x$ such that the distribution of \GCContent asymptotically converges towards a normal law having expectation in $\mu_x\cdot n\cdot(1+o(1))$ and standard deviation in $\sigma_x\cdot\sqrt{n}\cdot(1+o(1))$.
+Regarding the overall complexity claim for our approach, it was previously established~\cite{Waldispuhl2011} that, for each value of $x$, there exists constants $\mu_x$ and $\sigma_x$ such that the distribution of \GCContent asymptotically converges towards a normal law having expectation in $\mu_x\cdot n\cdot(1+o(1))$ and standard deviation in $\sigma_x\cdot\sqrt{n}\cdot(1+o(1))$.
 Furthermore, the concentration of the distribution, asserted by its limited standard deviation, therefore the expected number of attempts required to generate a valid sequence when $\kappa=0$ (resp.
-$\kappa\in\Omega(1/\sqrt n)$) grows linke $\Theta(\sqrt{n})$ (resp.
+$\kappa\in\Omega(1/\sqrt n)$) grows like $\Theta(\sqrt{n})$ (resp.
 $\Theta(1)$, i.e.
 a constant), leading to the announced complexities.
+Formally, since a suitable weight $x$ must be recomputed for each targeted structure and \GCContent, then the number $M$ of iterations required for the converge can be accounted for explicitly, leading to time complexities in $\Theta((M+\sqrt{n})\cdot k\cdot n)$ (if $\kappa=0 \Rightarrow$ no tolerance) and $\Theta(M\cdot k\cdot n)$ (if $\kappa>0$).
 
@@
 We specified a constraint mask to prevent stacking base-pairs from being modified and, whenever necessary, reestablished their content {\em a posteriori}, as \RNAinverse has been witnessed to take some liberties with constraints masks.
-As shown in Table~\ref{table:impact_on_gc} (Supplementary material), this postprocessing does not drastically alter the \GCContent, so the glocal approach reasonably addresses the contrained \GCContent design problem.
+As shown in Table~\ref{table:impact_on_gc} (Supplementary material), this postprocessing does not drastically alter the \GCContent, so the glocal approach reasonably addresses the constrained \GCContent design problem.

### 2013-03-09 ISMB/ResponseToReviewers-Incarnation-ISMB-2013.tex
Completed response to reviewers

@@
 \Comment{1) The $x^{\gc(s)}$ in equation (1) seems to me a novel contributing term that is of high importance. Perhaps its introduction should be highlighted more in the text.}
-\Answer{We tried to make this point clearer in our revised version.}
+\Answer{We tried to make this point clearer in our revised version.
+However, it should be noted that this contribution is not a goal in itself, but rather a mean to perform stochastic sampling with respect to the weighted distribution.
+This distribution has all sorts of desirable mathematical properties, among which the fact that induced distributions for additive features (here the \GCContent) are nicely concentrated, allowing for a rejection-based approach.}
 
@@
 Therefore, we completely rewrote our equations, figure and algorithm to adopt a classic dot-parenthesis representation.
-While the end result may be less elegant than tree-like representations, it has the merit of dropping the double subscripts, leading to a counterintuitive complexity claim in our original manuscript.}
+While the end result may be less elegant than expected with a tree-like representations, this rewrite has the merit of dropping the double subscripts, hopefully conveying a better intuition of the principles underlying the complexity claim in our original manuscript.}
 
@@
 \Comment{(5) recurring: "free-nucleotides in loops" -- isn't that the same, being "free" and in a loop? And: drop the hyphen.}
-\Answer{TODO}
+\Answer{The term of loop is sometimes used in a broader sense, which could be informally summarized as "a closed cyclic sequence of bases, connected either by the backbone or by base-pairs" (such that each connection but one goes in the $5'\to3'$ direction).
+This alternate definition was already used in the earlier works of M.
+Zuker, and maybe goes further back in the prehistory of RNA bioinformatics.
+Under this definition,  some of the bases involved in a loop may be paired, therefore we use the adjective to lift any ambiguity.
+}
 

### 2013-03-09 ISMB/results_ISMB.tex
some stuff in src
method over (except typos)

@@
 \subsection{Design}
- To benchmark our method, we first started by sampling about $1000$ sequences per structure. After correcting the free nucleotides with
+ To benchmark our method, we first started by sampling about $100$ sequences per structure. After correcting the free nucleotides with
  \texttt{RNAinverse} we computed their MFE with \texttt{RNAfold} from the \textit{Vienna Package 2.0}~\cite{Hofacker:1994}.
@@
 
-Also, a critical propertie in RNA sequence design is 
+Also, a critical property in RNA sequence design is 
 the frequency of the MFE. 

### 2013-03-09 ISMB/ResponseToReviewers-Incarnation-ISMB-2013.tex
One more typo for the road

@@
 We made this point clearer in the description of the adaptive sampling procedure, and the influence on the complexity of the number $M$ of iterations.
-Typically, 4 or 5 iterations are required to get a good approximation of the best value $x_{gc}$, and our binary search approach provably offers a linear convergence.}
+Typically, 4 or 5 iterations are required to get a good approximation of the best value $x_{gc}$, and our binary search approach provably offers a linear convergence (in term of the number of correct decimals) towards the closest representable value of $x_{gc}$.}
 
@@
 Therefore, we completely rewrote our equations, figure and algorithm to adopt a classic dot-parenthesis representation.
-While the end result may be less elegant than expected with a tree-like representations, this rewrite has the merit of dropping the double subscripts, hopefully conveying a better intuition of the principles underlying the complexity claim in our original manuscript.}
+While the end result may be less elegant than expected with a tree-like presentation, this rewrite has the merit of dropping the double subscripts, hopefully better conveying the  intuition underlying the complexity claim in our original manuscript.}
 

### 2013-03-10 ISMB/ResponseToReviewers-Incarnation-ISMB-2013.tex
updated answer to reviewers (Qs to Yann)

@@
 Mainly, with the exception of \RNASSD, none of the latest version of our competitor allows an explicit control over the \GCContent, resulting in designs that are usually highly concentrated in \GCContent (with means ranging from 50\% to 80\% depending on the software).
-TODO (Jerome)
+\textbf{FIXME: What should I add here?
+I think Vlad already added the histogram for the distribution of GC already (?)}
 }
@@
 \Comment{Is the method available?}
-\Answer{TODO (Jerome)}
+\Answer{\textbf{FIXME: Do you prefer a redirection to the git-hub repo or should we setup a basic web server?}}
 
@@
 However, it should be noted that this contribution is not a goal in itself, but rather a mean to perform stochastic sampling with respect to the weighted distribution.
-This distribution has all sorts of desirable mathematical properties, among which the fact that induced distributions for additive features (here the \GCContent) are nicely concentrated, allowing for a rejection-based approach.}
+This distribution has all sorts of desirable mathematical properties, among which the fact that induced distributions for additive features (here the \GCContent) are nicely concentrated, allowing for a rejection-based approach.
+\textbf{FIXME: Should we really says it's not a goal. This seems to diminish our work... I'll try to find a way to re-formulate that. I agree with the content though.}
+}
 
@@
 While such a postprocessing tends to reestablish \GCContent{}s that are  closer to 50\% \GC, this shift remains reasonable (at most ~6\% \GC).
-Furthermore, this shift is very predictable and offers the user the possibility to \emph{overshoot} for a more extreme \GCContent with the global approach, knowing that the postprocessing step will tend to average it.}
+Furthermore, this shift is very predictable and offers the user the possibility to \emph{overshoot} for a more extreme \GCContent with the global approach, knowing that the postprocessing step will tend to average it.
+\textbf{FIXME: We could eventually emphasize that we bring \RNAinverse in a region which is likely to statisfy the  MFE criterion and thus will require only minimal adjustments.}}
 
@@
 to capture non-canonical base-pairs which are increasingly considered by the field of RNA bioinformatics.
+\textbf{FIXME: Wow... that's the first time I see you refusing to be more formal :) }
 }
@@
 Under this definition,  some of the bases involved in a loop may be paired, therefore we use the adjective to lift any ambiguity.
+\textbf{FIXME: Alternatively we could use the term of ``unpaired nucleotides''?}
 }

### 2013-03-10 ISMB/main_ISMB.tex
Typos + Changed sampling figure
method over (except typos)
Typos
bioinfo template

@@
-\documentclass[11pt]{article}
-\usepackage[top=1in, bottom=1in,right=1in,left=1in]{geometry}                % See geometry.pdf to learn the layout options. There are lots.
-\geometry{letterpaper}                   % ... or a4paper or a5paper or ... 
+\documentclass{bioinfo}
+\copyrightyear{2013}
+\pubyear{2013}
+%\usepackage[top=1in, bottom=1in,right=1in,left=1in]{geometry}                % See geometry.pdf to learn the layout options. There are lots.
+%\geometry{letterpaper}                   % ... or a4paper or a5paper or ... 
 %\geometry{landscape}                % Activate for for rotated page geometry
 %\usepackage[parfill]{parskip}    % Activate to begin paragraphs with an empty line rather than an indent
-\usepackage[utf8]{inputenc}
+%\usepackage[utf8]{inputenc}
 \usepackage{graphicx}
@@
 \usepackage{relsize}
-\usepackage[numbers]{natbib}
+\usepackage{natbib}
 \usepackage[colorinlistoftodos]{todonotes}
@@
 \usepackage{relsize}
+\usepackage{multirow}
+\usepackage{cleveref}
+\usepackage{url}
+
 \usetikzlibrary{positioning}
@@
 \newtheorem{corollary}[theorem]{Corollary}
+\newtheorem{description}{Description}
 
-\newcommand{\ourprog}{\texttt{IncaRNAtion}\xspace}
+\crefmultiformat{equation}{Equations~(#2#1#3)}%
+{ and~(#2#1#3)}{, (#2#1#3)}{ and~(#2#1#3)}
 
@@
 
-\newcommand{\RNAmutants}{\texttt{RNAmutants}\xspace}
-\newcommand{\RNApyro}{\texttt{RNApyro}\xspace}
-\newcommand{\RNAinverse}{\texttt{RNAinverse}\xspace}
-\newcommand{\RNASSD}{\texttt{RNA-SSD}\xspace}
-\newcommand{\INFORNA}{\texttt{INFO-RNA}\xspace}
-\newcommand{\NUPACK}{\texttt{NUPACK}\xspace}
-\newcommand{\RNAiFOLD}{\texttt{RNAiFOLD}\xspace}
-\newcommand{\frankenstein}{\texttt{Frnakenstein}\xspace}
-\newcommand{\RNAexinv}{\texttt{RNAexinv}\xspace}
-\newcommand{\rnaDesign}{\texttt{rnaDesign}\xspace}
-\newcommand{\RNAensign}{\texttt{RNA-ensign}\xspace}
-\newcommand{\GCContent}{\Gb\Cb-content\xspace}
-\newcommand{\SFold}{\texttt{SFold}\xspace}
+\input{macros}
 
-\newcommand{\red}[1]{{\color{red}#1}}
-\newcommand{\farna}{\texttt{FARNA}\xspace}
-\newcommand{\mcfoldmcsym}{\texttt{MC-Pipeline}\xspace}
-\newcommand{\mcfold}{\texttt{MC-Fold}\xspace}
-\newcommand{\mcsym}{\texttt{MC-Sym}\xspace}
-\newcommand{\nast}{\texttt{NAST}\xspace}
-\newcommand{\ifoldrna}{\texttt{iFoldRNA}\xspace}
-\newcommand{\rnafold}{\texttt{RNAfold}\xspace}
-\newcommand{\rnasubopt}{\texttt{RNAsubopt}\xspace}
-\newcommand{\rnawolf}{\texttt{RNAwolf}\xspace}
-\newcommand{\rnastructure}{\texttt{RNAstructure}\xspace}
-\newcommand{\contrafold}{\texttt{contrafold}\xspace}
-\newcommand{\unafold}{\texttt{unafold}\xspace}
-\newcommand{\rnadd}{\texttt{RNA2D3D}\xspace}
-\newcommand{\assemble}{\texttt{assemble}\xspace}
-\newcommand{\fred}{\texttt{FR3D}\xspace}
-\newcommand{\rnajunction}{\texttt{RNAjunction}\xspace}
-\newcommand{\rnamotif}{\texttt{RNAmotif}\xspace}
-\newcommand{\treefolder}{\texttt{TreeFolder}\xspace}
-\newcommand{\barnacle}{\texttt{BARNACLE}\xspace}
-\newcommand{\contextfold}{\texttt{contextfold}\xspace}
-
-\newcommand{\RNASTRAND}{{\sf RNA STRAND}\xspace}
-
-%\newcommand{\citep}{\cite}
-\newcommand{\Z}[2]{\mathcal{Z}{\substack{(#1)\\ [#2]}}}
-\newcommand{\Y}[3]{\mathcal{Y}_{\substack{(#1)\\ [#3]}}^{#2}}
-\newcommand{\B}{\mathcal{B}}
-\newcommand{\Kron}{\delta}
-\newcommand{\ub}{\bullet}
-
-\newcommand{\Ab}{{\sf{A}}\xspace}
-\newcommand{\Cb}{{\sf{C}}\xspace}
-\newcommand{\Gb}{{\sf{G}}\xspace}
-\newcommand{\Ub}{{\sf{U}}\xspace}
-
-\newcommand{\gc}{{\#\text{{\sf gc}}}}
-\newcommand{\PE}[1]{E(#1)}
-\newcommand{\EI}{\text{EI}}
-\newcommand{\ES}{\text{{\sf Es}}}
-\newcommand{\ISO}{\text{ISO}}
-\newcommand{\Prob}{\text{P}}
-\newcommand{\Target}{\text{\sf S}}
-\newcommand{\CNBP}[1]{\gamma(#1)}
-
-\newcommand{\CFramed}[2][gray]{\begin{tcolorbox}[colframe=#1,title=To be included only if space allows\ldots,colback=white]#1\end{tcolorbox}}
-\renewcommand{\CFramed}[2][gray]{}
+\begin{document}
+\firstpage{1}
 
 \title{A weighted sampling algorithm for the design of RNA sequences with targeted secondary structure and nucleotides distribution}
-\author{Vladimir Reinharz$^1$, Yann Ponty$^{2,*}$, J\'er\^{o}me Waldisp\"{u}hl$^{1,*}$}
-\date{$^1$ School of Computer Science, McGill University, Montreal, Canada, $^2$ Laboratoire d'informatique, \'Ecole Polytechnique, Palaiseau, France.\\ \small $^*$Corresponding authors: \texttt{jeromew@cs.mcgill.ca}, \texttt{yann.ponty@lix.polytechnique.fr}}
+\author{Vladimir Reinharz$^1$, Yann Ponty$^{2,*}$, J\'er\^{o}me Waldisp\"{u}hl$^{1,*}$\footnote{to whom correspondence should be addressed}}
+\address{$^1$ School of Computer Science, McGill University, Montreal, Canada\\ $^2$ Laboratoire d'informatique, \'Ecole Polytechnique, Palaiseau, France.}
 
-%Last thing before BEGIN DOC for links to appear in pdf
+\history{Received on XXXXX; revised on XXXXX; accepted on XXXXX}
 
-%  \usepackage{hyperref}
-%  \hypersetup{colorlinks}
-%              citecolor=black,
-%              filecolor=black%
-%              linkcolor=black%
-%              urlcolor=black%
-%              pdftex}
+\editor{Associate Editor: XXXXXXX}
 
-\begin{document}
 \maketitle
@@
 \begin{abstract}
-\noindent
-\textbf{Motivations:} The design of RNA sequences folding into predefined secondary structures is a milestone for many synthetic biology and gene therapy studies.
+\section{Motivations:} The design of RNA sequences folding into predefined secondary structures is a milestone for many synthetic biology and gene therapy studies.
 Most of the current software uses similar local search strategies (i.e.
 a random seed is progressively adapted to acquire the desired folding propertues) and more importantly do not allow the user to control explicitly the nucleotide distribution such as the \GCContent in their sequences.
-However, the latter is an important criterion for large-scale applications as it could presumably be used to design sequences with better transcription rates and/or structural plasticity.\\
-\noindent
-\textbf{Results:}
+However, the latter is an important criterion for large-scale applications as it could presumably be used to design sequences with better transcription rates and/or structural plasticity.
+
+\section{Results:}
 In this paper, we introduce \ourprog, a novel algorithm to design RNA sequences folding into target secondary structures with a predefined nucleotide distribution.
@@
 To complete this study, we develop an hybrid method combining our global sampling approach with local search strategies.
-Remarkably, our glocal methodology overcomes both local and global approaches  for sampling sequences with a specific GC content and target structure.\\
+Remarkably, our glocal methodology overcomes both local and global approaches  for sampling sequences with a specific GC content and target structure.
+\section{Availability:} 
+Our method is freely available at WEBSITE INCARNATION
+\section{Contact:} \href{jeromew@cs.mcgill.ca}{jeromew@cs.mcgill.ca}, \href{yann.ponty@lix.polytechnique.fr}{yann.ponty@lix.polytechnique.fr}\\
 \noindent
@@
 
-\bibliographystyle{plainnat}
+\bibliographystyle{natbib}
 \bibliography{RNApyro}

### 2013-03-10 ISMB/results_ISMB.tex
Typos
bioinfo template

@@
 \label{sec:implementation}
-Our software, \ourprog, was implemented in {\tt Python~2.7}. We used
+Our software, \ourprog, is available at \oururl and was implemented in {\tt Python~2.7}. We used
 \RNAinverse from the \textit{Vienna Package 2.0}~\cite{Hofacker:1994}.
+
 All time benchmarks were run on a single AMD Opteron(tm) 6278 Processor  at 2.4 GHz with cache of 512 KB.
@@
 	\centering
-	\includegraphics[width=\textwidth]{Figures/time_rnastrand_clustered_rnainverse_100samples_fix}
+ 	\includegraphics[width=0.5\textwidth]{Figures/time_rnastrand_clustered_rnainverse_100samples_fix}
 	\caption{Average time in seconds to generate one sequence. We explicitly show 
@@
  	\centering
-	\includegraphics[width=\textwidth]{Figures/bins_distribution.png}
+	\includegraphics[width=0.5\textwidth]{Figures/bins_distribution.png}
 	\caption{Number of secondary structures per bin, according to our three clustering criteria.}
@@
 	\centering
-	\includegraphics[width=\textwidth]{Figures/mfe_struct_solve_nornainverse.png}
+	\includegraphics[width=0.5\textwidth]{Figures/mfe_struct_solve_nornainverse.png}
 	\caption{Quality of \ourprog results. The first row shows the percentage
@@
 	\centering
-	\includegraphics[width=\textwidth]{Figures/mfe_struct_solved}
+	\includegraphics[width=0.5\textwidth]{Figures/mfe_struct_solved}
 	\caption{Quality of \texttt{Incarnation+RNAinverse} results. The first row shows the percentage
@@
  	\centering
-	\includegraphics[scale=0.45]{Figures/rnastrand_clustered_rnainverse_100samples_struct_sens.png}
+	\includegraphics[width=0.5\textwidth]{Figures/rnastrand_clustered_rnainverse_100samples_struct_sens.png}
 	\caption{Structural sensitivity (i.e. $\#$ well predicted base pairs / $\#$ base pairs in target) of the sampled sequences MFE. }
@@
 	\centering
-	\includegraphics[width=\textwidth]{Figures/nb_sols_entropy.png}
+	\includegraphics[width=0.5\textwidth]{Figures/nb_sols_entropy.png}
 	\caption{Number of solutions generated and their average entropy. 
@@
 	\centering
-	\includegraphics[width=\textwidth]{Figures/freq.png}
+	\includegraphics[width=0.5\textwidth]{Figures/freq.png}
 	\caption{Average Boltzmann probability of the MFE.}
@@
 
-\label{fig:rnainverse}
+%\label{fig:rnainverse}

### 2013-03-10 ISMB/results_ISMB.tex
fix figs

--- 2013-03-09 ISMB/results_ISMB.tex
@@
 
-\begin{figure}[t!]
+\begin{figure*}[t!]
 	\centering
@@
 	\label{fig:time}	
-\end{figure}
+\end{figure*}
 
@@
 
- \begin{figure}[ht!]
+ \begin{figure*}[ht!]
  	\centering
@@
 	\label{fig:bins}
- \end{figure}
+ \end{figure*}
 
@@
 
-\begin{figure}[ht!]	
+\begin{figure*}[ht!]	
 	\centering
@@
 	\label{fig:mfe_struct_solved_noinverse}	
-\end{figure}
+\end{figure*}
 
-\begin{figure}[ht!]	
+\begin{figure*}[ht!]	
 	\centering
@@
 	\label{fig:mfe_struct_solved}	
-\end{figure}
+\end{figure*}
 
@@
 
-\begin{figure}[ht!]
+\begin{figure*}[ht!]
  	\centering
-	\includegraphics[scale=0.45]{Figures/rnastrand_clustered_rnainverse_100samples_struct_sens.png}
+	\includegraphics[width=\textwidth]{Figures/rnastrand_clustered_rnainverse_100samples_struct_sens.png}
 	\caption{Structural sensitivity (i.e. $\#$ well predicted base pairs / $\#$ base pairs in target) of the sampled sequences MFE. }
 	\label{fig:ss_sens}	
-\end{figure}
+\end{figure*}
 
@@
 
-\begin{figure}[ht!]
+\begin{figure*}[ht!]
 	\centering
@@
 	\label{fig:nb_sols_entropy}
-\end{figure}
+\end{figure*}
 
-\begin{figure}[ht!]
+\begin{figure*}[ht!]
 	\centering
@@
 	\label{fig:freq}
-\end{figure}
+\end{figure*}
 

### 2013-03-10 ISMB/ResponseToReviewers-Incarnation-ISMB-2013.tex
Typo in equations + Formatting

@@
 Furthermore, this shift is very predictable and offers the user the possibility to \emph{overshoot} for a more extreme \GCContent with the global approach, knowing that the postprocessing step will tend to average it.
-\textbf{FIXME: We could eventually emphasize that we bring \RNAinverse in a region which is likely to statisfy the  MFE criterion and thus will require only minimal adjustments.}}
+\textbf{FIXME: We could eventually emphasize that we bring \RNAinverse in a region which is likely to satisfy the  MFE criterion and thus will require only minimal adjustments.}}
 

### 2013-03-10 ISMB/methods_ISMB.tex
bioinfo template
fix figs
Typo in equations + Formatting

@@
 The recursion must consider different cases, depending on whether the first position in $\Struct$ is:
-\begin{description}
-\item[{\bf Case 1.}] Unpaired ($\Struct = \ub \Struct'$):
+\begin{itemize}
+\item {\bf Case 1.} Unpaired ($\Struct = \ub \Struct'$):
 \begin{align}
@@
 \end{align}
-\item[{\bf Case 2.}] Paired to last position ($\Struct = \op \Struct' \cp$), stacking onto a pre-existing closing pair ($\N=\BoolTrue$):
+\item {\bf Case 2.} Paired to last position ($\Struct = \op \Struct' \cp$), stacking onto a pre-existing closing pair ($\N=\BoolTrue$):
 \begin{align}
@@
 \end{align}
-\item[{\bf Case 3.}] Paired to some position ($\Struct = \op \Struct' \cp \Struct''$), but not stacking ($\N=\BoolFalse$ or $\Struct''\neq \varepsilon$):
+\item {\bf Case 3.} Paired to some position ($\Struct = \op \Struct' \cp \Struct''$), but not stacking ($\N=\BoolFalse$ or $\Struct''\neq \varepsilon$):
 \begin{align}
@@
 			\cdot e^{\frac{-\ES^{\beta}_{\varnothing\to a'b'}}{RT}}
-      \cdot\Z{\BoolFalse,\Struct'}{a',b'}
-      \cdot\Z{\BoolTrue,\Struct''}{b',b}.
+      \cdot\Z{\BoolTrue,\Struct'}{a',b'}
+      \cdot\Z{\BoolFalse,\Struct''}{b',b}.
 \label{eq:Z_rec}
 \end{align}
-\end{description}
+\end{itemize}
 
@@
 
-\begin{figure}
-\resizebox{\textwidth}{!}{\input{FigStochasticBacktrack.pgf}}
+\begin{figure*}
+\resizebox{\linewidth}{!}{\input{FigStochasticBacktrack.pgf}}
 \caption{Stochastic backtrack procedure for a given substructure $\Struct$: Either the first position is left unpaired (top), a base-pair is formed between the two extremities, stacking onto an exterior base-pair (middle), or paired without creating a stacking, defining two regions on which subsequent recursive calls are needed (bottom).
@@
 Positions indicated in red are assigned at the current stage of the backtrack.\label{fig:stochastic}}
-\end{figure}
+\end{figure*}
 
@@
 	\SetAlgoLined
-\SetKwFunction{Backtrack}{StocBack}
+\SetKwFunction{Backtrack}{SB}
 \SetKwFunction{Random}{Random}

### 2013-03-10 ISMB/main_ISMB.tex
fix availability
Typo in equations + Formatting
Protected biblio

@@
 \title{A weighted sampling algorithm for the design of RNA sequences with targeted secondary structure and nucleotides distribution}
-\author{Vladimir Reinharz$^1$, Yann Ponty$^{2,*}$, J\'er\^{o}me Waldisp\"{u}hl$^{1,*}$\footnote{to whom correspondence should be addressed}}
+\author{Vladimir Reinharz$^1$, Yann Ponty$^{2,*}$, J\'er\^{o}me Waldisp\"{u}hl$^{1}$\footnote{to whom correspondence should be addressed}}
 \address{$^1$ School of Computer Science, McGill University, Montreal, Canada\\ $^2$ Laboratoire d'informatique, \'Ecole Polytechnique, Palaiseau, France.}
@@
 \section{Availability:} 
-Our method is freely available at WEBSITE INCARNATION
+\ourprog is available at \href{http:// csb.cs.mcgill.ca/incarnation/}{csb.cs.mcgill.ca/incarnation/}
 \section{Contact:} \href{jeromew@cs.mcgill.ca}{jeromew@cs.mcgill.ca}, \href{yann.ponty@lix.polytechnique.fr}{yann.ponty@lix.polytechnique.fr}\\
@@
 \end{abstract}
-
-\newpage
 

### 2013-03-10 ISMB/methods_ISMB.tex
Protected biblio

@@
 
-The recursion must consider different cases, depending on whether the first position in $\Struct$ is:
-\begin{itemize}
-\item {\bf Case 1.} Unpaired ($\Struct = \ub \Struct'$):
-\begin{align}
-	\Z{\BoolTrue,\ub \Struct'}{a,b} = \Z{\BoolFalse,\ub \Struct'}{a,b}  &:=
+The general recursion scheme specializes, depending on the first position in $\Struct$.\\
+{\bf Case 1.} First position is unpaired ($\Struct = \ub \Struct'$):
+\begin{equation}
+	\Z{\BoolTrue,\ub \Struct'}{a,b} = \Z{\BoolFalse,\ub \Struct'}{a,b}  :=
       \sum_{a'\in \B}  
@@
 \label{eq:Z_unpaired}
-\end{align}
-\item {\bf Case 2.} Paired to last position ($\Struct = \op \Struct' \cp$), stacking onto a pre-existing closing pair ($\N=\BoolTrue$):
-\begin{align}
-	\Z{\BoolTrue,\op \Struct' \cp}{a,b} &:=
+\end{equation}
+{\bf Case 2.} First position is paired with last position ($\Struct = \op \Struct' \cp$), stacking onto a pre-existing exterior pair ($\N=\BoolTrue$):
+\begin{equation}
+	\Z{\BoolTrue,\op \Struct' \cp}{a,b} :=
       \sum_{a',b'\in \B^2}
@@
 \label{eq:Z_stack}
-\end{align}
-\item {\bf Case 3.} Paired to some position ($\Struct = \op \Struct' \cp \Struct''$), but not stacking ($\N=\BoolFalse$ or $\Struct''\neq \varepsilon$):
-\begin{align}
-	\Z{\N,\op \Struct' \cp \Struct''}{a,b} &:=
+\end{equation}
+{\bf Case 3.} First position is involved in a base-pair ($\Struct = \op \Struct' \cp \Struct''$), which is not stacking onto an exterior base-pair ($\N=\BoolFalse$ or $\Struct''\neq \varepsilon$):
+\begin{equation}
+	\Z{\N,\op \Struct' \cp \Struct''}{a,b} :=
 			 \displaystyle
@@
 \label{eq:Z_rec}
-\end{align}
-\end{itemize}
+\end{equation}
 
@@
 \SetKwFunction{Random}{Random}
-	rand $\leftarrow$ \Random$\left(\Z{\N,\Struct}{a,b}\right)$\tcp*[r]{Draw random number in $[0,\Z{\N,\Struct}{a,b}[$}
+\newcommand{\rand}{{r}}
+	$\rand \leftarrow $\Random$\left(\Z{\N,\Struct}{a,b}\right)$\tcp*[r]{Random real in $[0,\Z{\N,\Struct}{a,b}[$}
  \Switch{}{
-   \lCase(\tcp*[f]{Empty structure $\Rightarrow$ empty sequence}){$\Struct=\varepsilon$}{\Return{$\varepsilon$}}
+   \lCase(\tcp*[f]{Empty structure}){$\Struct=\varepsilon$}{\Return{$\varepsilon$}}
    \Case(\tcp*[f]{First position is unpaired}){$\Struct=\ub\, \Struct'$}{		
    \For{$a'\in\B$}{
-			rand $\leftarrow$ rand $- x^{\gc(a')}\cdot \Z{\BoolFalse,\Struct'}{a',b}$\;
-			\lIf{$\text{\rm rand}<0$}\Return{$a'.\Backtrack(a',b,\BoolFalse,\Struct')$}\;
+			$\rand \leftarrow \rand - x^{\gc(a')}\cdot \Z{\BoolFalse,\Struct'}{a',b}$\;
+			\lIf{$\rand<0$}\Return{$a'.\Backtrack(a',b,\BoolFalse,\Struct')$}\;
   		}
    }
-   \Case(\tcp*[f]{Extremities involved in stacking base pair}){$\Struct=\text{\rm\op}\, \Struct' \,\text{\rm\cp}$  {\rm \bf and} $\N=\BoolTrue$}
+   \Case(\tcp*[f]{Extremities are involved in stacking base pair}){$\Struct=\text{\rm\op}\, \Struct' \,\text{\rm\cp}$  {\rm \bf and} $\N=\BoolTrue$}
    {		
@@
     {
-				rand $\leftarrow$ rand $-
+				$\rand \leftarrow \rand -
 			 	x^{\gc(a'.b')}
@@
 			 	\cdot \Z{\BoolTrue,\Struct'}{a',b'}	$\;
-				\lIf{$\text{\rm rand}<0$}{
+				\lIf{$\rand<0$}{
 					\Return{$a'.\Backtrack(a',b',\BoolTrue,\Struct'). b'$}\;		
@@
 				\For{$(a',b')\in\B\times\B$}{
-					rand $\leftarrow$ rand $-	
+					$\rand \leftarrow \rand -	
        x^{\gc(a'.b')}
@@
 	    	  \cdot\Z{\BoolTrue,\Struct''}{b',b}$\;	
- 					\lIf{{\rm rand} $<0$}{
+ 					\lIf{\rand $<0$}{
 						\Return{$a'
-						.\Backtrack\left(a',b',\BoolFalse,\Struct'\right)
+						.\Backtrack\left(a',b',\BoolTrue,\Struct'\right)
 						.b'
-						.\Backtrack\left(b',b,\BoolTrue,\Struct''\right)
+						.\Backtrack\left(b',b,\BoolFalse,\Struct''\right)
 						$}\;	
@@
 
-\subsubsection{Self-adaptive sampling strategy and overall complexity}
+\subsubsection{Self-adaptive sampling strategy}
 
@@
 
+\subsubsection{Overall complexity}
+
 Regarding the overall complexity claim for our approach, it was previously established~\cite{Waldispuhl2011} that, for each value of $x$, there exists constants $\mu_x$ and $\sigma_x$ such that the distribution of \GCContent asymptotically converges towards a normal law having expectation in $\mu_x\cdot n\cdot(1+o(1))$ and standard deviation in $\sigma_x\cdot\sqrt{n}\cdot(1+o(1))$.
@@
 a constant), leading to the announced complexities.
-Formally, since a suitable weight $x$ must be recomputed for each targeted structure and \GCContent, then the number $M$ of iterations required for the converge can be accounted for explicitly, leading to time complexities in $\Theta((M+\sqrt{n})\cdot k\cdot n)$ (if $\kappa=0 \Rightarrow$ no tolerance) and $\Theta(M\cdot k\cdot n)$ (if $\kappa>0$).
+Formally, since a suitable weight $x$ must be recomputed for each targeted structure and \GCContent, then the number $M$ of iterations required for the converge can be accounted for explicitly, leading to time complexities in $\Theta((M+\sqrt{n})\cdot k\cdot n)$ (if $\kappa=0$, i.e.
+without any tolerance) and $\Theta(M\cdot k\cdot n)$ (if $\kappa>0$).
 

### 2013-03-11 ISMB/methods_ISMB.tex
workflow image+fixed biblio

@@
 
-%To that purpose, a Boltzmann weighted distribution is used, based on a pseudo-energy function $\PE{\cdot}$ which includes contributions for both the free-energy and its putative isostericity towards a multiple sequence alignment.
+%To that purpose, a Boltzmann weighted distribution is used, based on an energy function $\PE{\cdot}$ which includes contributions for both the free-energy and its putative isostericity towards a multiple sequence alignment.
 In this model, the probability that the nucleotide at a given position needs to be mutated (i.e.
-corresponds to a sequencing error) can be computed using a variant of the \emph{Inside-Outside algorithm}~\cite{Lari1990}.
+corresponds to a sequencing error) can be computed using a variant of the \emph{Inside-Outside algorithm}~\citep{Lari1990}.
 %
@@
 We use a simplified free-energy model which only includes additive contributions from stacking base-pairs.
-Using individual values from the Turner 2004 model (retrieved from the NNDB~\cite{Turner2010}).
+Using individual values from the Turner 2004 model (retrieved from the NNDB~\citep{Turner2010}).
 Given a candidate sequence $s$ for a secondary structure $\Struct$, the free-energy of any sequence $s$ of length $|\Struct|$  is given by
@@
 
-\subsubsection{\GCContent weighted pseudo-Boltzmann ensemble and distribution}
+\subsubsection{\GCWeighted Boltzmann ensemble and distribution}
 
-In order to counterbalance the documented tendency of sampling methods to generate \Gb\Cb-rich sequences~\cite{Levin:2012kx}, we introduce a parameter $x\in\mathbb{R}^+$, whose value will influence the \GCContent of generated sequences.
-For any secondary structure $\Struct$, the pseudo-Boltzmann factor of a sequence $s$ is $\B^{[x]}_{\Struct}(s)$  such that
+In order to counterbalance the documented tendency of sampling methods to generate \Gb\Cb-rich sequences~\citep{Levin:2012kx}, we introduce a parameter $x\in\mathbb{R}^+$, whose value will influence the \GCContent of generated sequences.
+For any secondary structure $\Struct$, the \GCWeighted-Boltzmann factor of a sequence $s$ is $\B^{[x]}_{\Struct}(s)$  such that
 \begin{equation}
@@
 
-Summing the pseudo-Boltzmann factor over all possible sequences of a given length $|\Struct|$, one obtains the pseudo-partition function $\mathcal{Z}_{\Struct}^{[x]}$, from which one defines the pseudo-Boltzmann probability $\Prob_{\Struct}^{[x]}(s)$ of each sequence $s$, respectively such that
+Summing the \GCWeighted-Boltzmann factor over all possible sequences of a given length $|\Struct|$, one obtains the \GCWeighted partition function $\mathcal{Z}_{\Struct}^{[x]}$, from which one defines the \GCWeighted Boltzmann probability $\Prob_{\Struct}^{[x]}(s)$ of each sequence $s$, respectively such that
 \begin{align}\mathcal{Z}_{\Struct}^{[x]} &= \sum_{\substack{|s|=n}}\B_{\Struct}^{[x]}(s)& \text{and}&&
@@
 
-\subsection{Sampling the pseudo-Boltzmann ensemble}
+\subsection{Linear-time stochastic sampling algorithm for the \GCWeighted-Boltzmann ensemble}
 
-Let us now describe a linear-time algorithm to sample sequences at random in the pseudo-Boltzmann distribution.
-This algorithm follows the general principles of the recursive approach to random generation~\cite{Wilf1977}, pioneered in the context of RNA by the \SFold algorithm~\cite{Ding2003}.
-The algorithm starts by precomputing the partition function restricted to each valid interval, and then performs a series of recursive stochastic backtracks, using precomputed values to decide on the probability of each alternative.
+Let us now describe a linear-time algorithm to sample sequences at random in the \GCWeighted Boltzmann distribution.
+This algorithm follows the general principles of the recursive approach to random generation~\citep{Wilf1977}, pioneered in the context of RNA by the \SFold algorithm~\citep{Ding2003}.
+The algorithm starts by precomputing the partition function restricted to each substructure occurring in the target structure, and then performs a series of recursive stochastic backtracks, using precomputed values to decide on the probability of each alternative.
 
-\subsubsection{Precomputing the pseudo-partition function}\label{sec:pf}
+\subsubsection{Precomputing the \GCWeighted partition function}\label{sec:pf}
 
-Firstly, a dynamic programming algorithm computes $\Z{\N,\Struct}{a,b}$ the pseudo-partition function ($x$ is omitted here for the sake of clarity) for a structure $\Struct$, assuming its (previously chosen) flanking nucleotides are $a$ and $b$ respectively, either forming a closing base-pair ($\N=\BoolTrue$) or not ($\N=\BoolFalse$).
+Firstly, a dynamic programming algorithm computes $\Z{\N,\Struct}{a,b}$ the \GCWeighted partition function ($x$ is omitted here for the sake of clarity) for a structure $\Struct$, assuming its (previously chosen) flanking nucleotides are $a$ and $b$ respectively, either forming a closing base-pair ($\N=\BoolTrue$) or not ($\N=\BoolFalse$).
 Remark that the empty structure only supports the empty sequence, having energy $0$, so one has
@@
 
-The general recursion scheme specializes, depending on the first position in $\Struct$.\\
+The general recursion scheme consists in three different terms, depending on the first position in $\Struct$:\\
 {\bf Case 1.} First position is unpaired ($\Struct = \ub \Struct'$):
@@
 	\SetAlgoLined
-\SetKwFunction{Backtrack}{SB}
+\SetKwFunction{Backtrack}{SB$_x$}
 \SetKwFunction{Random}{Random}
@@
 \subsubsection{Stochastic backtrack}
-Once the pseudo-partition functions have been computed and memorized, a stochastic backtrack starts from the target structure $\Target$ with any exterior bases $[a,b]$ and no nesting base-pair, corresponding to a call \Backtrack$\left(\varnothing,\varnothing,\BoolFalse,\Target\right)$ to Algorithm~\ref{alg:back}.
+Once the \GCWeighted partition functions have been computed and memorized, a stochastic backtrack starts from the target structure $\Target$ with any exterior bases $[a,b]$ and no nesting base-pair, corresponding to a call \Backtrack$\left(\varnothing,\varnothing,\BoolFalse,\Target\right)$ to Algorithm~\ref{alg:back}.
 At each step, a suitable assignment for one or several positions is chosen, using probabilities derived from the precomputation, as illustrated by Figure~\ref{fig:stochastic}.
@@
 $[gc-\kappa,gc+\kappa]$.
-Since sequences of arbitrary \GCContent may be generated by Algorithm~\ref{alg:back}, we use a rejection-based approach~\cite{Bodini2010}, previously adapted by the authors in a similar context~\cite{Waldispuhl2011}.
+Since sequences of arbitrary \GCContent may be generated by Algorithm~\ref{alg:back}, we use a rejection-based approach~\citep{Bodini2010}, previously adapted by the authors in a similar context~\citep{Waldispuhl2011}.
 This gives an algorithm which generates $k$ valid sequences in expected time $\Theta(k\cdot n\sqrt{n})$ when $\kappa=0$ (or $\Theta(k\cdot n)$ when $\kappa$ is a positive constant) and memory in $\Theta(k\cdot n)$.
-A complete analysis of the rejection process can be found in an earlier contribution~\cite{Waldispuhl2011}, but let us briefly outline the approach, and the main arguments used to establish its complexity.
+A complete analysis of the rejection process can be found in an earlier contribution~\citep{Waldispuhl2011}, but let us briefly outline the approach, and the main arguments used to establish its complexity.
 
-Our adaptive sampling approach simply generates sets of sequences by repeatedly running the stochastic backtrack algorithm.
-The average \GCContent induced by the current value of the $x$ parameter, can then be adequately estimated from the sample, or computed exactly using recent algorithmic advances~\cite{Ponty2011}.
+\begin{figure}
+\resizebox{\linewidth}{!}{\input{workflow.pgf}}
+\caption{General workflow of our adaptive sampling algorithm~\citep{Waldispuhl2011}.}
+\label{fig:workflow}
+\end{figure}
+
+As summarized by Figure~\ref{fig:workflow}, our adaptive sampling approach simply generates sets of sequences by repeatedly running the stochastic backtrack algorithm.
+The average \GCContent induced by the current value of the $x$ parameter, can then be adequately estimated from the sample, or computed exactly using recent algorithmic advances~\citep{Ponty2011}.
 The set of sequences is filtered to only retain valid sequences.
@@
 Consequently, for any targeted \GCContent $gc\in[0\%,100\%]$, there exists a unique value $x_{gc}$ such that generated sequences feature, on the average, the right \GCContent.
-In practice, a simple binary search~\cite{Waldispuhl2011} is used in our implementation, and typically converges after very few iterations.
-An optimal value for $x$ can also be derived analytically using interpolation after $\Theta(n)$ evaluations of $\Z{i,j}{a,b}$ for different candidate values of $x$, as previously noted~\cite{Waldispuhl2011} and implemented using the Fast-Fourier Transform~\cite{Senter2012}.
+In practice, a simple binary search~\citep{Waldispuhl2011} is used in our implementation, and typically converges after very few iterations.
+An optimal value for $x$ can also be derived analytically using interpolation after $\Theta(n)$ evaluations of $\Z{i,j}{a,b}$ for different candidate values of $x$, as previously noted~\citep{Waldispuhl2011} and could be implemented using the Fast-Fourier Transform~\citep{Senter2012}.
 
@@
 
-Regarding the overall complexity claim for our approach, it was previously established~\cite{Waldispuhl2011} that, for each value of $x$, there exists constants $\mu_x$ and $\sigma_x$ such that the distribution of \GCContent asymptotically converges towards a normal law having expectation in $\mu_x\cdot n\cdot(1+o(1))$ and standard deviation in $\sigma_x\cdot\sqrt{n}\cdot(1+o(1))$.
-Furthermore, the concentration of the distribution, asserted by its limited standard deviation, therefore the expected number of attempts required to generate a valid sequence when $\kappa=0$ (resp.
+It was previously established~\citep{Waldispuhl2011} that, for each value of $x$, there exists constants $\mu_x$ and $\sigma_x$ such that the distribution of \GCContent asymptotically converges towards a normal law having expectation in $\mu_x\cdot n\cdot(1+o(1))$ and standard deviation in $\sigma_x\cdot\sqrt{n}\cdot(1+o(1))$.
+Furthermore, the distribution of \GCContent is highly concentrated, as asserted by its limited standard deviation, therefore the expected number of attempts required to generate a valid sequence when $\kappa=0$ (resp.
 $\kappa\in\Omega(1/\sqrt n)$) grows like $\Theta(\sqrt{n})$ (resp.

### 2013-03-11 ISMB/main_ISMB.tex

@@
 \usepackage{relsize}
-\usepackage{natbib}
+%\usepackage{natbib}
 \usepackage[colorinlistoftodos]{todonotes}
@@
 \usetikzlibrary{snakes}
+\usetikzlibrary{shadows}
 \pgfdeclarelayer{background}
@@
 Most of the current software uses similar local search strategies (i.e.
-a random seed is progressively adapted to acquire the desired folding propertues) and more importantly do not allow the user to control explicitly the nucleotide distribution such as the \GCContent in their sequences.
+a random seed is progressively adapted to acquire the desired folding properties) and more importantly do not allow the user to control explicitly the nucleotide distribution such as the \GCContent in their sequences.
 However, the latter is an important criterion for large-scale applications as it could presumably be used to design sequences with better transcription rates and/or structural plasticity.

### 2013-03-11 ISMB/results_ISMB.tex
histo

@@
 \label{sec:implementation}
-Our software, \ourprog, is available at \oururl and was implemented in {\tt Python~2.7}. We used
+Our software, \ourprog, was implemented in {\tt Python~2.7}. We used
 \RNAinverse from the \textit{Vienna Package 2.0}~\cite{Hofacker:1994}.
-
 All time benchmarks were run on a single AMD Opteron(tm) 6278 Processor  at 2.4 GHz with cache of 512 KB.
@@
 Those are known secondary structure from a variety of organisms.
-We considered a subset of $50$ structures selected by Levin~\emph{et al}~\cite{Levin:2012kx}, 
+We considered a subset of $50$ structures selected by~\cite{Levin:2012kx}, 
 whose length ranges between $20$ and $100$ nucleotides. 

### 2013-03-11 ISMB/supplementary_ISMB.tex
bioinfo template
Typo
hist

@@
 Using the same dataset of $50$ structures, we generated $100$ samples
-per structure with \RNAinverse. They yield for most parameters
-a high entropy.
-Table.~\ref{tab:nb_rnainv} contains the number of samples with a given \GCContent generated by \RNAinverse. The vast majority resides around $50\%-70\%$, showing one of its major deficiencies. 
+per structure %with \RNAinverse. They yield for most parameters
+%a high entropy.
+%Table.~\ref{tab:nb_rnainv} contains the number of samples with a given \GCContent generated
+with  \RNAinverse, \INFORNA, \NUPACK and \frankenstein.
+They have a highly concentrated distribution of \GC content showing one of theirs major deficiencies.
+We present in Fig.~\ref{fig:gcdist} those distributions.
+
+\begin{figure}[ht]
+  \begin{center}
+    \includegraphics[width=0.5\textwidth]{Figures/histograme_5_gc_distribution_nornaexinv.png}
+  \end{center}
+  \caption{Overall \GCContent distribution for sequences designed using \RNAinverse, \INFORNA, \NUPACK and \frankenstein folding in the desired structure.}
+  \label{fig:gcdist}
+\end{figure}
+
  For all structures that have been solved 
@@
 
-\begin{table}[h!]
-	\begin{center}
-		\begin{tabular}{|c|ccccc|}
-		\hline
-		Target \GCContent & 10 & 30 & 50 & 70 & 90\\ \hline
-   $\#$\RNAinverse samples& 11 & 437 & 3362 & 1155 & 35\\ \hline
-		\end{tabular}
-	\end{center}
-  \caption{Overall \GCContent distribution for sequences designed using \RNAinverse.}
-	\label{tab:nb_rnainv}
-\end{table}
+%\begin{table}[h!]
+%	\begin{center}
+%		\begin{tabular}{|c|ccccc|}
+%		\hline
+%		Target \GCContent & 10 & 30 & 50 & 70 & 90\\ \hline
+%   $\#$\RNAinverse samples& 11 & 437 & 3362 & 1155 & 35\\ \hline
+%		\end{tabular}
+%	\end{center}
+%  \caption{Overall \GCContent distribution for sequences designed using \RNAinverse.}
+%	\label{tab:nb_rnainv}
+%\end{table}
 
-\begin{figure}[ht!]
+\begin{figure*}[ht!]
 	\centering
 	%\hspace{-5em}
-	\includegraphics[scale=0.4]{Figures/RNAinverse_data_100.png}
+	\includegraphics[width=0.5\textwidth]{Figures/RNAinverse_data_100.png}
 	\caption{Entropy and \GCContent  for structures solved by
@@
 	\label{fig:rnainverse}
-\end{figure}
+\end{figure*}
 {\em Récupérer Base-pair entropy}
@@
 \begin{center}
+\resizebox{0.5\textwidth}{!}{
 \begin{tabular}{|c|c|c|}
 \hline
-\multirow{3}{*}{Target \GCContent (\%)}& \multicolumn{2}{|c|}{Designed sequences \GCContent (\%)}\\ \cline{2-3}
+\multirow{3}{*}{Target \GCContent (\%)}& \multicolumn{2}{c|}{\GCContent (\%) of designed sequences}\\ \cline{2-3}
  & \ourprog & \ourprog + \RNAinverse\\
@@
 \end{tabular}
+}
 \end{center}

### 2013-03-11 ISMB/results_ISMB.tex
eq

@@
  To benchmark our method, we first started by sampling about $100$ sequences per structure. After correcting the free nucleotides with
- \texttt{RNAinverse} we computed their MFE with \texttt{RNAfold} from the \textit{Vienna Package 2.0}~\cite{Hofacker:1994}.
+ \RNAinverse we computed their MFE with \rnafold from the \textit{Vienna Package 2.0} by~\cite{Hofacker:1994}.
 
@@
 A desirable feature in sequence design, is to produce samples with a high
-diversity. Figure~\ref{fig:nb_sols_entropy} shows the number of correct
+diversity but with a conserved structure.
+Two useful measures are the entropy~\ref{eq:entropy} and the base pair entropy~\ref{eq:bpentropy}.
+The entropy is defined as:
+\begin{equation}
+  \sum_{\substack{i\\b \in (A,C,G,U)}} -p^b_i\log_4 p^b_i
+  \label{eq:entropy}
+\end{equation}
+where $p^b_i$ is the proportion of nucleotides $b$ at position $i$.
+The base pair entropy is defined as:
+\begin{equation}
+  \sum_{i<j}-p_{i,j}\log_2 p_{i,j}
+  \label{eq:bpentropy}
+\end{equation}
+where $p_{i,j}$ is the probability of having a base pair between positions $i$ and $j$. The base pair entropy reflects
+the distribution of structures weights in the Boltzmann ensemble. Thus, a low base pair entropy implies a highly conserved structure.
+To compute the base pairs probabilities, we used \rnafold from the \texttt{Vienna Package 2.0} developed by~\cite{Hofacker:1994}.
+
+  Figure~\ref{fig:nb_sols_entropy} shows the number of correct
 solutions as their average entropy and base pair entropy, since 

### 2013-03-11 ISMB/methods_ISMB.tex
Added highlight macro and fixed response to reviewers (still to be completed)

@@
 
+\highlight{
 \subsubsection{Precomputing the \GCWeighted partition function}\label{sec:pf}
 
-Firstly, a dynamic programming algorithm computes $\Z{\N,\Struct}{a,b}$ the \GCWeighted partition function ($x$ is omitted here for the sake of clarity) for a structure $\Struct$, assuming its (previously chosen) flanking nucleotides are $a$ and $b$ respectively, either forming a closing base-pair ($\N=\BoolTrue$) or not ($\N=\BoolFalse$).
+Firstly, a dynamic programming algorithm computes $\Z{\N,\Struct}{a,b}$ the \GCWeighted partition function (the dependency in $x$ is omitted here for the sake of clarity) for a structure $\Struct$, assuming its (previously chosen) flanking nucleotides are $a$ and $b$ respectively, either forming a closing base-pair ($\N=\BoolTrue$) or not ($\N=\BoolFalse$).
 Remark that the empty structure only supports the empty sequence, having energy $0$, so one has
@@
 Such values are entirely determined by $\Struct$ at any given step of the recursion, and their dependency can be summarized in a tree having $\Theta(|\Struct|)$.
-Therefore, the computation of $\Z{\N,\Target}{a,b}$ requires $\Theta(n)$ time and space using dynamic-programming.
+Therefore, the computation of $\Z{\N,\Target}{a,b}$ requires $\Theta(n)$ time and space using dynamic-programming.}
 
@@
 
+\begin{figure}
+\resizebox{\linewidth}{!}{\input{workflow.pgf}}
+\caption{General workflow of our adaptive sampling algorithm~\citep{Waldispuhl2011}.}
+\label{fig:workflow}
+\end{figure}
+
+\highlight{
 \subsubsection{Self-adaptive sampling strategy}
@@
 A complete analysis of the rejection process can be found in an earlier contribution~\citep{Waldispuhl2011}, but let us briefly outline the approach, and the main arguments used to establish its complexity.
-
-\begin{figure}
-\resizebox{\linewidth}{!}{\input{workflow.pgf}}
-\caption{General workflow of our adaptive sampling algorithm~\citep{Waldispuhl2011}.}
-\label{fig:workflow}
-\end{figure}
 
@@
 without any tolerance) and $\Theta(M\cdot k\cdot n)$ (if $\kappa>0$).
+}
 

### 2013-03-11 ISMB/ResponseToReviewers-Incarnation-ISMB-2013.tex

--- 2013-03-09 ISMB/ResponseToReviewers-Incarnation-ISMB-2013.tex
@@
 \usepackage{xspace}
+\usepackage{url}
 \usepackage{amssymb} % for much better looking tables
@@
 \Comment{Is the method available?}
-\Answer{TODO (Jerome)}
+\Answer{The method is available for download (and, in the near future, as a web server) at \oururl, as indicated in our revised version.}
 
@@
 \Comment{1) The $x^{\gc(s)}$ in equation (1) seems to me a novel contributing term that is of high importance. Perhaps its introduction should be highlighted more in the text.}
-\Answer{We tried to make this point clearer in our revised version.
-However, it should be noted that this contribution is not a goal in itself, but rather a mean to perform stochastic sampling with respect to the weighted distribution.
-This distribution has all sorts of desirable mathematical properties, among which the fact that induced distributions for additive features (here the \GCContent) are nicely concentrated, allowing for a rejection-based approach.}
+\Answer{We tried to emphasize this aspect in our revised version.
+However, it should be noted that this term is mainly a mean to perform stochastic sampling with respect to the weighted distribution, on which we prefer to put the emphasis.
+This distribution has all sorts of desirable mathematical properties, among which the fact that induced distributions for additive features (here the \GCContent) are nicely concentrated, allowing for a rejection-based approach.
+}
 
@@
 While such a postprocessing tends to reestablish \GCContent{}s that are  closer to 50\% \GC, this shift remains reasonable (at most ~6\% \GC).
-Furthermore, this shift is very predictable and offers the user the possibility to \emph{overshoot} for a more extreme \GCContent with the global approach, knowing that the postprocessing step will tend to average it.}
+This shift is very predictable and offers the user the possibility to \emph{overshoot} for a more extreme \GCContent with the global approach, knowing that the postprocessing step will tend to average it.
+Our interpretation of this limited shift, is that \ourprog brings \RNAinverse in a region which is likely to satisfy the  MFE criterion,  and thus will require only minimal adjustments compared to the ones required for a random initial sequence.}
 
@@
 However, adopting a pure tree-like representation would require elaborating on the correspondence between trees and structures, and might lead to technicalities in the context of stacking base-pairs.
-Therefore, we completely rewrote our equations, figure and algorithm to adopt a classic dot-parenthesis representation.
-While the end result may be less elegant than expected with a tree-like presentation, this rewrite has the merit of dropping the double subscripts, hopefully better conveying the  intuition underlying the complexity claim in our original manuscript.}
+Therefore, we completely rewrote our equations, figure and algorithm, adopting a classic dot-parenthesis representation.
+While the end result may be less elegant than could be expected with a tree-like presentation, this rewrite has the merit of dropping the double subscripts, hopefully better conveying the  intuition underlying the complexity claim in our original manuscript.}
 
 \Comment{(3) Figure 1 is also a bit misleading. Make it clear that the stochastic choice is NOT made over the three cases depicted. The case is always fixed for given $i$, just the bases are chosen.}
-\Answer{Absolutely.
+\Answer{Absolutely agreed.
 We changed this figure, which now explicitly distinguishes between the three different cases (different left-hand side diagrams), and also features the base emission probabilities.}
@@
 \Comment{(4) Section 2.2.3 is titled "Rejecting unsuitable candidates". It explains that candidates tend to be quite suitable, but somewhere, you should actually tell who is rejected and when ... }
-\Answer{We completely rewrote this section, which now starts by explicitly reminding the problem addressed by the algorithm.}
+\Answer{We completely rewrote this section, which now starts by explicitly (re)stating the problem addressed by the algorithm.}
 

### 2013-03-11 ISMB/results_ISMB.tex

@@
 Our software, \ourprog, was implemented in {\tt Python~2.7}. We used
-\RNAinverse from the \textit{Vienna Package 2.0}~\cite{Hofacker:1994}.
+\RNAinverse from the \textit{Vienna Package 2.0}~\citep{Hofacker:1994}.
 All time benchmarks were run on a single AMD Opteron(tm) 6278 Processor  at 2.4 GHz with cache of 512 KB.
@@
 To evaluate the quality of our method, we used secondary
-structures from the \RNASTRAND database~\cite{andronescu2008rna}.
+structures from the \RNASTRAND database~\citep{andronescu2008rna}.
 Those are known secondary structure from a variety of organisms.
@@
 diversity but with a conserved structure.
-Two useful measures are the entropy~\ref{eq:entropy} and the base pair entropy~\ref{eq:bpentropy}.
-The entropy is defined as:
-\begin{equation}
-  \sum_{\substack{i\\b \in (A,C,G,U)}} -p^b_i\log_4 p^b_i
+Two useful measures are the positional entropy and the base pair entropy.
+The positional entropy is defined as:
+\begin{align}
+  \sum_{\substack{i\\b \in (\Ab,\Cb,\Gb,\Ub)}} -p_{i\to b}\cdot \log_4(p_{i\to b}) &&& \text{(Positional entropy)}
   \label{eq:entropy}
-\end{equation}
-where $p^b_i$ is the proportion of nucleotides $b$ at position $i$.
+\end{align}
+where $p_{i\to b}$ is the proportion of nucleotides $b$ at position $i$.
+Intuitively, this measure captures the variability of sequences generated by a given method.
 The base pair entropy is defined as:
-\begin{equation}
-  \sum_{i<j}-p_{i,j}\log_2 p_{i,j}
+\begin{align}
+  \sum_{i<j}-p_{i,j}\cdot \log_2(p_{i,j})&&& \text{(Base-Pair entropy)}
   \label{eq:bpentropy}
-\end{equation}
-where $p_{i,j}$ is the probability of having a base pair between positions $i$ and $j$. The base pair entropy reflects
+\end{align}
+where $p_{i,j}$ is the Boltzman probability, i.e. probability that a random structure drawn with respect to 
+the Boltzmann distribution forms a base-pair between positions $i$ and $j$. The base pair entropy reflects
 the distribution of structures weights in the Boltzmann ensemble. Thus, a low base pair entropy implies a highly conserved structure.

### 2013-03-12 ISMB/results_ISMB.tex
more

@@
 
-\begin{figure*}[t!]
+\begin{figure}[ht!]
 	\centering
- 	\includegraphics[width=\textwidth]{Figures/time_rnastrand_clustered_rnainverse_100samples_fix}
-	\caption{Average time in seconds to generate one sequence. We explicitly show 
-	the time spent by \ourprog (full line) and \texttt{RNAinverse} (dashed) for various \GCContent. The first plot is in function
-	of the length of the structures. The second is in function of the stack
-	density (i.e. $2\cdot\#stacks/length$) and the latter in function of 
-	the proportion of free nucleotides (i.e. unpaired bases) within loops.}
+ 	\includegraphics[width=0.5\textwidth]{Figures/time_incar_rnassd}
+	\caption{
+  Average time in seconds to generate one sequence for \ourprog and \RNAinverse.}
+	%the time spent by \ourprog (full line) and \texttt{RNAinverse} (dashed) for various \GCContent. The first plot is in function
+	%of the length of the structures. The second is in function of the stack
+	%density (i.e. $2\cdot\#stacks/length$) and the latter in function of 
+	%the proportion of free nucleotides (i.e. unpaired bases) within loops.}
 	\label{fig:time}	
-\end{figure*}
+\end{figure}
 
@@
 	\centering
-	\includegraphics[width=\textwidth]{Figures/mfe_struct_solve_nornainverse.png}
+	\includegraphics[width=\textwidth]{Figures/mfe_struct_solved_vsrnainverse}
 	\caption{Quality of \ourprog results. The first row shows the percentage
-	of sampled sequences folding into the target. The second shows the 	
-	proportion	of structures for which at least one correct sequence was 
-	sampled.}
+	of sampled sequences folding into the target when using only \ourprog. The second shows 
+  after processing previous results with \RNAinverse.
+	}
 	\label{fig:mfe_struct_solved_noinverse}	
@@
 
-\begin{figure*}[ht!]	
-	\centering
-	\includegraphics[width=\textwidth]{Figures/mfe_struct_solved}
-	\caption{Quality of \texttt{Incarnation+RNAinverse} results. The first row shows the percentage
-	of sampled sequences folding into the target. The second shows the 	
-	proportion	of structures for which at least one correct sequence was 
-	sampled.}
-	\label{fig:mfe_struct_solved}	
-\end{figure*}
+%\begin{figure*}[ht!]	
+%	\centering
+%	\includegraphics[width=\textwidth]{Figures/mfe_struct_solved}
+%	\caption{Quality of \texttt{Incarnation+RNAinverse} results. The first row shows the percentage
+%	of sampled sequences folding into the target. The second shows the 	
+%	proportion	of structures for which at least one correct sequence was 
+%	sampled.}
+%	\label{fig:mfe_struct_solved}	
+%\end{figure*}
 
@@
 
+Using the same dataset of $50$ structures, we generated $100$ samples
+per structure with classical softwares who can not control the \GC content.%with \RNAinverse. They yield for most parameters
+%a high entropy.
+%Table.~\ref{tab:nb_rnainv} contains the number of samples with a given \GCContent generated
+with  \RNAinverse, \INFORNA, \NUPACK and \frankenstein.
+We present in Fig.~\ref{fig:gcdist} those distributions.
+They have a highly concentrated distribution of \GC content showing one of theirs major deficiencies, with no sequence below $30\%$\GC.
+Therefore the need to be able to explicitly
+ control the \GC content.
+
+\begin{figure}[ht!]
+  \begin{center}
+    \includegraphics[width=0.5\textwidth]{Figures/histograme_5_gc_distribution_nornaexinv.png}
+  \end{center}
+  \caption{Overall \GCContent distribution for sequences designed using \RNAinverse, \INFORNA, \NUPACK and \frankenstein folding in the desired structure.}
+  \label{fig:gcdist}
+\end{figure}
+
 A desirable feature in sequence design, is to produce samples with a high
-diversity but with a conserved structure.
-Two useful measures are the positional entropy and the base pair entropy.
+diversity.
+Two useful measures are the positional entropy and the sequence identity. %base pair entropy.
 The positional entropy is defined as:
@@
 Intuitively, this measure captures the variability of sequences generated by a given method.
-The base pair entropy is defined as:
+The positional entropy is defined over a set $\mathcal{S}$ of sequences of equal length as follows:
 \begin{align}
-  \sum_{i<j}-p_{i,j}\cdot \log_2(p_{i,j})&&& \text{(Base-Pair entropy)}
-  \label{eq:bpentropy}
+  \sum_{s^1,s^2 \in\mathcal{S}\times\mathcal{S}}\left(\frac{1}{|s^1|}{\sum_{\substack{i\\s^1_i\equiv s^2_i}}1}\right) &&& \text{Seq. identity}
 \end{align}
-where $p_{i,j}$ is the Boltzman probability, i.e. probability that a random structure drawn with respect to 
-the Boltzmann distribution forms a base-pair between positions $i$ and $j$. The base pair entropy reflects
-the distribution of structures weights in the Boltzmann ensemble. Thus, a low base pair entropy implies a highly conserved structure.
-To compute the base pairs probabilities, we used \rnafold from the \texttt{Vienna Package 2.0} developed by~\cite{Hofacker:1994}.
+where $s_i$ is the nucleotide at position $i$ in sequence $s$. The sequence identity is the ratio of perfectly conserved positions over
+all pairwise sequences in the set. It reflects how over all the sequences the same positions are conserved. 
+%The base pair entropy is defined as:
+%\begin{align}
+%  \sum_{i<j}-p_{i,j}\cdot \log_2(p_{i,j})&&& \text{(Base-Pair entropy)}
+%  \label{eq:bpentropy}
+%\end{align}
+%where $p_{i,j}$ is the Boltzman probability, i.e. probability that a random structure drawn with respect to 
+%the Boltzmann distribution forms a base-pair between positions $i$ and $j$. The base pair entropy reflects
+%the distribution of structures weights in the Boltzmann ensemble. Thus, a low base pair entropy implies a highly conserved structure.
+%To compute the base pairs probabilities, we used \rnafold from the \texttt{Vienna Package 2.0} developed by~\cite{Hofacker:1994}.
 
   Figure~\ref{fig:nb_sols_entropy} shows the number of correct
-solutions as their average entropy and base pair entropy, since 
+solutions as their average entropy and positions inside base pairs entropy, since 
 \ourprog  only influences the distribution of nucleotides inside 
@@
   \centering
-	\includegraphics[width=\textwidth]{Figures/nb_sols_entropy.png}
+  \includegraphics[width=\textwidth]{Figures/seq_identity_10_30}
+  \caption{Sequence identity of \ourprog and \RNAinverse for $10$ and $30\%$ of \GC.}
+\end{figure*}
+
+\begin{figure*}[ht!]
+	\centering
+	\includegraphics[width=\textwidth]{Figures/nb_sols_entropy}
 	\caption{Number of solutions generated and their average entropy. 
@@
 
-It should be noted, however, than \RNAinverse produces sequences whose \GCContent is extremely concentrated (see Table~\ref{tab:nb_rnainv} in Supplementary Material), thereby requiring a large number of runs before producing a suitable \GCContent sequence.
+It should be noted, however, than \RNAinverse produces sequences whose \GCContent is extremely concentrated (see Fig.), thereby requiring a large number of runs before producing a suitable \GCContent sequence.
 Therefore, a rejection algorithm based on \RNAinverse alone should not be considered a realistic solution for the design with constrained \GCContent.

### 2013-03-12 ISMB/supplementary_ISMB.tex
more

@@
 \section{Supplementary data}
-\subsection{Benchmark \ourprog vs \RNAinverse}
-Using the same dataset of $50$ structures, we generated $100$ samples
-per structure %with \RNAinverse. They yield for most parameters
+
+\subsection{Benchmark \ourprog+\RNAinverse}
+To emphasize the usefulness of processing \ourprog sequences with \RNAinverse, we present the number
+of structures for which at least one sequence was generated with the desired MFE in Figure.~\ref{fig:nb_sols}
+
+\begin{figure*}[ht!]
+  \centering
+  \includegraphics[width=\textwidth]{Figures/struct_solved_vsrnainverse.png}
+  \caption{The first row shows the number of structures for which one generated sequence has the 
+  structure as MFE when only using \ourprog. The second row shows when we process \ourprog results with \RNAinverse.}
+  \label{fig:nb_sols}
+\end{figure*}
+
+\subsection{Benchmark \ourprog vs \RNASSD}
+%Using the same dataset of $50$ structures, we generated $100$ samples
+%per structure %with \RNAinverse. They yield for most parameters
 %a high entropy.
 %Table.~\ref{tab:nb_rnainv} contains the number of samples with a given \GCContent generated
-with  \RNAinverse, \INFORNA, \NUPACK and \frankenstein.
+%with  \RNAinverse, \INFORNA, \NUPACK and \frankenstein.
 They have a highly concentrated distribution of \GC content showing one of theirs major deficiencies.
@@
 
-\begin{figure}[ht]
-  \begin{center}
-    \includegraphics[width=0.5\textwidth]{Figures/histograme_5_gc_distribution_nornaexinv.png}
-  \end{center}
-  \caption{Overall \GCContent distribution for sequences designed using \RNAinverse, \INFORNA, \NUPACK and \frankenstein folding in the desired structure.}
-  \label{fig:gcdist}
-\end{figure}
+%\begin{figure}[ht]
+%  \begin{center}
+%    \includegraphics[width=0.5\textwidth]{Figures/histograme_5_gc_distribution_nornaexinv.png}
+%  \end{center}
+%  \caption{Overall \GCContent distribution for sequences designed using \RNAinverse, \INFORNA, \NUPACK and \frankenstein folding in the desired structure.}
+%  \label{fig:gcdist}
+%\end{figure}
 
  For all structures that have been solved 
-by the three methods, only \RNAinverse, only \texttt{Incarnation} and
+by the three methods, only \RNASSD, only \texttt{Incarnation} and
 \texttt{Incarnation} followed by \RNAinverse, 
-we present for every different concentration of \GCContent
-the average sample entropy and the average base pairs entropy in Figure.~\ref{fig:rnainverse}.
+we present for different concentration of \GCContent the average sequence identity in Figure.~\ref{fig:identity_50_70_90}
+%the average sample entropy and the average base pairs entropy in Figure.~\ref{fig:rnainverse}.
 
+\begin{figure*}[ht!]
+  \centering
+  \includegraphics[width=\textwidth]{Figures/seq_identity_50_70_90.png}
+  \caption{Sequence identity by average \GCContent of $50$, $70$ and $90\%$ for \ourprog,\ourprog+\RNAinverse and \RNASSD.}
+  \label{fig:identity_50_70_90}
+\end{figure*}
 %\begin{table}[h!]
@@
 
-\begin{figure*}[ht!]
-	\centering
+%\begin{figure*}[ht!]
+%	\centering
 	%\hspace{-5em}
-	\includegraphics[width=0.5\textwidth]{Figures/RNAinverse_data_100.png}
-	\caption{Entropy and \GCContent  for structures solved by
-	the 3 methods.}
-	\label{fig:rnainverse}
-\end{figure*}
-{\em Récupérer Base-pair entropy}
+%	\includegraphics[width=0.5\textwidth]{Figures/RNAinverse_data_100.png}
+%	\caption{Entropy and \GCContent  for structures solved by
+%	the 3 methods.}
+%	\label{fig:rnainverse}
+%\end{figure*}
+%{\em Récupérer Base-pair entropy}
 
@@
 %\end{figure}
+
 \subsection{Limited impact on \GC of local-search postprocessing of \ourprog output}

### 2013-03-12 ISMB/results_ISMB.tex
graph better and freq. equation

@@
 A desirable feature in sequence design, is to produce samples with a high
-diversity.
-Two useful measures are the positional entropy and the sequence identity. %base pair entropy.
-The positional entropy is defined as:
-\begin{align}
-  \sum_{\substack{i\\b \in (\Ab,\Cb,\Gb,\Ub)}} -p_{i\to b}\cdot \log_4(p_{i\to b}) &&& \text{(Positional entropy)}
-  \label{eq:entropy}
-\end{align}
-where $p_{i\to b}$ is the proportion of nucleotides $b$ at position $i$.
-Intuitively, this measure captures the variability of sequences generated by a given method.
-The positional entropy is defined over a set $\mathcal{S}$ of sequences of equal length as follows:
+diversity and stable secondary structure. 
+Two useful measures are the sequence identity Boltzmann frequency.
+%The positional entropy is defined as:
+%\begin{align}
+%  \sum_{\substack{i\\b \in (\Ab,\Cb,\Gb,\Ub)}} -p_{i\to b}\cdot \log_4(p_{i\to b}) &&& \text{(Positional entropy)}
+%  \label{eq:entropy}
+%\end{align}
+%where $p_{i\to b}$ is the proportion of nucleotides $b$ at position $i$.
+%Intuitively, this measure captures the variability of sequences generated by a given method.
+The sequence identity is defined over a set $\mathcal{S}$ of sequences of equal length as follows:
 \begin{align}
@@
 \end{align}
-where $s_i$ is the nucleotide at position $i$ in sequence $s$. The sequence identity is the ratio of perfectly conserved positions over
-all pairwise sequences in the set. It reflects how over all the sequences the same positions are conserved. 
+where $s_i$ is the nucleotide at position $i$ in sequence $s$.
+Intuitively, this measure captures the variability of sequences generated by a given method.
+The Boltzmann frequency is defined, for a structure $\Struct$ and a sequence $s$ as:
+\begin{align}
+  e^{\frac{-\ES(s,\Struct)}{RT}}/\mathcal{Z}^s  &&&\text{Frequency}
+\end{align}
+where $\mathcal{Z}^s$ is the partition function of sequence $s$. This measure tells us how dominant is structure $\Struct$ in the Boltzmann ensemble of sequence $s$. A high value implies a 
+stable structure. To compute the frequency, we used \rnafold from the \texttt{Vienna Package 2.0} developed by~\cite{Hofacker:1994}.
+
 %The base pair entropy is defined as:

### 2013-03-12 ISMB/supplementary_ISMB.tex
graph better and freq. equation

@@
 \section{Supplementary data}
+
+\subsection{Benchmark other softwares}
+To evaluate the performances of \ourprog, we benchmark a set of classical softwares lacking \GCContent control.
+Those are \RNAinverse, \INFORNA, \NUPACK and \frankenstein.
+We present in Fig.~\ref{fig:no_gc_control} the average sequence identity and frequency
+for sequences generated them.
+
+\begin{figure*}[ht!]
+  \centering
+  \includegraphics[width=\textwidth]{Figures/entropy_nogc}
+  \caption{The average sequence identity and frequency for softwares without \GCContent control.}
+  \label{fig:no_gc_control}
+\end{figure*}
 
@@
 %the average sample entropy and the average base pairs entropy in Figure.~\ref{fig:rnainverse}.
-
 \begin{figure*}[ht!]
@@
 
-\begin{table}[h!]
+\begin{table}[ht!]
 \begin{center}

### 2013-03-12 ISMB/results_ISMB.tex
intermediate commit (results)

@@
 To ease the visualization of results, we clustered together structures
- having similar length, stacks density and proportion of free nucleotides in loops, leading to distributions of structures shown in Figure~\ref{fig:bins}.
+having similar length, stacks density and proportion of free nucleotides in loops,
+leading to distributions of structures shown in Figure~\ref{fig:bins}.
 
@@
 
+ \highlight{
 \subsection{Design}
- To benchmark our method, we first started by sampling about $100$ sequences per structure. After correcting the free nucleotides with
- \RNAinverse we computed their MFE with \rnafold from the \textit{Vienna Package 2.0} by~\cite{Hofacker:1994}.
+We ran our method as follows. First, we sampled approximately $100$ sequences per structure. Then, we use these sequences 
+as seed in \RNAinverse. Finally, we computed the MFE with the \rnafold program from the \textit{Vienna Package 2.0} \citep{Hofacker:1994}.
 
-A first criterion is the proportion of sequences accepting the target
-secondary structures as their MFE. A second criterion is the number of structures
-for which at least one sequence with the desired MFE was produced.
-Figure~\ref{fig:mfe_struct_solved_noinverse} shows the results obtained by \ourprog for these criteria. As mentioned earlier, the high rate of failure
-is due to the fact that no selection criterion is applied to
-unpaired nucleotides. A local strategy is thus needed.
-After processing \ourprog sequences with \RNAinverse to 
-locally optimize the free nucleotides, we obtained the results 
-summarized by Figure~\ref{fig:mfe_struct_solved}. We observed
-that length is, in general, not a good predictor for the hardness of designing a structure. 
-Instead, a high number of free nucleotides in the structure seems to be a 
-good measure of its design hardness. 
- The difficulty of 
-designing sequences for targets with a high number of free nucleotides 
- in loops appears clearly in the last column of Figure~\ref{fig:mfe_struct_solved}.
-Even with a \GCContent of $50\%$ and more, when at least
-one solution was found for almost all structures, most of the samples 
-do not fold into the target. A few structures remained unsolved under
-all \GCContent.
+\subsection{Success rate}
+We started by estimating the success rate of our methodology and computed the percentage of sequences with a MFE structure identical to the target secondary structure.
+Figure~\ref{fig:mfe_struct_solved_noinverse} shows our results.
+We clearly see that before the post-processing step (i.e.
+\RNAinverse) the sequences sampled by \ourprog have a low success rate (first row).
+As mentioned earlier, this could be explained by the fact that no selection criterion has been at this stage applied to unpaired nucleotides.
+Remarkably, after the local search optimization (with \RNAinverse) of nucleotides in unpaired regions (second row), we observe a dramatic improvement of our success rate.
+As expected, we observed that length is, in general, not a good predictor for the hardness of designing a structure.
+Instead, a high number of free nucleotides in the structure seems to be a  good measure of the hardness of its design.
+Similarly, these data also show that designing sequences with low \GCContent is challenging for all types of targets.
+}
 
@@
 	\includegraphics[width=\textwidth]{Figures/mfe_struct_solved_vsrnainverse}
-	\caption{Quality of \ourprog results. The first row shows the percentage
-	of sampled sequences folding into the target when using only \ourprog. The second shows 
-  after processing previous results with \RNAinverse.
-	}
+	\caption{Success rate \ourprog before after after \RNAinverse post-processing.
+The first row shows the percentage of sampled sequences folding into the target when using only \ourprog.
+The second shows after processing previous results with \RNAinverse.}
 	\label{fig:mfe_struct_solved_noinverse}	
@@
 
-To evaluate the global quality of \ourprog sequences, we show
-in Figure~\ref{fig:ss_sens} the ratio of well predicted base pairs in the
-MFE structure of our sampled sequences. We can observe that, in all cases,
-the hardest sequence to design are those with an extremely low \GCContent. As anticipated, those are the sequences with the weakest bonds.
-We notice that the most accurate sequences yield a \GCContent
-of $70\pm 10\%$.
+\highlight{ 
+We investigated further the quality of the sequences generated by \ourprog.
+In particular, we estimated the capacity of our methods to generate ``good'' sequences with desired folding capabilities regardless of the property  to fold exactly into the target structure.
+In Figure~\ref{fig:ss_sens}, we show the ratio of well predicted base pairs in the MFE structure of our sampled sequences.
+As above, we can observe that, in all cases, the sequences that are the hardest to design are those with an extremely low \GCContent.
+Indeed, the energetic contribution of the base pairs to the stability of the structure is weaker.
+Interestingly, we also notice that the most accurate sequences yield a \GCContent of $70\pm 10\%$.
+Overall, we observe that all our samples have good folding properties, and that there is a correlation between the ``precision'' of the samples and the hardness of the design.
+}
 
-As discussed in Sec.~\ref{sec:implementation}, we notice a highly decreased
-computational time needed to generate the sequences with $15\%$ free 
-nucleotides in the loops. We remark that those structures also yield 
-a much lower structural sensitivity.
+As discussed in Sec.~\ref{sec:implementation}, we noticed a highly decreased computational time needed to generate the sequences with $15\%$ free 
+nucleotides in the loops. We remark that those structures also yield a much lower structural sensitivity.
 
@@
 
-\subsection{Samples properties}
+\subsection{Properties of designed sequences}
 
-In this section, we further analyze the generated sequences that fold into the 
-target structure.
+In this section, we further analyze the generated sequences with a MFE structure that folds into the target structure.
 
-Using the same dataset of $50$ structures, we generated $100$ samples
-per structure with classical softwares who can not control the \GC content.%with \RNAinverse. They yield for most parameters
-%a high entropy.
-%Table.~\ref{tab:nb_rnainv} contains the number of samples with a given \GCContent generated
-with  \RNAinverse, \INFORNA, \NUPACK and \frankenstein.
-We present in Fig.~\ref{fig:gcdist} those distributions.
-They have a highly concentrated distribution of \GC content showing one of theirs major deficiencies, with no sequence below $30\%$\GC.
-Therefore the need to be able to explicitly
- control the \GC content.
+\highlight{
+Using the same dataset of $50$ structures, we generated $100$ samples per structure with classical softwares who do not control the \GCContent.
+Namely,  \RNAinverse, \INFORNA, \NUPACK and \frankenstein.
+We show the distribution of the \GCContent of the sequences produced with these softwares in Fig.~\ref{fig:gcdist} those distributions.
+As anticipated, we observe a clear bias toward high {\GCContent}s and a complete absence of sequence with less than $30\%$ of \GC.
+This striking results motivates a need for methods that enable to explicitly control the \GCContent and more precisely that enable to design sequences with low \GCContent (i.e.
+30\% or less).
+}
 
@@
 
-A desirable feature in sequence design, is to produce samples with a high
-diversity and stable secondary structure. 
-Two useful measures are the sequence identity Boltzmann frequency.
+\highlight{
+A desirable feature in sequence design, is to produce samples with a high sequence diversity and stable secondary structure. 
+Therefore, in the following we will use two useful measures which are the sequence identity of the samples, and the Boltzmann probability
+of the target structure in the low energy ensemble.
+}
 %The positional entropy is defined as:
@@
 %Intuitively, this measure captures the variability of sequences generated by a given method.
-The sequence identity is defined over a set $\mathcal{S}$ of sequences of equal length as follows:
+\highlight{
+The sequence identity is defined over a set $\mathcal{S}$ of aligned sequences (in our case, all sequences have the same length and can be trivially aligned) as :
 \begin{align}
@@
 where $s_i$ is the nucleotide at position $i$ in sequence $s$.
-Intuitively, this measure captures the variability of sequences generated by a given method.
-The Boltzmann frequency is defined, for a structure $\Struct$ and a sequence $s$ as:
+Intuitively, this measure captures the diversity of sequences generated by a given method.
+Next, the Boltzmann frequency is defined, for a structure $\Struct$ and a sequence $s$ as:
 \begin{align}
@@
 \end{align}
-where $\mathcal{Z}^s$ is the partition function of sequence $s$. This measure tells us how dominant is structure $\Struct$ in the Boltzmann ensemble of sequence $s$. A high value implies a 
-stable structure. To compute the frequency, we used \rnafold from the \texttt{Vienna Package 2.0} developed by~\cite{Hofacker:1994}.
+where $\mathcal{Z}^s$ is the partition function of sequence $s$.
+This measure tells us how dominant is a structure $\Struct$ in the Boltzmann ensemble of structure over a sequence $s$.
+A high value implies a stable structure.
+We compute this frequency with \rnafold from the \texttt{Vienna Package 2.0} \citep{Hofacker:1994}.
+}
 
@@
 	\centering
-  \includegraphics[width=\textwidth]{Figures/seq_identity_10_30}
-  \caption{Sequence identity of \ourprog and \RNAinverse for $10$ and $30\%$ of \GC.}
-\end{figure*}
-
-\begin{figure*}[ht!]
-	\centering
 	\includegraphics[width=\textwidth]{Figures/nb_sols_entropy}
@@
 	\label{fig:freq}
+\end{figure*}
+
+\begin{figure*}[ht!]
+  \centering
+  \includegraphics[width=\textwidth]{Figures/seq_identity_10_30}
+  \caption{Sequence identity of \ourprog and \RNAinverse for $10$ and $30\%$ of \GC.}
 \end{figure*}

### 2013-03-12 ISMB/results_ISMB.tex
intermediate commit 2

@@
 
+Before starting our benchmark, we asses the need for our methods and performed an analysis of the \GCContent drift achieved with state-of-the-art software.
+Using our dataset of $50$ structures, we generated $100$ samples per structure with classical softwares who do not control the \GCContent.
+Namely,  \RNAinverse, \INFORNA, \NUPACK and \frankenstein.
+We show the distribution of the \GCContent of the sequences produced with these softwares in Fig.~\ref{fig:gcdist} those distributions.
+As anticipated, we observe a clear bias toward high {\GCContent}s and a complete absence of sequence with less than $30\%$ of \GC.
+This striking results motivates a need for methods that enable to explicitly control the \GCContent and more precisely that enable to design sequences with low \GCContent (i.e.
+30\% or less).
+}
+
+\begin{figure}[ht!]
+  \begin{center}
+    \includegraphics[width=0.5\textwidth]{Figures/histograme_5_gc_distribution_nornaexinv.png}
+  \end{center}
+  \caption{Overall \GCContent distribution for sequences designed using \RNAinverse, \INFORNA, \NUPACK and \frankenstein folding in the desired structure.}
+  \label{fig:gcdist}
+\end{figure}
+
+\highlight{
 \subsection{Success rate}
@@
 
-\highlight{
-Using the same dataset of $50$ structures, we generated $100$ samples per structure with classical softwares who do not control the \GCContent.
-Namely,  \RNAinverse, \INFORNA, \NUPACK and \frankenstein.
-We show the distribution of the \GCContent of the sequences produced with these softwares in Fig.~\ref{fig:gcdist} those distributions.
-As anticipated, we observe a clear bias toward high {\GCContent}s and a complete absence of sequence with less than $30\%$ of \GC.
-This striking results motivates a need for methods that enable to explicitly control the \GCContent and more precisely that enable to design sequences with low \GCContent (i.e.
-30\% or less).
-}
-
-\begin{figure}[ht!]
-  \begin{center}
-    \includegraphics[width=0.5\textwidth]{Figures/histograme_5_gc_distribution_nornaexinv.png}
-  \end{center}
-  \caption{Overall \GCContent distribution for sequences designed using \RNAinverse, \INFORNA, \NUPACK and \frankenstein folding in the desired structure.}
-  \label{fig:gcdist}
-\end{figure}
-
-\highlight{
-A desirable feature in sequence design, is to produce samples with a high sequence diversity and stable secondary structure. 
-Therefore, in the following we will use two useful measures which are the sequence identity of the samples, and the Boltzmann probability
-of the target structure in the low energy ensemble.
-}
 %The positional entropy is defined as:
@@
 %Intuitively, this measure captures the variability of sequences generated by a given method.
+
 \highlight{
+A desirable feature in sequence design, is to produce samples with a high sequence diversity and stable secondary structure. 
+Therefore, in the following we will use two useful measures which are the sequence identity of the samples, and the Boltzmann probability
+of the target structure in the low energy ensemble.
+
 The sequence identity is defined over a set $\mathcal{S}$ of aligned sequences (in our case, all sequences have the same length and can be trivially aligned) as :
@@
 
-  Figure~\ref{fig:nb_sols_entropy} shows the number of correct
-solutions as their average entropy and positions inside base pairs entropy, since 
-\ourprog  only influences the distribution of nucleotides inside 
-stacked base pairs. The most constraining factor for generating valid
- solutions seems to be  low \GCContent and the percentage of free nucleotides.
+%Figure~\ref{fig:nb_sols_entropy} shows the number of correct solutions as their average entropy and positions inside base pairs entropy, since \ourprog  only influences the distribution of nucleotides inside stacked base pairs.
+The most constraining factor for generating valid solutions seems to be  low \GCContent and the percentage of free nucleotides.
 
-%Our method is able to generate
+\highlight{
+Figure~\ref{fig:nb_sols_entropy} shows the number of solutions generated (i.e.
+sequences with a MFE structure identical to the target structure) and the average sequence identity of the sequences.
+Here, we note that low {\GCContent}s have a strong (negative) influence on the number of sequences generated, and in parallel also affect negatively the sequence diversity.
+This observation emphasizes the difficulty to  design sequences with low \GCContent.
+Once again, large percentages of free nucleotides increase the difficulty of the task.
 
-Also, a critical property in RNA sequence design is 
-the frequency of the MFE. 
-The sequence should be in its target conformation long enough to
-perform the desired action. Presented in Figure~\ref{fig:freq} we see that
-there is a slow decline of the frequency with the increase in size. Yet,
-for an average \GCContent, the frequency stays over $10\%$ even
-at size of $100$ nucleotides.
+The thermodynamical stability of the target structure on the designed sequence is another important property when estimating the performance of RNA design algorithms.
+We estimate the quality of our solutions in Figure~\ref{fig:freq}.
+First, we observe a slow decline of the structure stability (i.e.
+the frequency) when the target structure increases in size.
+Yet, for an average \GCContent, the frequency stays over $10\%$ even at size of $100$ nucleotides.
+Next, we note that for the most difficult target structures (i.e.
+the longer ones or those with high percentages of unpaired nucleotides in loops) the \GCContent have a limited (almost null) influence on the stability of the target structure on the designed sequence.
+By contrast, this less true for easiest and small structures with only few free nucleotides in internal loops.
+}
 
@@
 	\includegraphics[width=\textwidth]{Figures/nb_sols_entropy}
-	\caption{Number of solutions generated and their average entropy. 
-	The last row presents only the entropy of the nucleotides inside base 
-	pairs.}
+	\caption{Number of solutions generated by \ourprog and their average sequence identity.
+The curves are parameterized by the {\GCContent} of the sequences.
+The data are plotted vs.
+the length of the target (left), density of stacked base pairs (centre) and number of unpaired nucleotides  in loops (right).}
 	\label{fig:nb_sols_entropy}
@@
 	\includegraphics[width=\textwidth]{Figures/freq.png}
-	\caption{Average Boltzmann probability of the MFE.}
+	\caption{Thermodynamical stability of the target structure.
+The curves report the average Boltzmann probability of the target structure (which is also the MFE structure) at various {\GCContent}s w.r.t.
+the length of the target (left), density of stacked base pairs (centre) and number of unpaired nucleotides  in loops (right).}
 	\label{fig:freq}
@@
 
-\begin{figure*}[ht!]
-  \centering
-  \includegraphics[width=\textwidth]{Figures/seq_identity_10_30}
-  \caption{Sequence identity of \ourprog and \RNAinverse for $10$ and $30\%$ of \GC.}
-\end{figure*}
-
+\highlight{
 \subsection{Global sampling vs Local search vs Glocal approach}
-
-Here, we are interested in estimating the impact of the design methodology on the performances.
-More precisely, we aim to determine the merits of a global sampling approach (\ourprog), compared to a glocal procedure (\ourprog + \RNAinverse) and a local search methodology (\RNAinverse).
+To conclude this study, we estimate the impact of the design methodology on the performances.
+More precisely, we aim to determine the merits of a global sampling approach (\ourprog), compared to a glocal procedure (\ourprog + \RNAinverse) and a local search methodology (\RNASSD).
+To our knowledge, \RNASSD is the only other software that implements an explicit control of the \GCContent.
+We focus our analysis on the design of sequences with low {\GCContent}s (30\% and less) as they are almost impossible to design with classical software (See Figure~\ref{fig:gcdist}).
 
@@
 Therefore, a rejection algorithm based on \RNAinverse alone should not be considered a realistic solution for the design with constrained \GCContent.
+}
 
-{\em Ajouter considérations Base Pair entropy: Se rappeler, que moins, c'est mieux !}
+\begin{figure*}[ht!]
+  \centering
+  \includegraphics[width=\textwidth]{Figures/seq_identity_10_30}
+  \caption{Sequence identity of \ourprog and \RNAinverse for $10$ and $30\%$ of \GC.}
+\end{figure*}
 

### 2013-03-12 ISMB/results_ISMB.tex
more done fig 9

--- 2013-03-12 ISMB/results_ISMB.tex
@@
   Figure~\ref{fig:nb_sols_entropy} shows the number of correct
-solutions as their average entropy and positions inside base pairs entropy, since 
-\ourprog  only influences the distribution of nucleotides inside 
-stacked base pairs. The most constraining factor for generating valid
+solutions \highlight{as their average sequence identity. Since \ourprog
+only influences the distribution of nucleotides inside stacked base pairs, 
+we consider the \GCContent only inside those base pairs.}
+  %entropy and positions inside base pairs entropy, since 
+%\ourprog  only influences the distribution of nucleotides inside 
+%stacked base pairs. 
+The most constraining factor for generating valid
  solutions seems to be  low \GCContent and the percentage of free nucleotides.
@@
 	\includegraphics[width=\textwidth]{Figures/nb_sols_entropy}
-	\caption{Number of solutions generated and their average entropy. 
-	The last row presents only the entropy of the nucleotides inside base 
-	pairs.}
+	\caption{Number of solutions generated with \ourprog+\RNAinverse. and their average sequence identity. 
+	The second row is the sequence identity for a $10\%$\GC and the last for $30\%$\GC inside the base pairs. }
 	\label{fig:nb_sols_entropy}
@@
 
-\begin{figure*}[ht!]
-  \centering
-  \includegraphics[width=\textwidth]{Figures/seq_identity_10_30}
-  \caption{Sequence identity of \ourprog and \RNAinverse for $10$ and $30\%$ of \GC.}
-\end{figure*}
+%\begin{figure*}[ht!]
+%  \centering
+%  \includegraphics[width=\textwidth]{Figures/seq_identity_10_30}
+%  \caption{Sequence identity of \ourprog and \RNAinverse for $10$ and $30\%$ of \GC.}
+%\end{figure*}
 
@@
 Here, we are interested in estimating the impact of the design methodology on the performances.
-More precisely, we aim to determine the merits of a global sampling approach (\ourprog), compared to a glocal procedure (\ourprog + \RNAinverse) and a local search methodology (\RNAinverse).
+More precisely, we aim to determine the merits of a global sampling approach (\ourprog), compared to a glocal procedure (\ourprog + \RNAinverse) and local search methodologies (\RNAinverse, \INFORNA, \NUPACK and \frankenstein).
 
 Our results (See Fig.
-\ref{fig:rnainverse} in Supplementary Material) indicate that a glocal approach produces sequences that are consistently more diverse than those obtained with both {\em pure} methods alone.
-Furthermore, while \ourprog outperforms \RNAinverse for extreme pressures on the \GCContent, \RNAinverse produces more entropic sequences than \ourprog for the central \GCContent regime ($30\%-70\%$).
+\ref{fig:no_gc_controle} in Supplementary Material) indicate that a glocal approach produces sequences that are consistently more diverse than those obtained with both {\em pure} methods alone.
+Furthermore, while \ourprog outperforms \RNAinverse for extreme pressures on the \GCContent, \RNAinverse produces sequences with a lower sequence identity than \ourprog for the central \GCContent regime ($30\%-70\%$).
 
-It should be noted, however, than \RNAinverse produces sequences whose \GCContent is extremely concentrated (see Fig.), thereby requiring a large number of runs before producing a suitable \GCContent sequence.
+It should be noted, however, that \RNAinverse, \INFORNA, \NUPACK  and \frankenstein produce sequences whose \GCContent is extremely concentrated (see Fig.~\ref{fig:gcdist}), thereby requiring a large number of runs before producing a suitable \GCContent sequence.
 Therefore, a rejection algorithm based on \RNAinverse alone should not be considered a realistic solution for the design with constrained \GCContent.

### 2013-03-12 ISMB/results_ISMB.tex

--- 2013-03-12 ISMB/results_ISMB.tex
@@
 
+  Figure~\ref{fig:nb_sols_entropy} shows the number of correct
+solutions \highlight{as their average sequence identity. Since \ourprog
+only influences the distribution of nucleotides inside stacked base pairs, 
+we consider the \GCContent only inside those base pairs.}
+  %entropy and positions inside base pairs entropy, since 
+%\ourprog  only influences the distribution of nucleotides inside 
+%stacked base pairs. 
+The most constraining factor for generating valid
+ solutions seems to be  low \GCContent and the percentage of free nucleotides.
+
+%Our method is able to generate 
 %Figure~\ref{fig:nb_sols_entropy} shows the number of correct solutions as their average entropy and positions inside base pairs entropy, since \ourprog  only influences the distribution of nucleotides inside stacked base pairs.
@@
 	\includegraphics[width=\textwidth]{Figures/nb_sols_entropy}
-	\caption{Number of solutions generated by \ourprog and their average sequence identity.
+	\caption{Number of solutions generated with \ourprog+\RNAinverse. and their average sequence identity. 
+	The second row is the sequence identity for a $10\%$\GC and the last for $30\%$\GC inside the base pairs. }
+%	\caption{Number of solutions generated by \ourprog and their average sequence identity.
 The curves are parameterized by the {\GCContent} of the sequences.
@@
 
+%<<<<<<< HEAD
+
+%\begin{figure*}[ht!]
+%  \centering
+%  \includegraphics[width=\textwidth]{Figures/seq_identity_10_30}
+%  \caption{Sequence identity of \ourprog and \RNAinverse for $10$ and $30\%$ of \GC.}
+%\end{figure*}
+
+%\subsection{Global sampling vs Local search vs Glocal approach}
+
+%Here, we are interested in estimating the impact of the design methodology on the performances.
+More precisely, we aim to determine the merits of a global sampling approach (\ourprog), compared to a glocal procedure (\ourprog + \RNAinverse) and local search methodologies (\RNAinverse, \INFORNA, \NUPACK and \frankenstein).
+%=======
 \highlight{
@@
 We focus our analysis on the design of sequences with low {\GCContent}s (30\% and less) as they are almost impossible to design with classical software (See Figure~\ref{fig:gcdist}).
+%>>>>>>> af029e5428ff6e8980ef29cd0002a6fd72501bd9
 
 Our results (See Fig.
-\ref{fig:rnainverse} in Supplementary Material) indicate that a glocal approach produces sequences that are consistently more diverse than those obtained with both {\em pure} methods alone.
-Furthermore, while \ourprog outperforms \RNAinverse for extreme pressures on the \GCContent, \RNAinverse produces more entropic sequences than \ourprog for the central \GCContent regime ($30\%-70\%$).
+\ref{fig:no_gc_controle} in Supplementary Material) indicate that a glocal approach produces sequences that are consistently more diverse than those obtained with both {\em pure} methods alone.
+Furthermore, while \ourprog outperforms \RNAinverse for extreme pressures on the \GCContent, \RNAinverse produces sequences with a lower sequence identity than \ourprog for the central \GCContent regime ($30\%-70\%$).
 
-It should be noted, however, than \RNAinverse produces sequences whose \GCContent is extremely concentrated (see Fig.), thereby requiring a large number of runs before producing a suitable \GCContent sequence.
+%<<<<<<< HEAD
+%It should be noted, however, that \RNAinverse, \INFORNA, \NUPACK  and \frankenstein produce sequences whose \GCContent is extremely concentrated (see Fig.~\ref{fig:gcdist}), thereby requiring a large number of runs before producing a suitable \GCContent sequence.
+Therefore, a rejection algorithm based on \RNAinverse alone should not be considered a realistic solution for the design with constrained \GCContent.
+%=======
+It should be noted, however, than \RNAinverse produces sequences whose \GCContent is extremely concentrated (see Fig.~\ref{fig:gcdist}), thereby requiring a large number of runs before producing a suitable \GCContent sequence.
 Therefore, a rejection algorithm based on \RNAinverse alone should not be considered a realistic solution for the design with constrained \GCContent.
 }
+%>>>>>>> af029e5428ff6e8980ef29cd0002a6fd72501bd9
 
-\begin{figure*}[ht!]
-  \centering
-  \includegraphics[width=\textwidth]{Figures/seq_identity_10_30}
-  \caption{Sequence identity of \ourprog and \RNAinverse for $10$ and $30\%$ of \GC.}
-\end{figure*}
+%\begin{figure*}[ht!]
+%  \centering
+%  \includegraphics[width=\textwidth]{Figures/seq_identity_10_30}
+%  \caption{Sequence identity of \ourprog and \RNAinverse for $10$ and $30\%$ of \GC.}
+%\end{figure*}
 

### 2013-03-12 ISMB/results_ISMB.tex
intermediate results 3

--- 2013-03-12 ISMB/results_ISMB.tex
@@
 More precisely, we aim to determine the merits of a global sampling approach (\ourprog), compared to a glocal procedure (\ourprog + \RNAinverse) and a local search methodology (\RNASSD).
-To our knowledge, \RNASSD is the only other software that implements an explicit control of the \GCContent.
-We focus our analysis on the design of sequences with low {\GCContent}s (30\% and less) as they are almost impossible to design with classical software (See Figure~\ref{fig:gcdist}).
+To our knowledge, \RNASSD, beside \ourprog, is the only software that implements an explicit control of the \GCContent.
 
-Our results (See Fig.
-\ref{fig:rnainverse} in Supplementary Material) indicate that a glocal approach produces sequences that are consistently more diverse than those obtained with both {\em pure} methods alone.
-Furthermore, while \ourprog outperforms \RNAinverse for extreme pressures on the \GCContent, \RNAinverse produces more entropic sequences than \ourprog for the central \GCContent regime ($30\%-70\%$).
+Here, we compare the running time and the sequence diversity of the solutions produced by each software.
+In addition, we focus on the design of sequences with low {\GCContent}s (30\% and less) as they are almost impossible to design with classical software (See Figure~\ref{fig:gcdist}).
 
-It should be noted, however, than \RNAinverse produces sequences whose \GCContent is extremely concentrated (see Fig.), thereby requiring a large number of runs before producing a suitable \GCContent sequence.
-Therefore, a rejection algorithm based on \RNAinverse alone should not be considered a realistic solution for the design with constrained \GCContent.
+Figure~\ref{fig:time} shows the running time of each software.
+These data demonstrate the efficiency and scalability of our techniques.
+In particular, this figure suggests that our strategy has the potential to be applied efficiently for designing sequences on long (and difficult) target secondary structures at low \GCContent -- A task that could have not been achieved before due time requirements.
+
+Next, we show in Figure~\ref{fig:benchmark_RNASSD} the average sequence identity achieved by the various methods.
+Our results show that at extremely low {\GCContent}s (i.e.
+10\%), \ourprog slightly outperforms \RNASSD while this advantage becomes less evident when the \GCContent increases.
 }
@@
   \caption{Sequence identity of \ourprog and \RNAinverse for $10$ and $30\%$ of \GC.}
+  \label{fig:benchmark_RNASSD}
 \end{figure*}

### 2013-03-12 ISMB/results_ISMB.tex
result completed
remove SSD from supp mat

@@
 We show the distribution of the \GCContent of the sequences produced with these softwares in Fig.~\ref{fig:gcdist} those distributions.
+
 As anticipated, we observe a clear bias toward high {\GCContent}s and a complete absence of sequence with less than $30\%$ of \GC.
@@
 30\% or less).
+In order to provide a complete overview of the performance of \ourprog, we provide additional statistics for these software in the supplementary material.
 }
@@
 10\%), \ourprog slightly outperforms \RNASSD while this advantage becomes less evident when the \GCContent increases.
+Our experiments on higher {\GCContent}s (i.e.
+50\% and above) showed that our glocal strategy and the local search approach perform similarly.
+Similarly, we did not find any clear evidence that a global, local or glocal approach outperforms others when we compare at the thermodynamical stability of the target structure (data not shown).
 }

### 2013-03-12 ISMB/supplementary_ISMB.tex
remove SSD from supp mat

@@
 
-\subsection{Benchmark \ourprog vs \RNASSD}
+%\subsection{Benchmark \ourprog vs \RNASSD}
 %Using the same dataset of $50$ structures, we generated $100$ samples
@@
 
- For all structures that have been solved 
-by the three methods, only \RNASSD, only \texttt{Incarnation} and
-\texttt{Incarnation} followed by \RNAinverse, 
-we present for different concentration of \GCContent the average sequence identity in Figure.~\ref{fig:identity_50_70_90}
+%For all structures that have been solved  by the three methods, only \RNASSD, only \texttt{Incarnation} and \texttt{Incarnation} followed by \RNAinverse, we present for different concentration of \GCContent the average sequence identity in Figure.~\ref{fig:identity_50_70_90}
 %the average sample entropy and the average base pairs entropy in Figure.~\ref{fig:rnainverse}.
-\begin{figure*}[ht!]
-  \centering
-  \includegraphics[width=\textwidth]{Figures/seq_identity_50_70_90.png}
-  \caption{Sequence identity by average \GCContent of $50$, $70$ and $90\%$ for \ourprog,\ourprog+\RNAinverse and \RNASSD.}
-  \label{fig:identity_50_70_90}
-\end{figure*}
+%\begin{figure*}[ht!]
+%  \centering
+%  \includegraphics[width=\textwidth]{Figures/seq_identity_50_70_90.png}
+%  \caption{Sequence identity by average \GCContent of $50$, $70$ and $90\%$ for \ourprog,\ourprog+\RNAinverse and \RNASSD.}
+%  \label{fig:identity_50_70_90}
+%\end{figure*}
 %\begin{table}[h!]

### 2013-03-12 ISMB/conclusion_ISMB.tex
Typos + Changed sampling figure

@@
 
-In this article, we described a novel algorithm, \ourprog, for the RNA design problem, i.e. the design of an RNA sequence adopting a predefined secondary structure as minimal free-energy fold.
+In this article, we described a novel algorithm, \ourprog, for the RNA secondary structure design problem, i.e.
+the design of an RNA sequence adopting a predefined secondary structure as its minimal free-energy fold.
 Implementing a global sampling approach, it optimizes affinity towards the target secondary structure, while granting the user full control over the \GCContent of the resulting sequences.
@@
 A global sampling approach would capture the positive aspects of design, optimizing affinity towards a given structure while allowing the specification of expressive systems of constraints.
-Designed sequences would serve as a seed for a restricted local approach which, by breaking unwanted symmetries, would perform the negative part of the design, 
-while ideally maintaining obedience to the constraints.
+Designed sequences would serve as a seed for a restricted local approach which, by breaking unwanted symmetries, would perform the negative part of the design, while ideally maintaining obedience to the constraints.
 Another perspective of this work is the incorporation of the full Turner energy model, which should in principle yield better designs for unpaired regions.

### 2013-03-12 ISMB/results_ISMB.tex

@@
 
+%Here, we are interested in estimating the impact of the design methodology on the performances.
+More precisely, we aim to determine the merits of a global sampling approach (\ourprog), compared to a glocal procedure (\ourprog + \RNAinverse) and local search methodologies (\RNAinverse, \INFORNA, \NUPACK and \frankenstein).
+%=======
 \highlight{
 \subsection{Global sampling vs Local search vs Glocal approach}
+
 To conclude this study, we estimate the impact of the design methodology on the performances.
@@
 }
+%>>>>>>> af029e5428ff6e8980ef29cd0002a6fd72501bd9
 
+%\begin{figure*}[ht!]
+%  \centering
+%  \includegraphics[width=\textwidth]{Figures/seq_identity_10_30}
+%  \caption{Sequence identity of \ourprog and \RNAinverse for $10$ and $30\%$ of \GC.}
+%\end{figure*}
 \begin{figure*}[ht!]

### 2013-03-12 ISMB/results_ISMB.tex
updated

@@
 
-  Figure~\ref{fig:nb_sols_entropy} shows the number of correct
-solutions \highlight{as their average sequence identity. Since \ourprog
-only influences the distribution of nucleotides inside stacked base pairs, 
-we consider the \GCContent only inside those base pairs.}
   %entropy and positions inside base pairs entropy, since 
@@
 %stacked base pairs.
-The most constraining factor for generating valid
- solutions seems to be  low \GCContent and the percentage of free nucleotides.
 
@@
 Figure~\ref{fig:nb_sols_entropy} shows the number of solutions generated (i.e.
-sequences with a MFE structure identical to the target structure) and the average sequence identity of the sequences.
+sequences with a MFE structure identical to the target structure).
 Here, we note that low {\GCContent}s have a strong (negative) influence on the number of sequences generated, and in parallel also affect negatively the sequence diversity.

### 2013-03-12 ISMB/ResponseToReviewers-Incarnation-ISMB-2013.tex

@@
 For most, source code packages are available.}
-\Answer{In addition to \RNAinverse~\cite{Hofacker:1994}, we systematically benchmarked \ourprog against an extended selection of available tools, including \NUPACK~\cite{Zadeh:2011fk}, \frankenstein~\cite{Lyngso:2012vn}, \RNASSD~\cite{Aguirre-Hernandez:2007kx}, \INFORNA~\cite{Busch:2006uq} and \RNAexinv~\cite{Avihoo:2011fk}.
+\Answer{In addition to \RNAinverse~\cite{Hofacker:1994}, when possible, we systematically benchmarked \ourprog against an extended selection of available tools, including \NUPACK~\cite{Zadeh:2011fk}, \frankenstein~\cite{Lyngso:2012vn}, \RNASSD~\cite{Aguirre-Hernandez:2007kx}, \INFORNA~\cite{Busch:2006uq}.
+Technical difficulties prevented us to include \RNAexinv~\cite{Avihoo:2011fk} in this draft, but the latter could be easily included later.
 While the page limitations associated with a conference submission do not allow for a an exhaustive discussion, we now report in the revised manuscript some of the most striking results.
 Mainly, with the exception of \RNASSD, none of the latest version of our competitor allows an explicit control over the \GCContent, resulting in designs that are usually highly concentrated in \GCContent (with means ranging from 50\% to 80\% depending on the software).
-\textbf{FIXME: What should I add here?
-I think Vlad already added the histogram for the distribution of GC already (?)}
+At low {\GCContent}s, our results showed that \ourprog is more efficient (i.e. fast and scalable to large structure) than local search approaches and generate more diverse sequences.
 }

### 2013-03-12 ISMB/results_ISMB.tex
more typos

@@
 
-Figure~\ref{fig:time} presents the average times spent running \ourprog and \RNAinverse to generate one sequence
+Figure~\ref{fig:time} presents the average times spent running \ourprog+\RNAinverse to generate one sequence
 with the required \GCContent. As expected, the time grows linearly
-in function of the length of the structures for \ourprog.  A highly intriguing feature is the substantial decrease of the time 
-required for running \RNAinverse when $15\%$ of nucleotides are unpaired in the target structure.
-However, one must remain careful interpreting this observation, as the structures within this class all originate from the PDB, and are relatively small (for the complete STRAND DB, the average length is $\sim526$nts, compared to $\sim38$nts around 15\% unpaired bases).
+in function of the length of the structures for \ourprog.  
+%A highly intriguing feature is the substantial decrease of the time 
+%required for running \RNAinverse when $15\%$ of nucleotides are unpaired in the target structure.
+%However, one must remain careful interpreting this observation, as the structures within this class all originate from the PDB, and are relatively small (for the complete STRAND DB, the average length is $\sim526$nts, compared to $\sim38$nts around 15\% unpaired bases).
 
@@
 
-As discussed in Sec.~\ref{sec:implementation}, we noticed a highly decreased computational time needed to generate the sequences with $15\%$ free 
-nucleotides in the loops. We remark that those structures also yield a much lower structural sensitivity.
+%As discussed in Sec.~\ref{sec:implementation}, we noticed a highly decreased computational time needed to generate the sequences with $15\%$ free 
+%nucleotides in the loops. 
+We noticed a highly decreased structural sensitivity for the sequences with $15\%$ free nucleotides in the loops. 
+%A highly intriguing feature is the substantial decrease of the time 
+%required for running \RNAinverse when $15\%$ of nucleotides are unpaired in the target structure.
+However, one must remain careful interpreting this observation, as the structures within this class all originate from the PDB, and are relatively small (for the complete STRAND DB, the average length is $\sim526$nts, compared to $\sim38$nts around 15\% unpaired bases).
 
@@
 where $\mathcal{Z}^s$ is the partition function of sequence $s$.
-This measure tells us how dominant is a structure $\Struct$ in the Boltzmann ensemble of structure over a sequence $s$.
+This measure tells us how dominant is a structure $\Struct$ in the Boltzmann ensemble of structures over a sequence $s$.
 A high value implies a stable structure.
@@
 the longer ones or those with high percentages of unpaired nucleotides in loops) the \GCContent have a limited (almost null) influence on the stability of the target structure on the designed sequence.
-By contrast, this less true for easiest and small structures with only few free nucleotides in internal loops.
+By contrast, this is less true for easiest and small structures with only few free nucleotides in internal loops.
 }

### 2013-03-12 ISMB/results_ISMB.tex
fix caption and fig identity nb_sols

@@
 	\centering
-	\includegraphics[width=\textwidth]{Figures/nb_sols_entropy}
-	\caption{Number of solutions generated with \ourprog+\RNAinverse. and their average sequence identity. 
-	The second row is the sequence identity for a $10\%$\GC and the last for $30\%$\GC inside the base pairs. }
+	\includegraphics[width=\textwidth]{Figures/nb_sols_seq_identity.png}
+	\caption{Number of solutions generated with \ourprog+\RNAinverse on the first row and their average sequence identity on the second.} 
 %	\caption{Number of solutions generated by \ourprog and their average sequence identity.

### 2013-04-09 src/IncaRNAtion.py
precision help -b

@@ -877,7 +877,7 @@ def help():
     -m <int> 
       The max penality for an invalid base pair, -1 for infinity
     -b <int> 
-      print 'n' optimal sequences
+      print 'n' stochasticly backtracked sequences
     -no_profile <> 
       Doesn't output a profile
     -s_gc <target_gc> <nb_samples> 

### 2013-04-11 ISMB/Camera/Waldispuhl.337.tex
camera ready all files / BUG FIGURES EXT TO FIX

\documentclass{bioinfo}
\copyrightyear{2013}
\pubyear{2013}
%\usepackage[top=1in, bottom=1in,right=1in,left=1in]{geometry}                % See geometry.pdf to learn the layout options. There are lots.
%\geometry{letterpaper}                   % ... or a4paper or a5paper or ... 
%\geometry{landscape}                % Activate for for rotated page geometry
%\usepackage[parfill]{parskip}    % Activate to begin paragraphs with an empty line rather than an indent
%\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm} %pour les newtheorem
\usepackage{epstopdf}
\usepackage{framed}
\usepackage{xspace}
%\usepackage{tcolorbox}
\usepackage{relsize}
%\usepackage{natbib}
\usepackage[colorinlistoftodos]{todonotes}

\usepackage{soul}

\colorlet{shadecolor}{yellow!80!orange!50}
\sethlcolor{shadecolor}
\makeatletter

\renewenvironment{snugshade}{%
 \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
 \colorbox{shadecolor}{##1}\hskip-\fboxsep
     % There is no \@totalrightmargin, so:
     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
 \MakeFramed {\advance\hsize-\width
   \@totalleftmargin\z@ \linewidth\hsize
   \@setminipage}}%
 {\par\unskip\endMakeFramed}

\makeatother

\newcommand{\highlight}[1]{{\begin{snugshade}{#1}\end{snugshade}}}

%\usepackage{subfig} %Subfigures
\usepackage{subcaption} %Subfigures

\usepackage{tikz}
\usepackage{amsmath}
\usepackage{ifthen}
\usepackage{relsize}
\usepackage{multirow}
\usepackage{cleveref}
\usepackage{url}

\usetikzlibrary{positioning}
\usetikzlibrary{matrix}
\usetikzlibrary{calc}
\usetikzlibrary{shapes}
\usetikzlibrary{arrows}
\usetikzlibrary{fit}
\usetikzlibrary{decorations}
\usetikzlibrary{snakes}
\usetikzlibrary{shadows}
\pgfdeclarelayer{background}
\pgfdeclarelayer{backbackground}
\pgfdeclarelayer{foreground}
\pgfsetlayers{backbackground,background,main,foreground}

\usepackage[noend,ruled,vlined]{algorithm2e}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{description}{Description}

\crefmultiformat{equation}{Equations~(#2#1#3)}%
{ and~(#2#1#3)}{, (#2#1#3)}{ and~(#2#1#3)}

\newcommand{\RNAmutants}{\texttt{RNAmutants}\xspace}
\newcommand{\RNApyro}{\texttt{RNApyro}\xspace}
\newcommand{\RNAinverse}{\texttt{RNAinverse}\xspace}
\newcommand{\RNASSD}{\texttt{RNA-SSD}\xspace}
\newcommand{\INFORNA}{\texttt{INFO-RNA}\xspace}
\newcommand{\NUPACK}{\texttt{NUPACK:Design}\xspace}
\newcommand{\RNAiFOLD}{\texttt{RNAiFOLD}\xspace}
\newcommand{\frankenstein}{\texttt{Frnakenstein}\xspace}
\newcommand{\RNAexinv}{\texttt{RNAexinv}\xspace}
\newcommand{\rnaDesign}{\texttt{rnaDesign}\xspace}
\newcommand{\RNAensign}{\texttt{RNA-ensign}\xspace}
\newcommand{\GC}{\Gb\Cb}
\newcommand{\GCContent}{\Gb\Cb-content\xspace}
\newcommand{\SFold}{\texttt{SFold}\xspace}
\newcommand{\GCWeighted}{\Gb\Cb-weighted\xspace}

\newcommand{\ourprog}{\texttt{IncaRNAtion}\xspace}

\newcommand{\red}[1]{{\color{red}#1}}
\newcommand{\farna}{\texttt{FARNA}\xspace}
\newcommand{\mcfoldmcsym}{\texttt{MC-Pipeline}\xspace}
\newcommand{\mcfold}{\texttt{MC-Fold}\xspace}
\newcommand{\mcsym}{\texttt{MC-Sym}\xspace}
\newcommand{\nast}{\texttt{NAST}\xspace}
\newcommand{\ifoldrna}{\texttt{iFoldRNA}\xspace}
\newcommand{\rnafold}{\texttt{RNAfold}\xspace}
\newcommand{\rnasubopt}{\texttt{RNAsubopt}\xspace}
\newcommand{\rnawolf}{\texttt{RNAwolf}\xspace}
\newcommand{\rnastructure}{\texttt{RNAstructure}\xspace}
\newcommand{\contrafold}{\texttt{contrafold}\xspace}
\newcommand{\unafold}{\texttt{unafold}\xspace}
\newcommand{\rnadd}{\texttt{RNA2D3D}\xspace}
\newcommand{\assemble}{\texttt{assemble}\xspace}
\newcommand{\fred}{\texttt{FR3D}\xspace}
\newcommand{\rnajunction}{\texttt{RNAjunction}\xspace}
\newcommand{\rnamotif}{\texttt{RNAmotif}\xspace}
\newcommand{\treefolder}{\texttt{TreeFolder}\xspace}
\newcommand{\barnacle}{\texttt{BARNACLE}\xspace}
\newcommand{\contextfold}{\texttt{contextfold}\xspace}

\newcommand{\RNASTRAND}{{\sf RNA STRAND}\xspace}

%\newcommand{\citep}{\cite}
\newcommand{\Z}[2]{\mathcal{Z}{\substack{[#2]\\#1}}}
\newcommand{\B}{\mathcal{B}}
\newcommand{\Kron}{\delta}
\newcommand{\ub}{\bullet}
\newcommand{\op}{\texttt{(}}
\newcommand{\cp}{\texttt{)}}

\newcommand{\Ab}{{\sf{A}}\xspace}
\newcommand{\Cb}{{\sf{C}}\xspace}
\newcommand{\Gb}{{\sf{G}}\xspace}
\newcommand{\Ub}{{\sf{U}}\xspace}

\newcommand{\gc}{{\#\text{{\sf gc}}}}
\newcommand{\PE}[1]{E(#1)}
\newcommand{\EI}{\text{EI}}
\newcommand{\ES}{{{E}}}
\newcommand{\ISO}{\text{ISO}}
\newcommand{\Prob}{\mathbb{P}}
\newcommand{\Target}{{S}^*}
\newcommand{\Struct}{{S}}
\newcommand{\N}{{N}}
\newcommand{\CNBP}[1]{\gamma(#1)}
\newcommand{\BoolTrue}{{\sf T}}
\newcommand{\BoolFalse}{{\sf F}}

\newcommand{\CFramed}[2][gray]{\begin{tcolorbox}[colframe=#1,title=To be included only if space allows\ldots,colback=white]#1\end{tcolorbox}}
\renewcommand{\CFramed}[2][gray]{}

\newcommand{\oururl}{\url{http://csb.cs.mcgill.ca/incarnation/}}

\begin{document}
\firstpage{1}

\title{A weighted sampling algorithm for the design of RNA sequences with targeted secondary structure and nucleotides distribution}
\author{Vladimir Reinharz$^1$, Yann Ponty$^{2,*}$, J\'er\^{o}me Waldisp\"{u}hl$^{1}$\footnote{to whom correspondence should be addressed}}
\address{$^1$ School of Computer Science, McGill University, Montreal, Canada\\ $^2$ Laboratoire d'informatique, \'Ecole Polytechnique, Palaiseau, France.}

\history{Received on XXXXX; revised on XXXXX; accepted on XXXXX}

\editor{Associate Editor: XXXXXXX}

\maketitle
%\section{}
%\subsection{}
\begin{abstract}
\section{Motivations:} The design of RNA sequences folding into predefined secondary structures is a milestone for many synthetic biology and gene therapy studies. Most of the current software uses similar local search strategies (i.e. a random seed is progressively adapted to acquire the desired folding properties) and more importantly do not allow the user to control explicitly the nucleotide distribution such as the \GCContent in their sequences. However, the latter is an important criterion for large-scale applications as it could presumably be used to design sequences with better transcription rates and/or structural plasticity.

\section{Results:}
In this paper, we introduce \ourprog, a novel algorithm to design RNA sequences folding into target secondary structures with a predefined nucleotide distribution.  \ourprog uses a global sampling approach and weighted sampling techniques. We show that our approach is fast (i.e. running time comparable or better than local search methods), seed-less (we remove the bias of the seed in local search heuristics), and successfully generates high-quality sequences (i.e. thermodynamically stable) for any \GCContent. To complete this study, we develop an hybrid method combining our global sampling approach with local search strategies. Remarkably, our glocal methodology overcomes both local and global approaches  for sampling sequences with a specific GC content and target structure.
\section{Availability:} 
\ourprog is available at \href{http:// csb.cs.mcgill.ca/incarnation/}{csb.cs.mcgill.ca/incarnation/}
\section{Contact:} \href{jeromew@cs.mcgill.ca}{jeromew@cs.mcgill.ca}, \href{yann.ponty@lix.polytechnique.fr}{yann.ponty@lix.polytechnique.fr}\\
\noindent
\textbf{Key words:} RNA, secondary structure, design, weighted sampling, GC-content.
\end{abstract}

\section{Introduction}
\label{sec:introduction}

At the core of the emerging field of synthetic biology resides our capacity to design and re-engineer molecules with target functions. RNA molecules are well tailored for such applications. The ease to synthesize them (they are directly transcribed from DNA) and the broad diversity of catalytic and regulation functions they can perform enable to integrate \textit{de-novo} logic  circuits within living cells \citep{Rodrigo:2012fk} or re-program existing regulation mechanisms \citep{Chang:2012uq}. Future advances and applications of these techniques in gene-therapy studies will strongly rely on efficient computational methods to design and re-engineer RNA molecules.

Most of RNA functions are, at least partially, encoded by the three-dimensional molecular structures, which are themselves primarily determined by the secondary structures. The development of efficient algorithms for designing RNA sequences with pre-defined secondary structures is thus a milestone to enter the synthetic biology era. \RNAinverse pioneered RNA secondary structure design algorithms. It has been developed and distributed with the Vienna RNA package \citep{Hofacker:1994}. However, only posterior experimental studies revealed the potential and practical impact of these techniques. Thereby, during the last 6 years many improvements and variants of \RNAinverse have been proposed. Conceptually, almost all of existing algorithms follow the same approach. First a seed sequence is selected, then a local search strategy is used to mutate the seed and find, in its vicinity, a sequence with desired folding properties. Using this strategy, \INFORNA \citep{Busch:2006uq}, \RNASSD \citep{Aguirre-Hernandez:2007kx} and \NUPACK \citep{Zadeh:2011fk} significantly improved the performance of RNA secondary structure design algorithms. More recent research studies aimed to include more constraints in the selection criteria. \RNAexinv focused on the design of sequences with enhanced thermodynamical and mutational robustness \citep{Avihoo:2011fk}, while \frankenstein enables to design RNA with multiple target structures \citep{Lyngso:2012vn}.

We recently introduced with \RNAensign a novel paradigm for the search strategy of RNA secondary structure design algorithm \citep{Levin:2012kx}. Instead of a local search approach, we proposed a global sampling strategy of the mutational landscape based on the \RNAmutants algorithm~\citep{Waldispuhl2008}. This methodology offered promising performances, but suffered from  prohibitive runtime and memory consumption. Following our work, Garcia-Martin~\emph{et al} proposed \RNAiFOLD~\citep{Garcia-Martin:2013aa}, an alternate methodology that uses constraint programming techniques to prune the mutational landscape. While also suffering from prohibitive running times, it is worth noting that this latter algorithm also proposes a seed-less approach to the RNA secondary structure design problem.

In this paper, we introduce \ourprog, a RNA secondary structure design algorithm that benefits of our recent algorithmic advances \citep{Reinharz:2013aa} to expand our original \RNAensign algorithm \citep{Levin:2012kx}. \ourprog addresses previous limitations of \RNAensign and offers new functionalities. First, while our previous program had a running time complexity of $\mathcal{O}(n^5)$, \ourprog now runs in linear-time and space complexity, allowing it to demonstrate similar speeds as any local search algorithm. Next, \ourprog is \textit{seed-less}. Unlike \RNAensign, it does not require a seed sequence to initiate its search. Finally, \ourprog implements a novel algorithm based on weighted sampling techniques~\citep{Bodini2010} that enables us to control, for the first time, \textit{explicitly} the \GCContent of the solution. This functionality is essential because wild-type sequences within living organisms often present medium or low \GCContent, presumably to offer better transcription rates and/or structural plasticity. Previous programs do not allow to control this parameter and tend to output sequences having high \GCContent{}s~\citep{Lyngso:2012vn}.

We demonstrate the performance of our algorithms on a set of real RNA structures extracted from the \RNASTRAND database \citep{andronescu2008rna}. To complete this study, we develop an hybrid method combining our global sampling approach with local search strategies such as the one implemented in \RNAinverse.  Remarkably, our glocal methodology overcomes both local and global approaches  for sampling sequences with a specific GC content and target structure.

\section{Methods}
\label{sec:methods}

We introduce a probabilistic model for the design of RNA sequences with a specific \GCContent and folding into a predefined secondary structure.
For the sake of simplicity, we choose to base this proof-of-concept implementation on a simplified free-energy function $\ES(\cdot)$, which only considers the contributions of 
stacked canonical base-pairs. We show how a modification of the dynamic programming scheme used in \RNAmutants allows for the sampling of good and diverse design candidates, in linear time and space complexities.

\subsection{Definitions}

A targeted secondary structure $\Target$ of length $n$ is given as a non-crossing arc-annotated sequence,  where 
$\Target_i$ stands for the base-pairing position of position $i$ in $\Target$ if any (and, reciprocally, $\Target_{\Target_i}=i$), or $-1$ otherwise. 
In addition, let us denote by $\gc(s)$ the number of occurrences of \Gb and \Cb in an RNA sequence $s$.

\subsubsection{Simplified energy model}
We use a simplified free-energy model which only includes additive contributions from stacking base-pairs. Using individual values from the Turner 2004 model (retrieved from the NNDB~\citep{Turner2010}). Given a candidate sequence $s$ for a secondary structure $\Struct$, the free-energy of any sequence $s$ of length $|\Struct|$  is given by
\begin{align*}
  \ES(s,\Struct) = \sum_{\substack{(i,j)\to (i',j')\in \Struct\\ \text{stacking pairs}}}\ES^{\beta}_{s_is_j\to s_{i'}s_{j'}} 
\end{align*}
where $\ES^{\beta}_{ab\to a'b'}$ is set to $0$ if $ab=\varnothing$ (no base-pair to stack onto), the tabulated free-energy of stacking pairs $(ab)/(a'b')$ in the Turner model if available, or $\beta\in[0,\infty]$ for non-Watson-Crick/Wobble pairs (i.e. not in $\{\Gb\Ub,\Ub\Gb,\Cb\Gb,\Gb\Cb, \Ab\Ub\text{ or }\Ub\Ab\}$). This latter parameter allows one to choose whether to simply penalize invalid base pairs ($\beta>0$), or forbid them altogether ($\beta = +\infty$). 
Position-specific sequence constraints can also be enforced at this level (details omitted for the sake of clarity) by assigning to $\ES$ a $+\infty$ penalty (leading to a null probability) in the presence of a base incompatible with a user-specified constraint mask.

\subsubsection{\GCWeighted Boltzmann ensemble and distribution}

In order to counterbalance the documented tendency of sampling methods to generate \Gb\Cb-rich sequences~\citep{Levin:2012kx}, we introduce a parameter $x\in\mathbb{R}^+$, whose value will influence the \GCContent of generated sequences. For any secondary structure $\Struct$, the \GCWeighted-Boltzmann factor of a sequence $s$ is $\B^{[x]}_{\Struct}(s)$  such that
\begin{equation}
\B_{\Struct}^{[x]}(s) = e^{\frac{-\ES(s,\Struct)}{RT}}\cdot x^{\gc(s)}
\label{def:genBoltz}
\end{equation}
where $R$ is the Boltzmann constant and $T$ the temperature in Kelvin.

Summing the \GCWeighted-Boltzmann factor over all possible sequences of a given length $|\Struct|$, one obtains the \GCWeighted partition function $\mathcal{Z}_{\Struct}^{[x]}$, from which one defines the \GCWeighted Boltzmann probability $\Prob_{\Struct}^{[x]}(s)$ of each sequence $s$, respectively such that 
\begin{align}\mathcal{Z}_{\Struct}^{[x]} &= \sum_{\substack{|s|=n}}\B_{\Struct}^{[x]}(s)& \text{and}&&
\Prob_{\Struct}^{[x]}(s) &= \frac{\B_{\Struct}^{[x]}(s)}{\mathcal{Z}_{\Struct}^{[x]}}.\label{def:distribution}\end{align}

\subsection{Linear-time stochastic sampling algorithm for the \GCWeighted-Boltzmann ensemble}

Let us now describe a linear-time algorithm to sample sequences at random in the \GCWeighted Boltzmann distribution. This algorithm follows the general principles of the recursive approach to random generation~\citep{Wilf1977}, pioneered in the context of RNA by the \SFold algorithm~\citep{Ding2003}. The algorithm starts by precomputing the partition function restricted to each substructure occurring in the target structure, and then performs a series of recursive stochastic backtracks, using precomputed values to decide on the probability of each alternative.

\subsubsection{Precomputing the \GCWeighted partition function}\label{sec:pf}

Firstly, a dynamic programming algorithm computes $\Z{\N,\Struct}{a,b}$ the \GCWeighted partition function (the dependency in $x$ is omitted here for the sake of clarity) for a structure $\Struct$, assuming its (previously chosen) flanking nucleotides are $a$ and $b$ respectively, either forming a closing base-pair ($\N=\BoolTrue$) or not ($\N=\BoolFalse$).
Remark that the empty structure only supports the empty sequence, having energy $0$, so one has
\begin{equation}
	\Z{\BoolTrue,\varepsilon}{a,b}=\Z{\BoolFalse,\varepsilon}{a,b}= e^{-0/RT} = 1.
	\label{eq:Z_in}
\end{equation}

The general recursion scheme consists in three different terms, depending on the first position in $\Struct$:\\
{\bf Case 1.} First position is unpaired ($\Struct = \ub \Struct'$):
\begin{equation}
	\Z{\BoolTrue,\ub \Struct'}{a,b} = \Z{\BoolFalse,\ub \Struct'}{a,b}  :=
      \sum_{a'\in \B}  
      x^{\gc(a')}
      \cdot\Z{\BoolFalse,\Struct'}{a',b}; 
\label{eq:Z_unpaired}
\end{equation}
{\bf Case 2.} First position is paired with last position ($\Struct = \op \Struct' \cp$), stacking onto a pre-existing exterior pair ($\N=\BoolTrue$):
\begin{equation}
	\Z{\BoolTrue,\op \Struct' \cp}{a,b} :=
      \sum_{a',b'\in \B^2}
			 x^{\gc(a'.b')}
			 \cdot e^{\frac{-\ES^{\beta}_{ab \to a'b'}}{RT}}
			 \cdot \Z{\BoolTrue,\Struct'}{a',b'};
\label{eq:Z_stack}
\end{equation}
{\bf Case 3.} First position is involved in a base-pair ($\Struct = \op \Struct' \cp \Struct''$), which is not stacking onto an exterior base-pair ($\N=\BoolFalse$ or $\Struct''\neq \varepsilon$):
\begin{equation}
	\Z{\N,\op \Struct' \cp \Struct''}{a,b} :=
			 \displaystyle
      \sum_{a',b'\in \B^2}
      x^{\gc(a'.b')}
			\cdot e^{\frac{-\ES^{\beta}_{\varnothing\to a'b'}}{RT}}
      \cdot\Z{\BoolTrue,\Struct'}{a',b'}
      \cdot\Z{\BoolFalse,\Struct''}{b',b}.
\label{eq:Z_rec}
\end{equation}

Remark that the number of combinations of $a$, $b$ and $\N$ remains bounded by a constant, thus the complexity of computing $\Z{\N,\Struct}{a,b}$ mainly depends on the values taken by $\Struct$ upon subsequent recursive calls. Such values are entirely determined by $\Struct$ at any given step of the recursion, and their dependency can be summarized in a tree having $\Theta(|\Struct|)$. Therefore, the computation of $\Z{\N,\Target}{a,b}$ requires $\Theta(n)$ time and space using dynamic-programming.

\begin{figure*}
\resizebox{\linewidth}{!}{\input{Waldispuhl.337.fig.1.pgf}}
\caption{Stochastic backtrack procedure for a given substructure $\Struct$: Either the first position is left unpaired (top), a base-pair is formed between the two extremities, stacking onto an exterior base-pair (middle), or paired without creating a stacking, defining two regions on which subsequent recursive calls are needed (bottom). For the empty structure (omitted here), the empty sequence is returned. Positions indicated in red are assigned at the current stage of the backtrack.\label{fig:stochastic}}
\end{figure*}

\begin{algorithm}[t]
\DontPrintSemicolon
	\SetAlgoLined
\SetKwFunction{Backtrack}{SB$_x$}
\SetKwFunction{Random}{Random}
\newcommand{\rand}{{r}}
	$\rand \leftarrow $\Random$\left(\Z{\N,\Struct}{a,b}\right)$\tcp*[r]{Random real in $[0,\Z{\N,\Struct}{a,b}[$}
 \Switch{}{
   \lCase(\tcp*[f]{Empty structure}){$\Struct=\varepsilon$}{\Return{$\varepsilon$}}
   \Case(\tcp*[f]{First position is unpaired}){$\Struct=\ub\, \Struct'$}{		
   \For{$a'\in\B$}{
			$\rand \leftarrow \rand - x^{\gc(a')}\cdot \Z{\BoolFalse,\Struct'}{a',b}$\;
			\lIf{$\rand<0$}\Return{$a'.\Backtrack(a',b,\BoolFalse,\Struct')$}\;
  		}
   }
   \Case(\tcp*[f]{Extremities are involved in stacking base pair}){$\Struct=\text{\rm\op}\, \Struct' \,\text{\rm\cp}$  {\rm \bf and} $\N=\BoolTrue$}
   {		
			\For{$(a',b')\in\B\times\B$}
    {
				$\rand \leftarrow \rand -
			 	x^{\gc(a'.b')}
			 	\cdot e^{{-\ES^{\beta}_{ab \to a'b'}}/{RT}}
			 	\cdot \Z{\BoolTrue,\Struct'}{a',b'}	$\;
				\lIf{$\rand<0$}{
					\Return{$a'.\Backtrack(a',b',\BoolTrue,\Struct'). b'$}\;		
				}
			}
		}
   \Other(\tcp*[f]{First position is paired without a stacking pair})
   {
				\tcp{$\Struct=\text{\rm\op}\, \Struct' \,\text{\rm\cp}\,\Struct''$}
				\For{$(a',b')\in\B\times\B$}{
					$\rand \leftarrow \rand -	
       x^{\gc(a'.b')}
					\cdot e^{\frac{-\ES^{\beta}_{\varnothing\to a'b'}}{RT}}
  	    	\cdot\Z{\BoolFalse,\Struct'}{a',b'}
	    	  \cdot\Z{\BoolTrue,\Struct''}{b',b}$\;	
 					\lIf{\rand $<0$}{
						\Return{$a'
						.\Backtrack\left(a',b',\BoolTrue,\Struct'\right)
						.b'
						.\Backtrack\left(b',b,\BoolFalse,\Struct''\right)
						$}\;	
					}
				}
   }
 }
\caption{\protect\Backtrack$\left(a,b,\N,\Struct\right)$\label{alg:back}}
\end{algorithm}
\subsubsection{Stochastic backtrack}
Once the \GCWeighted partition functions have been computed and memorized, a stochastic backtrack starts from the target structure $\Target$ with any exterior bases $[a,b]$ and no nesting base-pair, corresponding to a call \Backtrack$\left(\varnothing,\varnothing,\BoolFalse,\Target\right)$ to Algorithm~\ref{alg:back}.  At each step, a suitable assignment for one or several positions is chosen, using probabilities derived from the precomputation, as illustrated by Figure~\ref{fig:stochastic}. One or several recursive calls over the appropriate substructures are then performed. On each recursive call, the algorithm assigns at least one nucleotide to a -- previously unassigned -- position. Moreover, the number of executions of each loops is bounded by a constant. Consequently, the complexity of Algorithm~\ref{alg:back} is in $\Theta(n)$ time and space.

\begin{figure}
\resizebox{\linewidth}{!}{\input{Waldispuhl.337.fig.2.pgf}}
\caption{General workflow of our adaptive sampling algorithm~\protect\citep{Waldispuhl2011}.}
\label{fig:workflow}
\end{figure}

\subsubsection{Self-adaptive sampling strategy}

Let us remind that our goal is to produce a set of sequences whose \GCContent matches a prescribed value $gc$.
An absolute tolerance $\kappa$ may be allowed, so that the \GCContent of any valid sequence must fall in  
$[gc-\kappa,gc+\kappa]$. Since sequences of arbitrary \GCContent may be generated by Algorithm~\ref{alg:back}, we use a rejection-based approach~\citep{Bodini2010}, previously adapted by the authors in a similar context~\citep{Waldispuhl2011}. This gives an algorithm which generates $k$ valid sequences in expected time $\Theta(k\cdot n\sqrt{n})$ when $\kappa=0$ (or $\Theta(k\cdot n)$ when $\kappa$ is a positive constant) and memory in $\Theta(k\cdot n)$.
A complete analysis of the rejection process can be found in an earlier contribution~\citep{Waldispuhl2011}, but let us briefly outline the approach, and the main arguments used to establish its complexity.

As summarized by Figure~\ref{fig:workflow}, our adaptive sampling approach simply generates sets of sequences by repeatedly running the stochastic backtrack algorithm. The average \GCContent induced by the current value of the $x$ parameter, can then be adequately estimated from the sample, or computed exactly using recent algorithmic advances~\citep{Ponty2011}. The set of sequences is filtered to only retain valid sequences. The value of the parameter $x$ is then adapted to match the average \GCContent (induced by the value of $x$) with the targeted one.
It can be shown that the expected \GCContent is a continuous and strictly increasing monotonic function of $x$, whose limits are $0$ when $x=0$ and $n$ when $x\to +\infty$. Consequently, for any targeted \GCContent $gc\in[0\%,100\%]$, there exists a unique value $x_{gc}$ such that generated sequences feature, on the average, the right \GCContent.
In practice, a simple binary search~\citep{Waldispuhl2011} is used in our implementation, and typically converges after very few iterations. An optimal value for $x$ can also be derived analytically using interpolation after $\Theta(n)$ evaluations of $\Z{i,j}{a,b}$ for different candidate values of $x$, as previously noted~\citep{Waldispuhl2011} and could be implemented using the Fast-Fourier Transform~\citep{Senter2012}.

\subsubsection{Overall complexity}

It was previously established~\citep{Waldispuhl2011} that, for each value of $x$, there exists constants $\mu_x$ and $\sigma_x$ such that the distribution of \GCContent asymptotically converges towards a normal law having expectation in $\mu_x\cdot n\cdot(1+o(1))$ and standard deviation in $\sigma_x\cdot\sqrt{n}\cdot(1+o(1))$.
Furthermore, the distribution of \GCContent is highly concentrated, as asserted by its limited standard deviation, therefore the expected number of attempts required to generate a valid sequence when $\kappa=0$ (resp. $\kappa\in\Omega(1/\sqrt n)$) grows like $\Theta(\sqrt{n})$ (resp. $\Theta(1)$, i.e. a constant), leading to the announced complexities. Formally, since a suitable weight $x$ must be recomputed for each targeted structure and \GCContent, then the number $M$ of iterations required for the converge can be accounted for explicitly, leading to time complexities in $\Theta((M+\sqrt{n})\cdot k\cdot n)$ (if $\kappa=0$, i.e. without any tolerance) and $\Theta(M\cdot k\cdot n)$ (if $\kappa>0$).

\subsection{Postprocessing unpaired regions: A local/global (glocal) hybrid approach}
\label{subsec:glocal_method}
Due to our simplified energy model, unpaired regions are not subject to design constraints other than the \GCContent, leading to modest probabilities for refolded design candidates to match the targeted structure. To improve these performances and test the complementarity  of our global sampling approach with previous contributions based on local search, we used the \RNAinverse software to redesign unpaired regions. We specified a constraint mask to prevent stacking base-pairs from being modified and, whenever necessary, reestablished their content {\em a posteriori}, as \RNAinverse has been witnessed to take some liberties with constraints masks. As shown in Table~\ref{table:impact_on_gc} (Supplementary material), this postprocessing does not drastically alter the \GCContent, so the glocal approach reasonably addresses the constrained \GCContent design problem.

\section{Results}
\label{sec:results}

\subsection{Implementation}
\label{sec:implementation}
Our software, \ourprog, was implemented in {\tt Python~2.7}. We used
\RNAinverse from the \textit{Vienna Package 2.0}~\citep{Hofacker:1994}.
All time benchmarks were run on a single AMD Opteron(tm) 6278 Processor  at 2.4 GHz with cache of 512 KB.
The penalty $\beta$, associated with invalid base-pairs, was set to $15$.

\begin{figure}[ht!]
	\centering
 	\includegraphics[width=0.5\textwidth]{Waldispuhl.337.fig.3.pdf}
	\caption{
  Average time in seconds to generate one sequence for \ourprog and \RNAinverse.}
	\label{fig:time}	
\end{figure}

Figure~\ref{fig:time} presents the average times spent running \ourprog+\RNAinverse to generate one sequence
with the required \GCContent. As expected, the time grows linearly
in function of the length of the structures for \ourprog.

\subsection{Dataset}
To evaluate the quality of our method, we used secondary
structures from the \RNASTRAND database~\citep{andronescu2008rna}.
Those are known secondary structure from a variety of organisms.
We considered a subset of $50$ structures selected by~\cite{Levin:2012kx}, 
whose length ranges between $20$ and $100$ nucleotides. 
To ease the visualization of results, we clustered together structures
having similar length, stacks density and proportion of free nucleotides in loops,
leading to distributions of structures shown in Figure~\ref{fig:bins}.

 \begin{figure*}[ht!]
 	\centering
	\includegraphics[width=\textwidth]{Waldispuhl.337.fig.4.pdf}
	\caption{Number of secondary structures per bin, according to our three clustering criteria.}
	\label{fig:bins}
 \end{figure*}

\subsection{Design}
We ran our method as follows. First, we sampled approximately $100$ sequences per structure. Then, we use these sequences 
as seed in \RNAinverse. Finally, we computed the MFE with the \rnafold program from the \textit{Vienna Package 2.0} \citep{Hofacker:1994}.

Before starting our benchmark, we asses the need for our methods and performed an analysis of the \GCContent drift achieved with state-of-the-art software. Using our dataset of $50$ structures, we generated $100$ samples per structure with classical softwares who do not control the \GCContent. Namely,  \RNAinverse, \INFORNA, \NUPACK and \frankenstein.  We show the distribution of the \GCContent of the sequences produced with these softwares in Fig.~\ref{fig:gcdist} those distributions.

As anticipated, we observe a clear bias toward high {\GCContent}s and a complete absence of sequence with less than $30\%$ of \GC. This striking results motivates a need for methods that enable to explicitly control the \GCContent and more precisely that enable to design sequences with low \GCContent (i.e. 30\% or less). In order to provide a complete overview of the performance of \ourprog, we provide additional statistics for these software in the supplementary material.

\begin{figure}[ht!]
  \begin{center}
    \includegraphics[width=0.5\textwidth]{Waldispuhl.337.fig.5.pdf}
  \end{center}
  \caption{Overall \GCContent distribution for sequences designed using \RNAinverse, \INFORNA, \NUPACK and \frankenstein folding in the desired structure.}
  \label{fig:gcdist}
\end{figure}

\subsection{Success rate}
We started by estimating the success rate of our methodology and computed the percentage of sequences with a MFE structure identical to the target secondary structure. Figure~\ref{fig:mfe_struct_solved_noinverse} shows our results. We clearly see that before the post-processing step (i.e. \RNAinverse) the sequences sampled by \ourprog have a low success rate (first row). As mentioned earlier, this could be explained by the fact that no selection criterion has been at this stage applied to unpaired nucleotides. Remarkably, after the local search optimization (with \RNAinverse) of nucleotides in unpaired regions (second row), we observe a dramatic improvement of our success rate. As expected, we observed that length is, in general, not a good predictor for the hardness of designing a structure. Instead, a high number of free nucleotides in the structure seems to be a  good measure of the hardness of its design. Similarly, these data also show that designing sequences with low \GCContent is challenging for all types of targets.

\begin{figure*}[ht!]	
	\centering
	\includegraphics[width=\textwidth]{Waldispuhl.337.fig.6.pdf}
	\caption{Success rate \ourprog before and after \RNAinverse post-processing. The first row shows the percentage of sampled sequences folding into the target when using only \ourprog. The second shows after processing previous results with \RNAinverse.}
	\label{fig:mfe_struct_solved_noinverse}	
\end{figure*}

We investigated further the quality of the sequences generated by \ourprog. In particular, we estimated the capacity of our methods to generate ``good'' sequences with desired folding capabilities regardless of the property  to fold exactly into the target structure. In Figure~\ref{fig:ss_sens}, we show the ratio of well predicted base pairs in the MFE structure of our sampled sequences. As above, we can observe that, in all cases, the sequences that are the hardest to design are those with an extremely low \GCContent. Indeed, the energetic contribution of the base pairs to the stability of the structure is weaker. Interestingly, we also notice that the most accurate sequences yield a \GCContent of $70\pm 10\%$. Overall, we observe that all our samples have good folding properties, and that there is a correlation between the ``precision'' of the samples and the hardness of the design.

We noticed a highly decreased structural sensitivity for the sequences with $15\%$ free nucleotides in the loops. 
However, one must remain careful interpreting this observation, as the structures within this class all originate from the PDB, and are relatively small (for the complete STRAND DB, the average length is $\sim526$nts, compared to $\sim38$nts around 15\% unpaired bases).

\begin{figure*}[ht!]
 	\centering
	\includegraphics[width=\textwidth]{Waldispuhl.337.fig.7.pdf}
	\caption{Structural sensitivity (i.e. $\#$ well predicted base pairs / $\#$ base pairs in target) of the sampled sequences MFE. }
	\label{fig:ss_sens}	
\end{figure*}

\subsection{Properties of designed sequences}

In this section, we further analyze the generated sequences with a MFE structure that folds into the target structure.

A desirable feature in sequence design, is to produce samples with a high sequence diversity and stable secondary structure. 
Therefore, in the following we will use two useful measures which are the sequence identity of the samples, and the Boltzmann probability
of the target structure in the low energy ensemble.

The sequence identity is defined over a set $\mathcal{S}$ of aligned sequences (in our case, all sequences have the same length and can be trivially aligned) as :
\begin{align}
  \sum_{s^1,s^2 \in\mathcal{S}\times\mathcal{S}}\left(\frac{1}{|s^1|}{\sum_{\substack{i\\s^1_i\equiv s^2_i}}1}\right) &&& \text{Seq. identity}
\end{align}
where $s_i$ is the nucleotide at position $i$ in sequence $s$. Intuitively, this measure captures the diversity of sequences generated by a given method. Next, the Boltzmann frequency is defined, for a structure $\Struct$ and a sequence $s$ as:
\begin{align}
  e^{\frac{-\ES(s,\Struct)}{RT}}/\mathcal{Z}^s  &&&\text{Frequency}
\end{align}
where $\mathcal{Z}^s$ is the partition function of sequence $s$. This measure tells us how dominant is a structure $\Struct$ in the Boltzmann ensemble of structures over a sequence $s$. A high value implies a stable structure. We compute this frequency with \rnafold from the \texttt{Vienna Package 2.0} \citep{Hofacker:1994}.

Figure~\ref{fig:nb_sols_entropy} shows the number of solutions generated (i.e. sequences with a MFE structure identical to the target structure). Here, we note that low {\GCContent}s have a strong (negative) influence on the number of sequences generated, and in parallel also affect negatively the sequence diversity. This observation emphasizes the difficulty to  design sequences with low \GCContent. Once again, large percentages of free nucleotides increase the difficulty of the task.

The thermodynamical stability of the target structure on the designed sequence is another important property when estimating the performance of RNA design algorithms. We estimate the quality of our solutions in Figure~\ref{fig:freq}. First, we observe a slow decline of the structure stability (i.e. the frequency) when the target structure increases in size. Yet, for an average \GCContent, the frequency stays over $10\%$ even at size of $100$ nucleotides. Next, we note that for the most difficult target structures (i.e. the longer ones or those with high percentages of unpaired nucleotides in loops) the \GCContent have a limited (almost null) influence on the stability of the target structure on the designed sequence. By contrast, this is less true for easiest and small structures with only few free nucleotides in internal loops.

\begin{figure*}[ht!]
	\centering
	\includegraphics[width=\textwidth]{Waldispuhl.337.fig.8.pdf}
	\caption{Number of solutions generated with \ourprog+\RNAinverse on the first row and their average sequence identity on the second.} 
	\label{fig:nb_sols_entropy}
\end{figure*}

\begin{figure*}[ht!]
	\centering
	\includegraphics[width=\textwidth]{Waldispuhl.337.fig.9.pdf}
	\caption{Thermodynamical stability of the target structure. The curves report the average Boltzmann probability of the target structure (which is also the MFE structure) at various {\GCContent}s w.r.t. the length of the target (left), density of stacked base pairs (centre) and number of unpaired nucleotides  in loops (right).}
	\label{fig:freq}
\end{figure*}

\subsection{Global sampling vs Local search vs Glocal approach}

To conclude this study, we estimate the impact of the design methodology on the performances. More precisely, we aim to determine the merits of a global sampling approach (\ourprog), compared to a glocal procedure (\ourprog + \RNAinverse) and a local search methodology (\RNASSD). To our knowledge, \RNASSD, beside \ourprog, is the only software that implements an explicit control of the \GCContent.

Here, we compare the running time and the sequence diversity of the solutions produced by each software. In addition, we focus on the design of sequences with low {\GCContent}s (30\% and less) as they are almost impossible to design with classical software (See Figure~\ref{fig:gcdist}).

Figure~\ref{fig:time} shows the running time of each software. These data demonstrate the efficiency and scalability of our techniques. In particular, this figure suggests that our strategy has the potential to be applied efficiently for designing sequences on long (and difficult) target secondary structures at low \GCContent -- A task that could have not been achieved before due time requirements.

Next, we show in Figure~\ref{fig:benchmark_RNASSD} the average sequence identity achieved by the various methods. Our results show that at extremely low {\GCContent}s (i.e. 10\%), \ourprog slightly outperforms \RNASSD while this advantage becomes less evident when the \GCContent increases. Our experiments on higher {\GCContent}s (i.e. 50\% and above) showed that our glocal strategy and the local search approach perform similarly. Similarly, we did not find any clear evidence that a global, local or glocal approach outperforms others when we compare at the thermodynamical stability of the target structure (data not shown).

\begin{figure*}[ht!]
  \centering
  \includegraphics[width=\textwidth]{Waldispuhl.337.fig.10.pdf}
  \caption{Sequence identity of \ourprog and \RNAinverse for $10$ and $30\%$ of \GC.}
  \label{fig:benchmark_RNASSD}
\end{figure*}

\section{Conclusion}

\label{sec:conclusion}

In this article, we described a novel algorithm, \ourprog, for the RNA secondary structure design problem, i.e. the design of an RNA sequence adopting a predefined secondary structure as its minimal free-energy fold. Implementing a global sampling approach, it optimizes affinity towards the target secondary structure, while granting the user full control over the \GCContent of the resulting sequences.
This extended control does not necessarily induce additional computational demands, and we showed the linear complexity of both the preprocessing stage and the generation of candidate sequences for the design, allowing for the design of larger and more complex secondary structures in a matter of minutes on a single processor (e.g. $\sim$28 mins for 100 candidate sequences for a $\sim$1500nts 16s rRNA). We evaluated the method on a benchmark composed of target secondary structures extracted from the \RNASTRAND database. We observed good overall success rate, with the notable exception of very low targeted \GCContent ($10\%$), and a good to excellent entropy within designed candidates.
Finally, we implemented an hybrid approach, using the \RNAinverse software as a post-processing step for unpaired regions. This approach greatly increased the success rate of the method, allowing for the design of highly diverse candidates for almost all of the structures in our benchmark, while largely preserving the targeted \GCContent.

In the future, we would like to complement this study by further investigating the potential of hybrid local/global -- or {\em glocal} -- approaches. A global sampling approach would capture the positive aspects of design, optimizing affinity towards a given structure while allowing the specification of expressive systems of constraints. Designed sequences would serve as a seed for a restricted local approach which, by breaking unwanted symmetries, would perform the negative part of the design, while ideally maintaining obedience to the constraints. Another perspective of this work is the incorporation of the full Turner energy model, which should in principle yield better designs for unpaired regions.

\section{Acknowledgments}
\label{sec:acknowledgments}
The authors would like to thank Rob Knight for his suggestions and comments.
This work was funded by the French Agence Nationale de la Recherche (ANR) through the {\sc Magnum} {\tt ANR 2010 BLAN 0204} 
project (to YP), the
FQRNT team grant 232983 (to VR and JW) and NSERC Discovery grant 219671 (to JW). This paper has been developed as a result of a mobility stay funded by the Erasmus Mundus Programme of the European Commission under the Transatlantic Partnership for Excellence in Engineering – TEE Project (VR).

\bibliographystyle{natbib}
\bibliography{Waldispuhl.337.bib}

%\clearpage

%\input{supplementary_ISMB.tex}

\end{document}

### 2013-04-11 ISMB/main_ISMB.tex
Added highlight macro and fixed response to reviewers (still to be completed)
Last stuff
intermediate commit 2
camera ready all files / BUG FIGURES EXT TO FIX

@@
 \usepackage{epstopdf}
+\usepackage{framed}
 \usepackage{xspace}
-\usepackage{tcolorbox}
+%\usepackage{tcolorbox}
 \usepackage{relsize}
@@
 \usepackage[colorinlistoftodos]{todonotes}
+
+\usepackage{soul}
+
+\colorlet{shadecolor}{yellow!80!orange!50}
+\sethlcolor{shadecolor}
+\makeatletter
+
+\renewenvironment{snugshade}{%
+ \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
+ \colorbox{shadecolor}{##1}\hskip-\fboxsep
+     % There is no \@totalrightmargin, so:
+     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
+ \MakeFramed {\advance\hsize-\width
+   \@totalleftmargin\z@ \linewidth\hsize
+   \@setminipage}}%
+ {\par\unskip\endMakeFramed}
+
+\makeatother
+
+\newcommand{\highlight}[1]{{\begin{snugshade}{#1}\end{snugshade}}}
 
@@
 
-\maketitle
-%\section{}
-%\subsection{}
-\begin{abstract}
-\section{Motivations:} The design of RNA sequences folding into predefined secondary structures is a milestone for many synthetic biology and gene therapy studies.
+%\maketitle
+%%\section{}
+%%\subsection{}
+%\begin{abstract}
+%\section{Motivations:} The design of RNA sequences folding into predefined secondary structures is a milestone for many synthetic biology and gene therapy studies.
 Most of the current software uses similar local search strategies (i.e.
@@
 However, the latter is an important criterion for large-scale applications as it could presumably be used to design sequences with better transcription rates and/or structural plasticity.
-
-\section{Results:}
-In this paper, we introduce \ourprog, a novel algorithm to design RNA sequences folding into target secondary structures with a predefined nucleotide distribution.
+%
+%\section{Results:}
+%In this paper, we introduce \ourprog, a novel algorithm to design RNA sequences folding into target secondary structures with a predefined nucleotide distribution.
 \ourprog uses a global sampling approach and weighted sampling techniques.
@@
 Remarkably, our glocal methodology overcomes both local and global approaches  for sampling sequences with a specific GC content and target structure.
-\section{Availability:} 
-\ourprog is available at \href{http:// csb.cs.mcgill.ca/incarnation/}{csb.cs.mcgill.ca/incarnation/}
-\section{Contact:} \href{jeromew@cs.mcgill.ca}{jeromew@cs.mcgill.ca}, \href{yann.ponty@lix.polytechnique.fr}{yann.ponty@lix.polytechnique.fr}\\
-\noindent
-\textbf{Key words:} RNA, secondary structure, design, weighted sampling, GC-content.
-\end{abstract}
-
-\input{introduction_ISMB.tex}
-
-\input{methods_ISMB.tex}
-
-\input{results_ISMB.tex}
-
-\input{conclusion_ISMB.tex}
-
-\input{acknowledgments_ISMB.tex}
-
-\newpage
-
-\bibliographystyle{natbib}
-\bibliography{RNApyro}
-
-\newpage
+%\section{Availability:} 
+%\ourprog is available at \href{http:// csb.cs.mcgill.ca/incarnation/}{csb.cs.mcgill.ca/incarnation/}
+%\section{Contact:} \href{jeromew@cs.mcgill.ca}{jeromew@cs.mcgill.ca}, \href{yann.ponty@lix.polytechnique.fr}{yann.ponty@lix.polytechnique.fr}\\
+%\noindent
+%\textbf{Key words:} RNA, secondary structure, design, weighted sampling, GC-content.
+%\end{abstract}
+%
+%
+%\input{introduction_ISMB.tex}
+%
+%\input{methods_ISMB.tex}
+%
+%\input{results_ISMB.tex}
+%
+%\input{conclusion_ISMB.tex}
+%
+%\input{acknowledgments_ISMB.tex}
+%
+%\bibliographystyle{natbib}
+%\bibliography{RNApyro}
+%
+%\clearpage
 

### 2013-04-11 ISMB/supplementary_ISMB.tex
typos corrected

@@
 We present in Fig.~\ref{fig:no_gc_control} the average sequence identity and frequency
-for sequences generated them.
+for the sequences that generated them.
 
@@
   \centering
-  \includegraphics[width=\textwidth]{Figures/entropy_nogc}
+  \includegraphics[width=\textwidth]{Figures/entropy_nogc.pdf}
   \caption{The average sequence identity and frequency for softwares without \GCContent control.}
@@
   \centering
-  \includegraphics[width=\textwidth]{Figures/struct_solved_vsrnainverse.png}
+  \includegraphics[width=\textwidth]{Figures/struct_solved_vsrnainverse.pdf}
   \caption{The first row shows the number of structures for which one generated sequence has the 

### 2013-04-11 ISMB/Camera/Waldispuhl.337.tex
typos corrected
fix fig links
typo title

@@
 
-\title{A weighted sampling algorithm for the design of RNA sequences with targeted secondary structure and nucleotides distribution}
+\title{A weighted sampling algorithm for the design of RNA sequences with targeted secondary structure and nucleotide distribution}
 \author{Vladimir Reinharz$^1$, Yann Ponty$^{2,*}$, J\'er\^{o}me Waldisp\"{u}hl$^{1}$\footnote{to whom correspondence should be addressed}}
@@
 	\centering
- 	\includegraphics[width=0.5\textwidth]{Waldispuhl.337.fig.3.pdf}
+ 	\includegraphics[width=0.5\textwidth]{{Waldispuhl.337.fig.3}.pdf}
 	\caption{
@@
  	\centering
-	\includegraphics[width=\textwidth]{Waldispuhl.337.fig.4.pdf}
+	\includegraphics[width=\textwidth]{{Waldispuhl.337.fig.4}.pdf}
 	\caption{Number of secondary structures per bin, according to our three clustering criteria.}
@@
   \begin{center}
-    \includegraphics[width=0.5\textwidth]{Waldispuhl.337.fig.5.pdf}
+    \includegraphics[width=0.5\textwidth]{{Waldispuhl.337.fig.5}.pdf}
   \end{center}
@@
 	\centering
-	\includegraphics[width=\textwidth]{Waldispuhl.337.fig.6.pdf}
+	\includegraphics[width=\textwidth]{{Waldispuhl.337.fig.6}.pdf}
 	\caption{Success rate \ourprog before and after \RNAinverse post-processing.
@@
  	\centering
-	\includegraphics[width=\textwidth]{Waldispuhl.337.fig.7.pdf}
+	\includegraphics[width=\textwidth]{{Waldispuhl.337.fig.7}.pdf}
 	\caption{Structural sensitivity (i.e. $\#$ well predicted base pairs / $\#$ base pairs in target) of the sampled sequences MFE. }
@@
 	\centering
-	\includegraphics[width=\textwidth]{Waldispuhl.337.fig.8.pdf}
+	\includegraphics[width=\textwidth]{{Waldispuhl.337.fig.8}.pdf}
 	\caption{Number of solutions generated with \ourprog+\RNAinverse on the first row and their average sequence identity on the second.} 
@@
 	\centering
-	\includegraphics[width=\textwidth]{Waldispuhl.337.fig.9.pdf}
+	\includegraphics[width=\textwidth]{{Waldispuhl.337.fig.9}.pdf}
 	\caption{Thermodynamical stability of the target structure.
@@
 More precisely, we aim to determine the merits of a global sampling approach (\ourprog), compared to a glocal procedure (\ourprog + \RNAinverse) and a local search methodology (\RNASSD).
-To our knowledge, \RNASSD, beside \ourprog, is the only software that implements an explicit control of the \GCContent.
+To the best of our knowledge, \RNASSD, beside \ourprog, is the only software that implements an explicit control of the \GCContent.
 
@@
   \centering
-  \includegraphics[width=\textwidth]{Waldispuhl.337.fig.10.pdf}
+  \includegraphics[width=\textwidth]{{Waldispuhl.337.fig.10}.pdf}
   \caption{Sequence identity of \ourprog and \RNAinverse for $10$ and $30\%$ of \GC.}

### 2013-04-19 ISMB/Camera/Waldispuhl.337.tex
added figures pgf -> pdf (1,2)
moved fig , redid fig time

@@
 
+\begin{figure}
+%\resizebox{\linewidth}{!}{\input{{Waldispuhl.337.fig.2}.pdf}}
+  \centering
+  \includegraphics[width=0.45\textwidth]{{Waldispuhl.337.fig.1}.pdf}
+\caption{General workflow of our adaptive sampling algorithm~\protect\citep{Waldispuhl2011}.}
+\label{fig:workflow}
+\end{figure}
+
 \begin{figure*}
-\resizebox{\linewidth}{!}{\input{Waldispuhl.337.fig.1.pgf}}
+%\resizebox{\linewidth}{!}{\input{{Waldispuhl.337.fig.1}.pdf}}
+  \centering
+  \includegraphics[width=\textwidth]{{Waldispuhl.337.fig.2}.pdf}
 \caption{Stochastic backtrack procedure for a given substructure $\Struct$: Either the first position is left unpaired (top), a base-pair is formed between the two extremities, stacking onto an exterior base-pair (middle), or paired without creating a stacking, defining two regions on which subsequent recursive calls are needed (bottom).
@@
 
-\begin{figure}
-\resizebox{\linewidth}{!}{\input{Waldispuhl.337.fig.2.pgf}}
-\caption{General workflow of our adaptive sampling algorithm~\protect\citep{Waldispuhl2011}.}
-\label{fig:workflow}
-\end{figure}
-
 \subsubsection{Self-adaptive sampling strategy}
@@
 We specified a constraint mask to prevent stacking base-pairs from being modified and, whenever necessary, reestablished their content {\em a posteriori}, as \RNAinverse has been witnessed to take some liberties with constraints masks.
-As shown in Table~\ref{table:impact_on_gc} (Supplementary material), this postprocessing does not drastically alter the \GCContent, so the glocal approach reasonably addresses the constrained \GCContent design problem.
+As shown in the supplementary material, this postprocessing does not drastically alter the \GCContent, so the glocal approach reasonably addresses the constrained \GCContent design problem.
 
@@
 
- \begin{figure*}[ht!]
+ \begin{figure*}[hb!]
  	\centering
@@
 
-\begin{figure}[ht!]
+\begin{figure}[hb!]
   \begin{center}

### 2013-07-19 Talk-ISMB/Figures/Pseudo-RNAinverse.pgf.tex
talk ismb added

%!TEX root = ../axearn.tex
\begin{tikzpicture}

 \newcommand{\BSep}{9pt}
 \newcommand{\ESep}{30pt}
 \newcommand{\HSep}{250pt}
  \tikzstyle{base}=[circle,draw,thick,inner sep=0,minimum width=20pt,fill=white]
%  \tikzstyle{basered}=[circle,draw,thick,inner sep=0,minimum width=20pt,fill=red]
  \tikzstyle{basesmall}=[circle,draw,thick,inner sep=0,minimum width=10pt,fill=white]
  \tikzstyle{basephantom}=[base,dashed]
    \tikzstyle{baseinvisible}=[base,dashed,opacity=0]
  \tikzstyle{linez}=[draw,snake=coil, segment aspect=.2,%
line after snake=0pt, 
        segment length=10pt,thick]
  \tikzstyle{lined}=[linez,draw,snake=none,thick]
  \tikzstyle{line}=[linez,draw,snake=none,thick]
  \tikzstyle{bp}=[in=90,out=90,draw,line width=1.5pt,blue,looseness=1.2]
  \tikzstyle{bptransparent}=[in=90,out=90,draw,line width=1.5pt,blue!20,looseness=1.2]  
  \tikzstyle{redbp}=[in=90,out=90,draw,line width=1.5pt,red,looseness=1.2]
  \tikzstyle{inblock}=[trapezium,trapezium angle=83, fill=blue!20, draw=blue!40,line width=1.5pt, inner sep=0]
    \tikzstyle{inblockgray}=[trapezium,trapezium angle=83, fill=lightgray!20, draw=lightgray,line width=1.5pt, inner sep=0]
   \tikzstyle{outblock}=[trapezium,trapezium angle=83, fill=red!50, draw=red!40,line width=1.5pt, inner sep=0]
  \tikzstyle{lbl}=[inner sep=0]
  \tikzstyle{arr}=[line width=1.3pt,->]
%

\begin{scope}[yshift=185pt]
  \node[] (tot) at (0,0) {\Huge Sequence:};
	\visible<2-2>{
		\node[right=\ESep of tot, baseinvisible] (iso-a-1) {$nt^1$};
		\foreach \i in {2,...,17}{
			\pgfmathtruncatemacro\j{\i - 1}
			\ifthenelse{\i>3 \and \i<9}{
					\node[right=\BSep of iso-a-\j, base] (iso-a-\i) {$nt^{\i}$};		
					\ifthenelse{\i>4}
						{\path[lined] (iso-a-\j) -- (iso-a-\i);}
						{}
				}{
					\node[right=\BSep of iso-a-\j, baseinvisible] (iso-a-\i) {$nt^{\i}$};
				}
		}
			\path[linez] (iso-a-1) to (iso-a-4);
			\path[linez] (iso-a-8) to (iso-a-17);
			\draw[bp] (iso-a-4) to (iso-a-8);
	}

	\visible<3>{
		\node[right=\ESep of tot, baseinvisible] (iso-a-1) {$nt^1$};
		\foreach \i in {2,...,17}{
			\pgfmathtruncatemacro\j{\i - 1}
			\ifthenelse{\i>3 \and \i<9}{
				\ifthenelse{\i=4}{\node[right=\BSep of iso-a-\j, base,fill=blue] (iso-a-\i) {$nt^{\i}$};}{				        
				\ifthenelse{\i=5}{\node[right=\BSep of iso-a-\j, base,fill=green] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=6}{\node[right=\BSep of iso-a-\j, base,fill=yellow] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=7}{\node[right=\BSep of iso-a-\j, base,fill=blue] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=8}{\node[right=\BSep of iso-a-\j, base,fill=red] (iso-a-\i) {$nt^{\i}$};}{
				}}}}}
				\ifthenelse{\i>4}
					{\path[lined] (iso-a-\j) -- (iso-a-\i);}
					{}
			}{
				\node[right=\BSep of iso-a-\j, baseinvisible] (iso-a-\i) {$nt^{\i}$};
			}
		}
			\path[linez] (iso-a-1) to (iso-a-4);
			\path[linez] (iso-a-8) to (iso-a-17);
			\draw[bp] (iso-a-4) to (iso-a-8);
	}

	\visible<4>{
		\node[right=\ESep of tot, baseinvisible] (iso-a-1) {$nt^1$};
		\foreach \i in {2,...,17}{
			\pgfmathtruncatemacro\j{\i - 1}
			\ifthenelse{\i>3 \and \i<9}{
				\ifthenelse{\i=4}{\node[right=\BSep of iso-a-\j, base,fill=blue] (iso-a-\i) {$nt^{\i}$};}{				        
				\ifthenelse{\i=5}{\node[right=\BSep of iso-a-\j, base,fill=green!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=6}{\node[right=\BSep of iso-a-\j, base,fill=yellow!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=7}{\node[right=\BSep of iso-a-\j, base,fill=blue!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=8}{\node[right=\BSep of iso-a-\j, base,fill=red] (iso-a-\i) {$nt^{\i}$};}{
				}}}}}
				\ifthenelse{\i>4}
					{\path[lined] (iso-a-\j) -- (iso-a-\i);}
					{}
			}{
				\node[right=\BSep of iso-a-\j, baseinvisible] (iso-a-\i) {$nt^{\i}$};
			}
		}
			\path[linez] (iso-a-1) to (iso-a-4);
			\path[linez] (iso-a-8) to (iso-a-17);
			\draw[bp] (iso-a-4) to (iso-a-8);
	}

	\visible<5>{
		\node[right=\ESep of tot, baseinvisible] (iso-a-1) {$nt^1$};
		\foreach \i in {2,...,17}{
			\pgfmathtruncatemacro\j{\i - 1}
			\ifthenelse{\i>3 \and \i<9}{
				\ifthenelse{\i=4}{\node[right=\BSep of iso-a-\j, base,fill=green] (iso-a-\i) {$nt^{\i}$};}{				        
				\ifthenelse{\i=5}{\node[right=\BSep of iso-a-\j, base,fill=green!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=6}{\node[right=\BSep of iso-a-\j, base,fill=yellow!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=7}{\node[right=\BSep of iso-a-\j, base,fill=blue!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=8}{\node[right=\BSep of iso-a-\j, base,fill=yellow] (iso-a-\i) {$nt^{\i}$};}{
				}}}}}
				\ifthenelse{\i>4}
					{\path[lined] (iso-a-\j) -- (iso-a-\i);}
					{}
			}{
				\node[right=\BSep of iso-a-\j, baseinvisible] (iso-a-\i) {$nt^{\i}$};
			}
		}
			\path[linez] (iso-a-1) to (iso-a-4);
			\path[linez] (iso-a-8) to (iso-a-17);
			\draw[bp] (iso-a-4) to (iso-a-8);
	}

	\visible<6>{
		\node[right=\ESep of tot, baseinvisible] (iso-a-1) {$nt^1$};
		\foreach \i in {2,...,17}{
			\pgfmathtruncatemacro\j{\i - 1}
			\ifthenelse{\i>3 \and \i<16}{
				\ifthenelse{\i=4}{\node[right=\BSep of iso-a-\j, base,fill=green!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=5}{\node[right=\BSep of iso-a-\j, base,fill=green!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=6}{\node[right=\BSep of iso-a-\j, base,fill=yellow!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=7}{\node[right=\BSep of iso-a-\j, base,fill=blue!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=8}{\node[right=\BSep of iso-a-\j, base,fill=yellow!20] (iso-a-\i) {$nt^{\i}$};}{
				\node[right=\BSep of iso-a-\j, base] (iso-a-\i) {$nt^{\i}$};
				}}}}}
				\ifthenelse{\i>4}
					{\path[lined] (iso-a-\j) -- (iso-a-\i);}
					{}
			}{
				\node[right=\BSep of iso-a-\j, baseinvisible] (iso-a-\i) {$nt^{\i}$};
			}
		}
			\path[linez] (iso-a-1) to (iso-a-4);
			\path[linez] (iso-a-15) to (iso-a-17);
			\draw[bptransparent] (iso-a-4) to (iso-a-8);
			\draw[bp] (iso-a-9) to (iso-a-15);
			\draw[bp] (iso-a-10) to (iso-a-14);
	}

	\visible<7>{
		\node[right=\ESep of tot, baseinvisible] (iso-a-1) {$nt^1$};
		\foreach \i in {2,...,17}{
			\pgfmathtruncatemacro\j{\i - 1}
			\ifthenelse{\i>3 \and \i<16}{
				\ifthenelse{\i=4}{\node[right=\BSep of iso-a-\j, base,fill=green!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=5}{\node[right=\BSep of iso-a-\j, base,fill=green!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=6}{\node[right=\BSep of iso-a-\j, base,fill=yellow!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=7}{\node[right=\BSep of iso-a-\j, base,fill=blue!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=8}{\node[right=\BSep of iso-a-\j, base,fill=yellow!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=9}{\node[right=\BSep of iso-a-\j, base,fill=yellow] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=10}{\node[right=\BSep of iso-a-\j, base,fill=red] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=11}{\node[right=\BSep of iso-a-\j, base,fill=red] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=12}{\node[right=\BSep of iso-a-\j, base,fill=blue] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=13}{\node[right=\BSep of iso-a-\j, base,fill=yellow] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=14}{\node[right=\BSep of iso-a-\j, base,fill=green] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=15}{\node[right=\BSep of iso-a-\j, base,fill=green] (iso-a-\i) {$nt^{\i}$};}{
				}}}}}}}}}}}}
				\ifthenelse{\i>4}
					{\path[lined] (iso-a-\j) -- (iso-a-\i);}
					{}
			}{
				\node[right=\BSep of iso-a-\j, baseinvisible] (iso-a-\i) {$nt^{\i}$};
			}
		}
			\path[linez] (iso-a-1) to (iso-a-4);
			\path[linez] (iso-a-15) to (iso-a-17);
			\draw[bptransparent] (iso-a-4) to (iso-a-8);
			\draw[bp] (iso-a-9) to (iso-a-15);
			\draw[bp] (iso-a-10) to (iso-a-14);
	}

	\visible<8>{
		\node[right=\ESep of tot, baseinvisible] (iso-a-1) {$nt^1$};
		\foreach \i in {2,...,17}{
			\pgfmathtruncatemacro\j{\i - 1}
			\ifthenelse{\i>3 \and \i<16}{
				\ifthenelse{\i=4}{\node[right=\BSep of iso-a-\j, base,fill=green!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=5}{\node[right=\BSep of iso-a-\j, base,fill=green!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=6}{\node[right=\BSep of iso-a-\j, base,fill=yellow!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=7}{\node[right=\BSep of iso-a-\j, base,fill=blue!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=8}{\node[right=\BSep of iso-a-\j, base,fill=yellow!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=9}{\node[right=\BSep of iso-a-\j, base,fill=red] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=10}{\node[right=\BSep of iso-a-\j, base,fill=red!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=11}{\node[right=\BSep of iso-a-\j, base,fill=red!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=12}{\node[right=\BSep of iso-a-\j, base,fill=blue!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=13}{\node[right=\BSep of iso-a-\j, base,fill=yellow!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=14}{\node[right=\BSep of iso-a-\j, base,fill=green!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=15}{\node[right=\BSep of iso-a-\j, base,fill=blue] (iso-a-\i) {$nt^{\i}$};}{
				}}}}}}}}}}}}
				\ifthenelse{\i>4}
					{\path[lined] (iso-a-\j) -- (iso-a-\i);}
					{}
			}{
				\node[right=\BSep of iso-a-\j, baseinvisible] (iso-a-\i) {$nt^{\i}$};
			}
		}
			\path[linez] (iso-a-1) to (iso-a-4);
			\path[linez] (iso-a-15) to (iso-a-17);
			\draw[bptransparent] (iso-a-4) to (iso-a-8);
			\draw[bp] (iso-a-9) to (iso-a-15);
			\draw[bp] (iso-a-10) to (iso-a-14);
	}

	\visible<9>{
		\node[right=\ESep of tot, baseinvisible] (iso-a-1) {$nt^1$};
		\foreach \i in {2,...,17}{
			\pgfmathtruncatemacro\j{\i - 1}
			\ifthenelse{\i>3 \and \i<16}{
				\ifthenelse{\i=4}{\node[right=\BSep of iso-a-\j, base,fill=green!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=5}{\node[right=\BSep of iso-a-\j, base,fill=green!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=6}{\node[right=\BSep of iso-a-\j, base,fill=yellow!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=7}{\node[right=\BSep of iso-a-\j, base,fill=blue!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=8}{\node[right=\BSep of iso-a-\j, base,fill=yellow!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=9}{\node[right=\BSep of iso-a-\j, base,fill=red!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=10}{\node[right=\BSep of iso-a-\j, base,fill=red!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=11}{\node[right=\BSep of iso-a-\j, base,fill=red!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=12}{\node[right=\BSep of iso-a-\j, base,fill=blue!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=13}{\node[right=\BSep of iso-a-\j, base,fill=yellow!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=14}{\node[right=\BSep of iso-a-\j, base,fill=green!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=15}{\node[right=\BSep of iso-a-\j, base,fill=blue!20] (iso-a-\i) {$nt^{\i}$};}{
				}}}}}}}}}}}}
				\ifthenelse{\i>4}
					{\path[lined] (iso-a-\j) -- (iso-a-\i);}
					{}
			}{
				\node[right=\BSep of iso-a-\j, baseinvisible] (iso-a-\i) {$nt^{\i}$};
			}
		}
			\path[linez] (iso-a-1) to (iso-a-4);
			\path[linez] (iso-a-15) to (iso-a-17);
			\draw[bp] (iso-a-4) to (iso-a-8);
			\draw[bp] (iso-a-9) to (iso-a-15);
			\draw[bp] (iso-a-10) to (iso-a-14);
	}

	\visible<10>{
		\node[right=\ESep of tot, baseinvisible] (iso-a-1) {$nt^1$};
		\foreach \i in {2,...,17}{
			\pgfmathtruncatemacro\j{\i - 1}
			\ifthenelse{\i>3 \and \i<16}{
				\ifthenelse{\i=4}{\node[right=\BSep of iso-a-\j, base,fill=blue] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=5}{\node[right=\BSep of iso-a-\j, base,fill=green!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=6}{\node[right=\BSep of iso-a-\j, base,fill=yellow!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=7}{\node[right=\BSep of iso-a-\j, base,fill=blue!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=8}{\node[right=\BSep of iso-a-\j, base,fill=yellow] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=9}{\node[right=\BSep of iso-a-\j, base,fill=green] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=10}{\node[right=\BSep of iso-a-\j, base,fill=red!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=11}{\node[right=\BSep of iso-a-\j, base,fill=red!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=12}{\node[right=\BSep of iso-a-\j, base,fill=blue!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=13}{\node[right=\BSep of iso-a-\j, base,fill=yellow!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=14}{\node[right=\BSep of iso-a-\j, base,fill=yellow] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=15}{\node[right=\BSep of iso-a-\j, base,fill=blue] (iso-a-\i) {$nt^{\i}$};}{
				}}}}}}}}}}}}
				\ifthenelse{\i>4}
					{\path[lined] (iso-a-\j) -- (iso-a-\i);}
					{}
			}{
				\node[right=\BSep of iso-a-\j, baseinvisible] (iso-a-\i) {$nt^{\i}$};
			}
		}
			\path[linez] (iso-a-1) to (iso-a-4);
			\path[linez] (iso-a-15) to (iso-a-17);
			\draw[bp] (iso-a-4) to (iso-a-8);
			\draw[bp] (iso-a-9) to (iso-a-15);
			\draw[bp] (iso-a-10) to (iso-a-14);
	}

	\visible<11>{
		\node[right=\ESep of tot, baseinvisible] (iso-a-1) {$nt^1$};
		\foreach \i in {2,...,17}{
			\pgfmathtruncatemacro\j{\i - 1}
			\ifthenelse{\i>1}{
				\ifthenelse{\i=4}{\node[right=\BSep of iso-a-\j, base,fill=blue!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=5}{\node[right=\BSep of iso-a-\j, base,fill=green!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=6}{\node[right=\BSep of iso-a-\j, base,fill=yellow!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=7}{\node[right=\BSep of iso-a-\j, base,fill=blue!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=8}{\node[right=\BSep of iso-a-\j, base,fill=yellow!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=9}{\node[right=\BSep of iso-a-\j, base,fill=green!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=10}{\node[right=\BSep of iso-a-\j, base,fill=red!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=11}{\node[right=\BSep of iso-a-\j, base,fill=red!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=12}{\node[right=\BSep of iso-a-\j, base,fill=blue!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=13}{\node[right=\BSep of iso-a-\j, base,fill=yellow!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=14}{\node[right=\BSep of iso-a-\j, base,fill=yellow!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=15}{\node[right=\BSep of iso-a-\j, base,fill=blue!20] (iso-a-\i) {$nt^{\i}$};}{
					\node[right=\BSep of iso-a-\j, base] (iso-a-\i) {$nt^{\i}$};
				}}}}}}}}}}}}
			}{
				\node[right=\BSep of iso-a-\j, baseinvisible] (iso-a-\i) {$nt^{\i}$};
			}
			\ifthenelse{\i>2}{\path[lined] (iso-a-\j) -- (iso-a-\i);}{}
		}
			\path[linez] (iso-a-1) to (iso-a-2);
			\draw[bp] (iso-a-2) to (iso-a-17);			
			\draw[bp] (iso-a-3) to (iso-a-16);
			\draw[bp] (iso-a-4) to (iso-a-8);
			\draw[bp] (iso-a-9) to (iso-a-15);
			\draw[bp] (iso-a-10) to (iso-a-14);
	}

	\visible<12>{
		\node[right=\ESep of tot, baseinvisible] (iso-a-1) {$nt^1$};
		\foreach \i in {2,...,17}{
			\pgfmathtruncatemacro\j{\i - 1}
			\ifthenelse{\i>1}{
				\ifthenelse{\i=2}{\node[right=\BSep of iso-a-\j, base,fill=yellow] (iso-a-\i) {$nt^{\i}$};}{				        
				\ifthenelse{\i=3}{\node[right=\BSep of iso-a-\j, base,fill=green] (iso-a-\i) {$nt^{\i}$};}{				        
				\ifthenelse{\i=4}{\node[right=\BSep of iso-a-\j, base,fill=blue!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=5}{\node[right=\BSep of iso-a-\j, base,fill=red] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=6}{\node[right=\BSep of iso-a-\j, base,fill=yellow!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=7}{\node[right=\BSep of iso-a-\j, base,fill=green] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=8}{\node[right=\BSep of iso-a-\j, base,fill=yellow!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=9}{\node[right=\BSep of iso-a-\j, base,fill=green!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=10}{\node[right=\BSep of iso-a-\j, base,fill=blue] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=11}{\node[right=\BSep of iso-a-\j, base,fill=red!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=12}{\node[right=\BSep of iso-a-\j, base,fill=blue!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=13}{\node[right=\BSep of iso-a-\j, base,fill=yellow!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=14}{\node[right=\BSep of iso-a-\j, base,fill=yellow!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=15}{\node[right=\BSep of iso-a-\j, base,fill=green] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=16}{\node[right=\BSep of iso-a-\j, base,fill=blue] (iso-a-\i) {$nt^{\i}$};}{				        
				\ifthenelse{\i=17}{\node[right=\BSep of iso-a-\j, base,fill=green] (iso-a-\i) {$nt^{\i}$};}{				        
					\node[right=\BSep of iso-a-\j, base] (iso-a-\i) {$nt^{\i}$};
				}}}}}}}}}}}}}}}}
			}{
				\node[right=\BSep of iso-a-\j, baseinvisible] (iso-a-\i) {$nt^{\i}$};
			}
			\ifthenelse{\i>2}{\path[lined] (iso-a-\j) -- (iso-a-\i);}{}
		}
			\path[linez] (iso-a-1) to (iso-a-2);
			\draw[bp] (iso-a-2) to (iso-a-17);			
			\draw[bp] (iso-a-3) to (iso-a-16);
			\draw[bp] (iso-a-4) to (iso-a-8);
			\draw[bp] (iso-a-9) to (iso-a-15);
			\draw[bp] (iso-a-10) to (iso-a-14);
	}

	\visible<13>{
		\node[right=\ESep of tot, baseinvisible] (iso-a-1) {$nt^1$};
		\foreach \i in {2,...,17}{
			\pgfmathtruncatemacro\j{\i - 1}
			\ifthenelse{\i>1}{
				\ifthenelse{\i=2}{\node[right=\BSep of iso-a-\j, base,fill=yellow!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=3}{\node[right=\BSep of iso-a-\j, base,fill=green!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=4}{\node[right=\BSep of iso-a-\j, base,fill=blue!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=5}{\node[right=\BSep of iso-a-\j, base,fill=green] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=6}{\node[right=\BSep of iso-a-\j, base,fill=yellow!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=7}{\node[right=\BSep of iso-a-\j, base,fill=red] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=8}{\node[right=\BSep of iso-a-\j, base,fill=yellow!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=9}{\node[right=\BSep of iso-a-\j, base,fill=green!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=10}{\node[right=\BSep of iso-a-\j, base,fill=blue!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=11}{\node[right=\BSep of iso-a-\j, base,fill=red!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=12}{\node[right=\BSep of iso-a-\j, base,fill=blue!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=13}{\node[right=\BSep of iso-a-\j, base,fill=yellow!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=14}{\node[right=\BSep of iso-a-\j, base,fill=yellow!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=15}{\node[right=\BSep of iso-a-\j, base,fill=green!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=16}{\node[right=\BSep of iso-a-\j, base,fill=red] (iso-a-\i) {$nt^{\i}$};}{				        
				\ifthenelse{\i=17}{\node[right=\BSep of iso-a-\j, base,fill=green!20] (iso-a-\i) {$nt^{\i}$};}{
					\node[right=\BSep of iso-a-\j, base] (iso-a-\i) {$nt^{\i}$};
				}}}}}}}}}}}}}}}}
			}{
				\node[right=\BSep of iso-a-\j, baseinvisible] (iso-a-\i) {$nt^{\i}$};
			}
			\ifthenelse{\i>2}{\path[lined] (iso-a-\j) -- (iso-a-\i);}{}
		}
			\path[linez] (iso-a-1) to (iso-a-2);
			\draw[bp] (iso-a-2) to (iso-a-17);			
			\draw[bp] (iso-a-3) to (iso-a-16);
			\draw[bp] (iso-a-4) to (iso-a-8);
			\draw[bp] (iso-a-9) to (iso-a-15);
			\draw[bp] (iso-a-10) to (iso-a-14);
	}

	\visible<14>{
		\node[right=\ESep of tot, base] (iso-a-1) {$nt^1$};
		\foreach \i in {2,...,17}{
			\pgfmathtruncatemacro\j{\i - 1}
			\ifthenelse{\i>1}{
				\ifthenelse{\i=2}{\node[right=\BSep of iso-a-\j, base,fill=yellow!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=3}{\node[right=\BSep of iso-a-\j, base,fill=green!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=4}{\node[right=\BSep of iso-a-\j, base,fill=blue!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=5}{\node[right=\BSep of iso-a-\j, base,fill=green!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=6}{\node[right=\BSep of iso-a-\j, base,fill=yellow!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=7}{\node[right=\BSep of iso-a-\j, base,fill=red!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=8}{\node[right=\BSep of iso-a-\j, base,fill=yellow!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=9}{\node[right=\BSep of iso-a-\j, base,fill=green!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=10}{\node[right=\BSep of iso-a-\j, base,fill=blue!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=11}{\node[right=\BSep of iso-a-\j, base,fill=red!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=12}{\node[right=\BSep of iso-a-\j, base,fill=blue!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=13}{\node[right=\BSep of iso-a-\j, base,fill=yellow!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=14}{\node[right=\BSep of iso-a-\j, base,fill=yellow!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=15}{\node[right=\BSep of iso-a-\j, base,fill=green!20] (iso-a-\i) {$nt^{\i}$};}{
				\ifthenelse{\i=16}{\node[right=\BSep of iso-a-\j, base,fill=red!20] (iso-a-\i) {$nt^{\i}$};}{				        
				\ifthenelse{\i=17}{\node[right=\BSep of iso-a-\j, base,fill=green!20] (iso-a-\i) {$nt^{\i}$};}{					\node[right=\BSep of iso-a-\j, base] (iso-a-\i) {$nt^{\i}$};
				}}}}}}}}}}}}}}}}
			}{
				\node[right=\BSep of iso-a-\j, baseinvisible] (iso-a-\i) {$nt^{\i}$};
			}
			\ifthenelse{\i>1}{\path[lined] (iso-a-\j) -- (iso-a-\i);}{}
		}
			\draw[bp] (iso-a-2) to (iso-a-17);			
			\draw[bp] (iso-a-3) to (iso-a-16);
			\draw[bp] (iso-a-4) to (iso-a-8);
			\draw[bp] (iso-a-9) to (iso-a-15);
			\draw[bp] (iso-a-10) to (iso-a-14);
	}

\end{scope}

\begin{scope}[yshift=-20pt, xshift=70pt]
	\node[baseinvisible] (a-0) at (0,0) {};

	\foreach \i in {1,...,17}{
		\pgfmathtruncatemacro\j{\i - 1}
		\node[right=\BSep of a-\j, base] (a-\i) {};		
		\node[lbl,below=3pt of a-\i] (i-\i) {$\i$};
		\ifthenelse{\i>1}{\path[lined] (a-\j) -- (a-\i);}{}
	}

	\begin{pgfonlayer}{background}	
		\visible<1>{
			\coordinate (northr1) at ($ (a-9.north) + (0,20pt) $);
			 \node[rectangle,inner sep=2pt,fit=(a-1.west) (northr1) (a-17.east) (i-1) (i-17)] (r1) {};
			 \path[inblock]   (r1.south west) to  (r1.south east) to (r1.north east) 		to[out=90,in=90] (r1.north west) to (r1.south west) ;
			 \draw[bp] (a-2) to (a-17);
			 \draw[bp] (a-3) to (a-16);
			 \draw[bp] (a-4) to (a-8);
			 \draw[bp] (a-9) to (a-15);
			 \draw[bp] (a-10) to (a-14);
  		}
	\end{pgfonlayer}{background}

	\begin{pgfonlayer}{background}	
		\visible<2-5>{
			\coordinate (northr1) at ($ (a-5.north) + (0,20pt) $);
			\node[rectangle,inner sep=2pt,fit=(a-4.west) (northr1) (a-8.east) (i-4) (i-8)] (r1) {};
			\path[inblock]   (r1.south west) to  (r1.south east) to (r1.north east)	to[out=90,in=90] (r1.north west) to (r1.south west) ;

			\draw[bp] (a-4) to (a-8);

			\draw[bptransparent] (a-2) to (a-17);
			\draw[bptransparent] (a-3) to (a-16);
			\draw[bptransparent] (a-9) to (a-15);
			\draw[bptransparent] (a-10) to (a-14);

  		}
	\end{pgfonlayer}{background}

	\begin{pgfonlayer}{background}	
		\visible<6-8>{
			\coordinate (northr1) at ($ (a-5.north) + (0,20pt) $);
			\node[rectangle,inner sep=2pt,fit=(a-4.west) (northr1) (a-8.east) (i-4) (i-8)] (r1) {};
			\path[inblockgray]   (r1.south west) to  (r1.south east) to (r1.north east)	to[out=90,in=90] (r1.north west) to (r1.south west) ;

			\coordinate (northr1) at ($ (a-12.north) + (0,20pt) $);
			\node[rectangle,inner sep=2pt,fit=(a-9.west) (northr1) (a-15.east) (i-9) (i-15)] (r1) {};
			\path[inblock]   (r1.south west) to  (r1.south east) to (r1.north east)	to[out=90,in=90] (r1.north west) to (r1.south west) ;

			\draw[bptransparent] (a-2) to (a-17);
			\draw[bptransparent] (a-3) to (a-16);
			\draw[bptransparent] (a-4) to (a-8);
			\draw[bp] (a-9) to (a-15);
			\draw[bp] (a-10) to (a-14);			  
  		}
	\end{pgfonlayer}{background}

	\begin{pgfonlayer}{background}	
		\visible<9-10>{
			\coordinate (northr1) at ($ (a-9.north) + (0,20pt) $);
			\node[rectangle,inner sep=2pt,fit=(a-4.west) (northr1) (a-15.east) (i-4) (i-15)] (r1) {};
			\path[inblock]   (r1.south west) to  (r1.south east) to (r1.north east)	to[out=90,in=90] (r1.north west) to (r1.south west) ;

			\draw[bptransparent] (a-2) to (a-17);
			\draw[bptransparent] (a-3) to (a-16);
			\draw[bp] (a-4) to (a-8);
			\draw[bp] (a-9) to (a-15);
			\draw[bp] (a-10) to (a-14);			  
  		}
	\end{pgfonlayer}{background}

	\begin{pgfonlayer}{background}	
		\visible<11-13>{
			\coordinate (northr1) at ($ (a-9.north) + (0,20pt) $);
			\node[rectangle,inner sep=2pt,fit=(a-2.west) (northr1) (a-17.east) (i-2) (i-17)] (r1) {};
			\path[inblock]   (r1.south west) to  (r1.south east) to (r1.north east)	to[out=90,in=90] (r1.north west) to (r1.south west) ;

			\draw[bp] (a-2) to (a-17);
			\draw[bp] (a-3) to (a-16);
			\draw[bp] (a-4) to (a-8);
			\draw[bp] (a-9) to (a-15);
			\draw[bp] (a-10) to (a-14);			  
  		}
	\end{pgfonlayer}{background}

	\begin{pgfonlayer}{background}	
		\visible<14>{
			\coordinate (northr1) at ($ (a-9.north) + (0,20pt) $);
			\node[rectangle,inner sep=2pt,fit=(a-1.west) (northr1) (a-17.east) (i-1) (i-17)] (r1) {};
			\path[inblock]   (r1.south west) to  (r1.south east) to (r1.north east)	to[out=90,in=90] (r1.north west) to (r1.south west) ;
			\draw[bp] (a-2) to (a-17);
			\draw[bp] (a-3) to (a-16);
			\draw[bp] (a-4) to (a-8);
			\draw[bp] (a-9) to (a-15);
			\draw[bp] (a-10) to (a-14);			  
  		}
	\end{pgfonlayer}{background}

\end{scope}

\end{tikzpicture}

### 2013-07-19 Talk-ISMB/Raw_data/make_graph_gc.py

@@ -1,0 +1,82 @@
+import os
+import sys
+import numpy as np
+from matplotlib import rc,use
+use('pgf')
+#rc('font',**{'family':'sans-serif','sans-serif':['Helvetica']})
+rc('text', usetex=True)
+import matplotlib.mlab as mlab
+import matplotlib.pyplot as plt
+from scipy import integrate
+
+COLORS = ['r','b','g','y','k']
+
+RAW_DATA = 'dump_gc_data.txt'
+BIN_WIDTH = 0.005
+
+def get_bin(x,width):
+    x = int(x*10000)
+    width = int(width*10000)
+    bin = x/width
+    return bin
+
+def make_list_gc_per_round():
+    raw_data = [x.strip() for x in open(RAW_DATA)]
+    l_data = []
+    for i,line in enumerate(raw_data):
+        if i % 2 == 0:
+            weight = float(line)
+        else:
+            gc_contents = line.split()
+            l_data.append((weight, gc_contents))
+    return l_data
+
+def cluster_data(l_data,width):
+    pass
+
+def main():
+    MAX_Y =1175 
+    l_data = make_list_gc_per_round()
+    fig = plt.figure()
+    ax = fig.add_subplot(111)
+    ax.set_title(r'Target 15% GC')
+    ax.set_xlabel('GC content')
+    ax.set_ylabel('Quantity')
+    ax.set_ylim((0,MAX_Y))
+    ax.set_xlim((0,1))
+    ax.vlines(0.10,0,MAX_Y,color='k',linestyles='dashed')
+    ax.vlines(0.20,0,MAX_Y,color='k',linestyles='dashed')
+    #ax.get_yaxis().set_ticks([])
+    total_generated = 0
+    
+    for i,data in enumerate(l_data):
+        print i
+        weight = data[0]
+        a = np.array([float(x) for x in data[1]])
+        sigma,mu = np.std(a),np.average(a)
+        x = np.linspace(0,1,100)
+        ax.plot(x,130*mlab.normpdf(x,mu,sigma),COLORS[i],
+                label= '$\Pr(G\mid C) = %0.4f$' % (2*weight),linewidth=6)
+        total_generated += 1000*integrate.quad(lambda x: mlab.normpdf(x,mu,sigma),
+                                          0.1,0.2)[0]
+        total_generated = int(total_generated)
+        #total_generated += max([mlab.normpdf(x,mu,sigma) for x in 
[…]

### 2013-07-21 Talk-ISMB/Raw_data/make_graph_gc.py
presentation

@@ -3,13 +3,13 @@ import sys
 import numpy as np
 from matplotlib import rc,use
 use('pgf')
-#rc('font',**{'family':'sans-serif','sans-serif':['Helvetica']})
+rc('font',**{'family':'sans-serif','sans-serif':['Helvetica']})
 rc('text', usetex=True)
 import matplotlib.mlab as mlab
 import matplotlib.pyplot as plt
 from scipy import integrate
 
-COLORS = ['r','b','g','y','k']
+COLORS = ['r','b','g','y','c']
 
 RAW_DATA = 'dump_gc_data.txt'
 BIN_WIDTH = 0.005
@@ -48,7 +48,7 @@ def main():
     ax.vlines(0.20,0,MAX_Y,color='k',linestyles='dashed')
     #ax.get_yaxis().set_ticks([])
     total_generated = 0
-    
+    previous_number = 0 
     for i,data in enumerate(l_data):
         print i
         weight = data[0]
@@ -56,7 +56,7 @@ def main():
         sigma,mu = np.std(a),np.average(a)
         x = np.linspace(0,1,100)
         ax.plot(x,130*mlab.normpdf(x,mu,sigma),COLORS[i],
-                label= '$\Pr(G\mid C) = %0.4f$' % (2*weight),linewidth=6)
+                label= '$x^{\#GC} = %0.4f$' % (weight),linewidth=6)
         total_generated += 1000*integrate.quad(lambda x: mlab.normpdf(x,mu,sigma),
                                           0.1,0.2)[0]
         total_generated = int(total_generated)
@@ -66,8 +66,9 @@ def main():
         #linestyle='dashed',label='Tot round %s' % i)
         ax.plot([0.1,0.2],[total_generated,total_generated],'%s--' % COLORS[i],
                 linewidth=5,label='Tot round %s is %s' % ((i+1),total_generated))
-        ax.fill_between([0.1,0.2],[total_generated]*2,[0]*2,facecolor=COLORS[i],
-                       alpha=0.7)
+        ax.fill_between([0.1,0.2],[total_generated]*2,[previous_number]*2,facecolor=COLORS[i],
+                       )
+        previous_number = total_generated
         ax.legend()
         #fig.show()
         fig.savefig('dist_%s.pdf' % i, dpi=1200)

### 2015-09-28 ISMB/introduction_ISMB.tex
changed energy values, new energy methods, forward

--- 2013-03-03 ISMB/introduction_ISMB.tex
@@
 RNA molecules are well tailored for such applications.
-The ease to synthesize them (they are directly transcribed from DNA) and the broad diversity of catalytic and regulation functions they can perform enable to integrate \textit{de-novo} logic  circuits within living cells \cite{Rodrigo:2012fk} or re-program existing regulation mechanisms \cite{Chang:2012uq}.
+The ease to synthesize them (they are directly transcribed from DNA) and the broad diversity of catalytic and regulation functions they can perform enable to integrate \textit{de-novo} logic  circuits within living cells \citep{Rodrigo:2012fk} or re-program existing regulation mechanisms \citep{Chang:2012uq}.
 Future advances and applications of these techniques in gene-therapy studies will strongly rely on efficient computational methods to design and re-engineer RNA molecules.
@@
 \RNAinverse pioneered RNA secondary structure design algorithms.
-It has been developed and distributed with the Vienna RNA package \cite{Hofacker:1994}.
+It has been developed and distributed with the Vienna RNA package \citep{Hofacker:1994}.
 However, only posterior experimental studies revealed the potential and practical impact of these techniques.
@@
 First a seed sequence is selected, then a local search strategy is used to mutate the seed and find, in its vicinity, a sequence with desired folding properties.
-Using this strategy, \INFORNA \cite{Busch:2006uq}, \RNASSD \cite{Aguirre-Hernandez:2007kx} and \NUPACK \cite{Zadeh:2011fk} significantly improved the performance of RNA secondary structure design algorithms.
+Using this strategy, \INFORNA \citep{Busch:2006uq}, \RNASSD \citep{Aguirre-Hernandez:2007kx} and \NUPACK \citep{Zadeh:2011fk} significantly improved the performance of RNA secondary structure design algorithms.
 More recent research studies aimed to include more constraints in the selection criteria.
-\RNAexinv focused on the design of sequences with enhanced thermodynamical and mutational robustness \cite{Avihoo:2011fk}, while \frankenstein enables to design RNA with multiple target structures \cite{Lyngso:2012vn}.
+\RNAexinv focused on the design of sequences with enhanced thermodynamical and mutational robustness \citep{Avihoo:2011fk}, while \frankenstein enables to design RNA with multiple target structures \citep{Lyngso:2012vn}.
 
-We recently introduced with \RNAensign a novel paradigm for the search strategy of RNA secondary structure design algorithm \cite{Levin:2012kx}.
-Instead of a local search approach, we proposed a global sampling strategy of the mutational landscape based on the \RNAmutants algorithm~\cite{Waldispuhl2008}.
+We recently introduced with \RNAensign a novel paradigm for the search strategy of RNA secondary structure design algorithm \citep{Levin:2012kx}.
+Instead of a local search approach, we proposed a global sampling strategy of the mutational landscape based on the \RNAmutants algorithm~\citep{Waldispuhl2008}.
 This methodology offered promising performances, but suffered from  prohibitive runtime and memory consumption.
-Following our work, Garcia-Martin and co-workers proposed \RNAiFOLD, an alternate methodology that uses constraint programming techniques to prune the mutational landscape.
+Following our work, Garcia-Martin~\emph{et al} proposed \RNAiFOLD~\citep{Garcia-Martin:2013aa}, an alternate methodology that uses constraint programming techniques to prune the mutational landscape.
 While also suffering from prohibitive running times, it is worth noting that this latter algorithm also proposes a seed-less approach to the RNA secondary structure design problem.
 
-In this paper, we introduce \ourprog, a RNA secondary structure design algorithm that benefits of our recent algorithmic advances \cite{Reinharz:2013aa} to expand our original \RNAensign algorithm \cite{Levin:2012kx}.
+In this paper, we introduce \ourprog, a RNA secondary structure design algorithm that benefits of our recent algorithmic advances \citep{Reinharz:2013aa} to expand our original \RNAensign algorithm \citep{Levin:2012kx}.
 \ourprog addresses previous limitations of \RNAensign and offers new functionalities.
@@
 Unlike \RNAensign, it does not require a seed sequence to initiate its search.
-Finally, \ourprog implements a novel algorithm using weighted sampling techniques~\cite{Bodini2010} that enables us to control, for the first time, \textit{explicitly} the \GCContent of the solution.
+Finally, \ourprog implements a novel algorithm based on weighted sampling techniques~\citep{Bodini2010} that enables us to control, for the first time, \textit{explicitly} the \GCContent of the solution.
 This functionality is essential because wild-type sequences within living organisms often present medium or low \GCContent, presumably to offer better transcription rates and/or structural plasticity.
-Previous programs do not allow to control this parameter and tend to output sequences with high \GCContent{}s.
+Previous programs do not allow to control this parameter and tend to output sequences having high \GCContent{}s~\citep{Lyngso:2012vn}.
 
-We demonstrate the performance of our algorithms on a set of real RNA structures extracted from the \RNASTRAND database \cite{andronescu2008rna}.
+We demonstrate the performance of our algorithms on a set of real RNA structures extracted from the \RNASTRAND database \citep{andronescu2008rna}.
 To complete this study, we develop an hybrid method combining our global sampling approach with local search strategies such as the one implemented in \RNAinverse.

### 2015-09-28 ISMB/supplementary_ISMB.tex
changed energy values, new energy methods, forward

@@
 However, as summarized in Table~\ref{table:impact_on_gc}, we observed that the local search heuristic used to design nucleotides in loop regions has a very limited impact on the \GCContent.
-For each class of \GCContent, we reported the observed \GCContent in the sequence initially generated by \ourprog, and the observed \GCContent after the \RNAinverse postprocessing (as defined in Section \ref{subsec:glocal_method}).
+For each class of \GCContent, we reported the observed \GCContent in the sequence initially generated by \ourprog, and the observed \GCContent after the \RNAinverse postprocessing (as defined in Section 2.3).
 Our results show that the \GCContent is relatively well conserved (less than 6\% variation), with a general tendency of the postprocessing step to bring the \GCContent back to 50\%.

### 2015-09-28 Talk-ISMB/Raw_data/make_graph_gc.py

@@ -1,0 +1,83 @@
+import os
+import sys
+import numpy as np
+from matplotlib import rc,use
+use('pgf')
+rc('font',**{'family':'sans-serif','sans-serif':['Helvetica']})
+rc('text', usetex=True)
+import matplotlib.mlab as mlab
+import matplotlib.pyplot as plt
+from scipy import integrate
+
+COLORS = ['r','b','g','y','c']
+
+RAW_DATA = 'dump_gc_data.txt'
+BIN_WIDTH = 0.005
+
+def get_bin(x,width):
+    x = int(x*10000)
+    width = int(width*10000)
+    bin = x/width
+    return bin
+
+def make_list_gc_per_round():
+    raw_data = [x.strip() for x in open(RAW_DATA)]
+    l_data = []
+    for i,line in enumerate(raw_data):
+        if i % 2 == 0:
+            weight = float(line)
+        else:
+            gc_contents = line.split()
+            l_data.append((weight, gc_contents))
+    return l_data
+
+def cluster_data(l_data,width):
+    pass
+
+def main():
+    MAX_Y =1175 
+    l_data = make_list_gc_per_round()
+    fig = plt.figure()
+    ax = fig.add_subplot(111)
+    ax.set_title(r'Target 15% GC')
+    ax.set_xlabel('GC content')
+    ax.set_ylabel('Quantity')
+    ax.set_ylim((0,MAX_Y))
+    ax.set_xlim((0,1))
+    ax.vlines(0.10,0,MAX_Y,color='k',linestyles='dashed')
+    ax.vlines(0.20,0,MAX_Y,color='k',linestyles='dashed')
+    #ax.get_yaxis().set_ticks([])
+    total_generated = 0
+    previous_number = 0 
+    for i,data in enumerate(l_data):
+        print i
+        weight = data[0]
+        a = np.array([float(x) for x in data[1]])
+        sigma,mu = np.std(a),np.average(a)
+        x = np.linspace(0,1,100)
+        ax.plot(x,130*mlab.normpdf(x,mu,sigma),COLORS[i],
+                label= '$x^{\#GC} = %0.4f$' % (weight),linewidth=6)
+        total_generated += 1000*integrate.quad(lambda x: mlab.normpdf(x,mu,sigma),
+                                          0.1,0.2)[0]
+        total_generated = int(total_generated)
+        #total_generated += max([mlab.normpdf(x,mu,sigma) for x in 
+        #                        np.linspace(0.1,0.2,1000)])
[…]

### 2015-09-28 src/IncaRNAtion.py

@@ -1,0 +1,1092 @@
+###############################################################################
+#Copyright (c) 2013, Vladimir Reinharz, Yann Ponty & Jerome Waldispuhl        #
+#All rights reserved.                                                         #
+#                                                                             #
+#Redistribution and use in source and binary forms, with or without           #
+#modification, are permitted provided that the following conditions are met:  #
+#* Redistributions of source code must retain the above copyright             #
+#notice, this list of conditions and the following disclaimer.                #
+#* Redistributions in binary form must reproduce the above copyright          #
+#notice, this list of conditions and the following disclaimer in the          #
+#documentation and/or other materials provided with the distribution.         #
+#* Neither the name of the <organization> nor the                             #
+#names of its contributors may be used to endorse or promote products         #
+#derived from this software without specific prior written permission.        #
+#                                                                             #
+#THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"  #
+#AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE    #
+#IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE   #
+#ARE DISCLAIMED. IN NO EVENT SHALL <COPYRIGHT HOLDER> BE LIABLE FOR ANY       #
+#DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES   #
+#(INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR SERVICES; #
[…]

### 2015-09-28 src/parser/IncaRNAtion_ss.py

@@ -1,0 +1,989 @@
+###############################################################################
+#Copyright (c) 2013, Vladimir Reinharz, Yann Ponty & Jerome Waldispuhl        #
+#All rights reserved.                                                         #
+#                                                                             #
+#Redistribution and use in source and binary forms, with or without           #
+#modification, are permitted provided that the following conditions are met:  #
+#* Redistributions of source code must retain the above copyright             #
+#notice, this list of conditions and the following disclaimer.                #
+#* Redistributions in binary form must reproduce the above copyright          #
+#notice, this list of conditions and the following disclaimer in the          #
+#documentation and/or other materials provided with the distribution.         #
+#* Neither the name of the <organization> nor the                             #
+#names of its contributors may be used to endorse or promote products         #
+#derived from this software without specific prior written permission.        #
+#                                                                             #
+#THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"  #
+#AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE    #
+#IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE   #
+#ARE DISCLAIMED. IN NO EVENT SHALL <COPYRIGHT HOLDER> BE LIABLE FOR ANY       #
+#DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES   #
+#(INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR SERVICES; #
+#LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION) HOWEVER CAUSED AND  #
+#ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT LIABILITY, OR TORT   #
+#(INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY OUT OF THE USE OF THIS#
[…]

### 2015-09-28 src/parser/ss_decomposition.py

@@ -1,0 +1,128 @@
+from pprint import pprint
+
+
+def nb_strand(nbs):
+    nb = 1 
+    for x in sorted(nbs)[:-1]:
+        if not x+1 in nbs:
+            nb +=1 
+    return nb
+
+
+def ss_to_bp(ss):
+    bp = []
+    l = []
+    for i, x in enumerate(ss):
+        if x == '(':
+            l.append(i)
+        elif x == ')':
+            bp.append((l.pop(), i))
+    return bp
+
+
+def make_tree(ss):
+    def make_sub_tree(start, end, bps):
+        t = []
+        x = start
+        while x <= end:
+            flag_aft_bp = False
+
+            if x in bps:
+                y, z = x, bps[x]
+                while True:
+                    if y+1 in bps and z-1 == bps[y+1]:
+                        y += 1
+                        z -= 1
+                    else:
+                        t.append((x,y,z,bps[x], make_sub_tree(y+1,z-1,bps)))
+                        x = bps[x] + 1
+                        flag_aft_bp = True
+                        break
+            if x > end:
+                break
+            if not flag_aft_bp:
+                t.append(x)
+                x += 1
+        return t
+    s = 0
+    e = len(ss)-1
+    bps = ss_to_bp(ss)
+    d_bps = {x[0]:x[1] for x in bps}
+    return [-1] + make_sub_tree(s,e,d_bps) + [len(ss)]
+
+
+def make_components(tree):
+    l_components = []
+    this_comp = set()
+    sub_comp = []
+    flag = False
+    if not tree:
+        return []
+    if tree[0] == -1:
+        flag = True
+        e = tree[-1]
+        tree = tree[1:-1]
+
+    for x in tree:
+        if isinstance(x, tuple):
+            if not flag or not (x[0] == 0 and x[3] == e-1):
+                this_comp.add(x[0])
+                this_comp.add(x[3])
+            stack = [z for z in range(x[0], x[1]+1)]
+            stack.extend(z for z in range(x[2], x[3]+1))
+            l_components.append(set(stack))
+            sub_comp.extend(make_components(x[-1]))
+            sub_comp[-1].add(x[1])
+            sub_comp[-1].add(x[2])
+        else:
+            this_comp.add(x)
+    l_components.extend(sub_comp)
+    if this_comp:
+        l_components.append(this_comp)
+    return l_components
+
+
+def decompose(ss):
+    #get the decomposition of the ss in elements
[…]

### 2015-09-28 src/parser/values/parser.py

@@ -1,0 +1,217 @@
+import re
+import itertools
+import sys
+
+nucleotides = ['A', 'C', 'G', 'U']
+
+
+
+#copy from original Incarnation
+IUPACBASES = {
+  'A':['A'],
+  'C':['C'],
+  'G':['G'],
+  'U':['U'],
+  'N':['A','C','G','U'],
+  'R':['A','G'],
+  'Y':['C','U'],
+  'S':['G','C'],
+  'W':['A','U'],
+  'K':['G','U'],
+  'M':['A','C'],
+  'B':['C','G','U'],
+  'D':['A','G','U'],
+  'H':['A','C','U'],
+  'V':['A','C','G']
+  }
+STACKING_ENERGY = {k:sys.maxint for k in itertools.product(
+                  IUPACBASES['N'], repeat=4)}
+#Adjust with turner04
+#The order of the nucleotides is from 5' -> 3'
+STACKING_ENERGY.update({('A', 'A', 'U', 'U'):-0.9,
+                        ('A', 'C', 'G', 'U'):-2.2,
+                        ('A', 'G', 'C', 'U'):-2.1,
+                        ('A', 'G', 'U', 'U'):-0.6,
+                        ('A', 'U', 'A', 'U'):-1.1,
+                        ('A', 'U', 'G', 'U'):-1.4,
+                        ('C', 'A', 'U', 'G'):-2.1,
+                        ('C', 'G', 'U', 'G'):-1.4,
+                        ('C', 'U', 'A', 'G'):-2.1,
+                        ('C', 'U', 'G', 'G'):-2.1,
+                        ('G', 'A', 'U', 'C'):-2.4,
+                        ('G', 'C', 'G', 'C'):-3.4,
+                        ('G', 'G', 'C', 'C'):-3.3,
+                        ('G', 'G', 'U', 'C'):-1.5,
+                        ('G', 'U', 'A', 'C'):-2.2,
+                        ('G', 'U', 'G', 'C'):-2.5,
+                        ('G', 'A', 'U', 'U'):-1.3,
+                        ('G', 'C', 'G', 'U'):-2.5,
+                        ('G', 'G', 'C', 'U'):-2.1,
+                        ('G', 'G', 'U', 'U'):-0.5,
+                        ('G', 'U', 'A', 'U'):-1.4,
+                        ('G', 'U', 'G', 'U'):1.3,
+                        ('U', 'A', 'U', 'A'):-1.3,
[…]
